\documentclass[11pt,onecolumn]{IEEEtran}
\input{orius_ieee_detailed_preamble.tex}

\title{ORIUS: A Universal Runtime Safety Layer for Physical AI\\under Degraded Observation ---\\Battery Witness Row with Bounded AV and Healthcare Validation}
\author{Pratik Niroula}
\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
Physical AI systems operating in safety-critical domains face a fundamental hazard:
\emph{degraded observation}.  When telemetry is delayed, dropped, stale, or drifted,
a controller may evaluate its action as safe with respect to the observed state while
the true physical state has already crossed a constraint boundary.  We call this the
\emph{Observation--Action Safety Gap} (OASG).

We present \textbf{ORIUS} (Observation--Reality Integrity for Universal Safety), a
typed runtime layer that wraps any upstream controller and enforces safety through a
five-stage kernel: \textbf{Detect}~degraded telemetry, \textbf{Calibrate}~the
observation-consistent state set, \textbf{Constrain}~the admissible action set,
\textbf{Shield}~unsafe candidate actions via repair or fallback, and
\textbf{Certify}~every released action with an auditable safety certificate.

The ORIUS architecture is \emph{universal by design}: the kernel, benchmark schema,
contract stack, and governance surface are domain-agnostic.  Domain specifics are
isolated in thin adapter modules that implement six typed methods.  This paper
reports the active \emph{three-domain defended lane} with one theorem-grade witness
row and two bounded promoted rows:

\begin{enumerate}[nosep]
  \item \textbf{Battery Energy Storage}: dispatch optimisation under degraded SoC
        telemetry, validated on locked CPSBench replay artifacts.  Battery remains
        the witness row with the deepest theorem-to-code-to-artifact closure; the
        promoted runtime slice achieves zero true-state violations.
  \item \textbf{Autonomous Vehicles}: ego-speed and gap prediction under degraded
        perception, validated on the full Waymo Open Motion corpus (7~validation
        shards, 1{,}975~scenarios, 80{,}975~feature steps, 9{,}348~runtime test steps,
        9{,}348~ certificates, all DC3S contracts satisfied).
  \item \textbf{Medical and Healthcare Monitoring}: bounded alert-release monitoring
        under degraded vital-sign observations, promoted through a MIMIC-backed
        healthcare row with bounded runtime and governance claims.
\end{enumerate}

All promoted rows maintain 100\% certificate-chain validity, 100\% audit completeness,
and zero contract failures on their active runtime surfaces.  All results are reproducible
from locked artifacts with SHA-256 hashes.  We report a strict defended theorem core,
the complete shift-aware adaptive uncertainty mechanism, and a three-domain synthesis
showing that the same kernel survives domain change without semantic loss while keeping
battery-specific proof depth clearly distinguished from the bounded AV and Healthcare rows.
\end{abstract}

\begin{IEEEkeywords}
runtime safety, degraded observation, conformal prediction, safety certificates,
cyber-physical systems, battery energy storage, autonomous vehicles, healthcare monitoring, OASG
\end{IEEEkeywords}

\tableofcontents
\newpage

%% ===== MAIN BODY =====
\input{detailed_sections/detailed_intro.tex}
\input{detailed_sections/detailed_problem_oasg.tex}
\input{detailed_sections/detailed_related_work.tex}
\input{detailed_sections/detailed_runtime_architecture.tex}
\input{detailed_sections/detailed_contracts_governance.tex}
\input{detailed_sections/detailed_theory_bridge.tex}
\input{detailed_sections/detailed_shift_aware.tex}
\input{detailed_sections/detailed_battery_validation.tex}
\input{detailed_sections/detailed_av_validation.tex}
\input{detailed_sections/detailed_cross_domain.tex}
\input{detailed_sections/detailed_reproducibility.tex}
\input{detailed_sections/detailed_limitations.tex}
\input{detailed_sections/detailed_conclusion.tex}

%% ===== APPENDIX =====
\appendices
\input{detailed_sections/detailed_appendix.tex}

\bibliographystyle{IEEEtran}
\bibliography{../bibliography/orius_monograph}

\end{document}
