% === Section 5: Contracts and Governance ===
\section{Contract Stack and Governance Surface}\label{sec:contracts}

Every \dcs{} step produces intermediate typed objects.  The \textbf{contract stack}
defines the obligations that each object must satisfy for the step to be valid.  The
\textbf{governance surface} aggregates per-step certificates into episode- and
campaign-level summaries that auditors can inspect.

\subsection{Typed Contract Objects}

\Cref{tab:contract_objects} lists the principal contract objects, their
associated assumptions, and their role in the kernel.

\begin{table}[h]
\caption{Typed contract objects produced by the \dcs{} kernel.}
\label{tab:contract_objects}
\centering\small
\begin{tabular}{lll}
\toprule
\textbf{Contract object} & \textbf{Assumptions} & \textbf{Role} \\
\midrule
\texttt{ObservationPacket}            & A1       & Raw telemetry + parsed state \\
\texttt{ReliabilityAssessment}        & A2, A5, A6 & Detection output: $\wt$, flags \\
\texttt{ObservationConsistentStateSet} & A2, A5   & Uncertainty envelope $\Ut$ \\
\texttt{SafetySpec}                   & A3, A4, A8 & Domain constraints + fallback \\
\texttt{SafeActionSet}               & A3, A4   & Tightened admissible actions \\
\texttt{RepairDecision}              & A4       & Shield output \\
\texttt{SafetyCertificate}           & ---      & Auditable record binding all stages \\
\bottomrule
\end{tabular}
\end{table}

\subsection{ContractVerifier}

The \texttt{ContractVerifier} is a second-pass checker that validates a
\texttt{SafetyCertificate} against a set of \emph{contract predicates}.  Each
predicate is a boolean function of the certificate's payload.  The verifier evaluates
all predicates and returns a \texttt{VerificationResult} with:
\begin{itemize}[nosep]
  \item \texttt{chain\_valid}: conjunction of all predicate outcomes.
  \item \texttt{failure\_rows}: count of certificates that failed at least one predicate.
  \item \texttt{required\_payload\_pass\_rate}: fraction of certificates with all
        required fields present.
  \item \texttt{audit\_completeness\_rate}: fraction of steps with a valid certificate.
\end{itemize}

\subsection{Contract Predicates}

The predicate set is domain-agnostic.  The following predicates are evaluated for
every certificate:

\begin{enumerate}[label=\textbf{CP\arabic*},nosep]
  \item \textbf{Monotone inflation}: the inflation factor $\alpha \ge 1.0$.
  \item \textbf{Weight bounds}: $\wt \in [0, 1]$.
  \item \textbf{Interval non-degenerate}: for every target, $\mathrm{upper} \ge
        \mathrm{lower}$.
  \item \textbf{Action within bounds}: released action satisfies all tightened
        constraints.
  \item \textbf{Certificate completeness}: all required fields are present and
        non-null.
  \item \textbf{Timestamp monotonicity}: certificate timestamps are non-decreasing
        within an episode.
\end{enumerate}

The contract predicate CP1 (monotone inflation) is the subject of Theorem~T4 and
was the focus of a bug fix validated with 200/200 trial runs at zero failures.

\subsection{Domain Instantiation Protocol}

A \texttt{DomainInstantiation} protocol class specifies the typed methods that an
adapter must implement.  The protocol is defined as a Python
\texttt{typing.Protocol} in \texttt{contracts.py}:

\begin{verbatim}
class DomainInstantiation(Protocol):
    @property
    def domain_id(self) -> str: ...
    def assess(self, obs: ObservationPacket,
               history: Sequence[...]) -> ReliabilityAssessment: ...
    def calibrate(self, state: Mapping, uncertainty: Mapping,
                  quantile: float,
                  reliability: ReliabilityAssessment)
                  -> ObservationConsistentStateSet: ...
    def constrain(self, ocss: ObservationConsistentStateSet,
                  spec: SafetySpec) -> SafeActionSet: ...
    def shield(self, candidate: Mapping, safe_set: SafeActionSet,
               spec: SafetySpec) -> RepairDecision: ...
    def certify(self, ...) -> SafetyCertificate: ...
\end{verbatim}

Any adapter that satisfies this \texttt{Protocol} can be plugged into the kernel.

\subsection{Governance Surface}\label{sec:governance_surface}

At the episode level, the governance surface aggregates certificates into a summary
\texttt{runtime\_governance\_summary.csv} with the metrics in
\cref{tab:governance_metrics}.

\begin{table}[h]
\caption{Governance summary metrics.}
\label{tab:governance_metrics}
\centering\small
\begin{tabular}{ll}
\toprule
\textbf{Metric} & \textbf{Value (Battery)} \\
\midrule
\texttt{certificate\_rows}            & 288 \\
\texttt{chain\_valid}                 & 1.0 (true) \\
\texttt{failure\_rows}                & 0 \\
\texttt{required\_payload\_pass\_rate} & 1.0 \\
\texttt{expiry\_metadata\_presence\_rate} & 0.0 \\
\texttt{expiry\_consistency\_rate}     & 1.0 \\
\texttt{audit\_completeness\_rate}     & 1.0 \\
\bottomrule
\end{tabular}
\end{table}

The battery witness row achieves 100\% chain validity on its active runtime surface.
The all-zip grouped nuPlan AV row reports a 99.992\% certificate-valid release rate
and 100\% audit completeness on 1{,}531{,}104 runtime steps.

\subsection{CertOS Verification}

The CertOS module performs an offline re-verification of the entire certificate chain.
For each active runtime certificate surface:
\begin{itemize}[nosep]
  \item \textbf{Battery}: 288 certificates verified, \texttt{chain\_valid = true},
        0 failures.
  \item \textbf{AV}: 1{,}531{,}104 all-zip grouped nuPlan certificates emitted,
        certificate-valid release rate $0.999924$,
        249 empirical runtime violations under the promoted replay contract.
\end{itemize}

\subsection{Safety Architecture Comparison}

\Cref{tab:safety_comparison} positions ORIUS/\dcs{} against five alternative
safety-layer paradigms.  The comparison focuses on structural capabilities rather
than domain-specific numbers.

\begin{table}[h]
\caption{Safety architecture comparison.  \checkmark = supported natively,
$\circ$ = partially supported, --- = not supported.}
\label{tab:safety_comparison}
\centering\scriptsize
\begin{tabular}{p{2.4cm}ccccccc}
\toprule
\textbf{Capability} & \textbf{DC3S} & \textbf{Simplex} & \textbf{GP-MPC}
  & \textbf{DRO} & \textbf{Safe RL} & \textbf{Vanilla CQR} \\
\midrule
Obs-quality aware           & \checkmark & ---  & ---  & ---  & ---  & --- \\
Reliability-adaptive margin & \checkmark & ---  & $\circ$ & ---  & ---  & --- \\
Conformal coverage guarantee& \checkmark & ---  & ---  & ---  & ---  & \checkmark \\
Domain-agnostic kernel      & \checkmark & $\circ$ & ---  & ---  & ---  & \checkmark \\
Typed adapter protocol      & \checkmark & ---  & ---  & ---  & ---  & --- \\
Certificate governance      & \checkmark & ---  & ---  & ---  & ---  & --- \\
Graceful degradation (T8)   & \checkmark & \checkmark & $\circ$ & ---  & ---  & --- \\
Formal proof programme      & \checkmark & \checkmark & $\circ$ & \checkmark & $\circ$ & $\circ$ \\
Runtime overhead $<$ 1\,ms  & \checkmark & \checkmark & ---  & ---  & ---  & \checkmark \\
Three-row validated\textsuperscript{$\dagger$} & \checkmark & ---  & ---  & ---  & ---  & --- \\
\bottomrule
\end{tabular}
\end{table}

{\footnotesize $\dagger$ Active defended lane: Battery witness row plus bounded AV and Healthcare rows with locked artifacts.}

\paragraph{Key differentiators.}
\begin{itemize}[nosep]
  \item \textbf{Simplex}~\cite{sha1998simplex,sha2001using} provides safety switching between a
        performance controller and a verified safe controller, but does not adapt
        the safety margin to telemetry reliability.
  \item \textbf{GP-MPC} (Gaussian Process Model Predictive Control) provides
        uncertainty-aware control but requires a Gaussian process model per domain,
        with $O(n^3)$ scaling precluding real-time use at large state dimensions.
  \item \textbf{DRO} (Distributionally Robust Optimisation) handles distributional
        ambiguity but does not produce per-step safety certificates or adapt to
        observation-quality signals.
  \item \textbf{Safe RL} incorporates safety constraints into the reward, but
        typically provides asymptotic guarantees rather than per-step certificates.
  \item \textbf{Vanilla CQR} achieves conformal coverage but is quality-ignorant:
        the same interval width is used regardless of telemetry degradation.
\end{itemize}

DC3S is unique in combining observation-quality awareness, reliability-adaptive margins,
a formal proof programme, domain-agnostic typed adapters, and per-step certificate
governance in a single runtime layer with sub-millisecond overhead.
