% === Section 1: Introduction ===
\section{Introduction}\label{sec:intro}

The deployment of AI-driven controllers in safety-critical physical systems---battery
energy storage, autonomous vehicles, and clinical monitoring---has
progressed from laboratory prototypes to field-deployed assets.  A pervasive but
under-addressed hazard accompanies this transition: the controller takes actions
conditioned on \emph{observations} of the true physical state, yet observations are
never perfect.  Sensors stall, telemetry packets arrive late or out-of-order,
calibration drifts silently, and network links drop entire channels.  When the
observation pipeline degrades, the gap between what the controller \emph{believes} to
be the state and what the state \emph{actually is} may grow unboundedly, and the
controller may take an action that is formally safe in the observed state space yet
catastrophically unsafe in the true state space.

We call this gap the \textbf{Observation--Action Safety Gap} (OASG).  The OASG is
distinct from standard notions of model uncertainty, adversarial perturbation, or
out-of-distribution detection.\footnote{Out-of-distribution (OOD) detectors identify
distributional mismatches but do not synthesise conservative action sets; adversarial
robustness certifies bounded input perturbations but does not handle telemetry
infrastructure faults (packet loss, reordering, drift).}  It requires a runtime-level
mechanism that:
\begin{enumerate}[nosep]
  \item Detects telemetry degradation at every control step.
  \item Quantifies the resulting \emph{set of states consistent with the degraded
        observation}.
  \item Tightens the admissible action space so that safety is guaranteed for
        every member of that set.
  \item Shields or repairs any candidate action that exits the tightened set.
  \item Issues an auditable, tamper-evident certificate for every released action.
\end{enumerate}

No existing framework---simplex architectures~\cite{sha2001using},
runtime-verification monitors~\cite{leucker2009brief}, Bayesian OOD
detectors~\cite{hendrycks2017baseline}, or conformal-prediction wrappers~\cite{romano2019conformalized}---addresses
all five requirements simultaneously while remaining \emph{domain-agnostic},
\ie~applicable to any cyber-physical domain via a thin adapter rather than a domain-specific redesign.

\subsection{The ORIUS Framework}

This paper presents \textbf{ORIUS} (Observation--Reality Integrity for Universal
Safety), a typed runtime safety layer whose five-stage kernel---\textbf{Detect},
\textbf{Calibrate}, \textbf{Constrain}, \textbf{Shield}, \textbf{Certify}, abbreviated
\dcs{}---closes the OASG for any upstream controller.  The kernel is defined over abstract
typed contracts (\cref{sec:contracts}); domain specifics are confined to an adapter
module that implements six typed methods.  A formal theorem chain (T1--T11) connects
the runtime invariants to end-to-end safety guarantees under stated assumptions
(A1--A13).

\subsection{Scope and Claims}

Unlike broader planning drafts that motivated ORIUS with additional candidate rows,
this manuscript restricts its defended empirical lane to \textbf{three active domains}
with locked, reproducible artifacts and explicit evidence tiers:

\begin{enumerate}[nosep]
  \item \textbf{Battery Energy Storage (BESS)}: dispatch optimisation under degraded
        state-of-charge telemetry on locked CPSBench replay artifacts.  Battery is the
        theorem-grade witness row and carries the deepest artifact-to-proof closure.
  \item \textbf{Autonomous Vehicles (AV)}: longitudinal prediction of ego-speed and
        relative gap under degraded perception, validated on the completed all-zip
        grouped nuPlan replay surface with explicit non-closure of full
        autonomous-driving field deployment.
  \item \textbf{Medical and Healthcare Monitoring}: bounded monitoring and alert release
        under degraded vital-sign observations.  Healthcare is promoted as a bounded
        MIMIC-backed runtime row; its claims remain narrower than the battery witness row.
\end{enumerate}

The three active rows share the same kernel binary, the same contract
stack, the same governance pipeline, and the same CertOS certificate chain.

\subsection{Contributions}

The principal contributions of this paper are:

\begin{enumerate}[nosep]
  \item A complete formal architecture for runtime safety under degraded observation,
        expressed as five typed stages with a supporting contract algebra
        (\cref{sec:architecture}).
  \item An 11-theorem proof chain connecting observation reliability to true-state
        safety guarantees (\cref{sec:theory}).
  \item A shift-aware adaptive uncertainty mechanism that dynamically adjusts
        prediction intervals without offline recalibration (\cref{sec:shift}).
  \item A bounded three-domain empirical lane:
        \begin{itemize}[nosep]
          \item \textbf{Battery witness row}: zero true-state violations on the promoted
                runtime slice, with the deepest theorem-to-artifact alignment.
          \item \textbf{AV bounded row}: 1{,}531{,}104 all-zip grouped nuPlan replay
                steps and certificates; ORIUS reduces TSVR from 0.289250 for the
                no-wrapper baseline to 0.000163 on the promoted replay row, while the
                harder relative-gap shift table remains explicitly marked as bounded.
          \item \textbf{Healthcare bounded row}: TSVR reduced from 0.194489 to 0.000000 on
                the promoted MIMIC monitoring surface, with grouped calibration and
                governance evidence reported explicitly as bounded claims.
        \end{itemize}
  \item A fully reproducible artifact bundle with SHA-256 hashes, locked
        dependencies, and a single-command pipeline (\cref{sec:reproducibility}).
\end{enumerate}

\paragraph{What this paper proves and what it does not.}
The theorem chain (T1--T11) is stated on the typed-contract algebra, but the
repo-truthful rigor posture is tracked theorem-by-theorem in the active audit
rather than flattened to ``proved in full generality.'' Empirical confirmation
is restricted to the three domains above.  In
particular:
\begin{itemize}[nosep]
  \item The Battery domain achieves $\TSVR = 0$ (zero true-state violations) and
        serves as a \emph{reference witness}: it demonstrates the full pipeline
        end-to-end with the deepest artifact-to-theorem closure.
  \item The AV domain is a \emph{bounded proof-validated} row: ORIUS is promoted only
        under the longitudinal TTC + predictive-entry-barrier contract, with bounded
        nuPlan replay artifacts and no claim of full road deployment closure.
  \item The Healthcare domain is a \emph{bounded proof-validated} row: ORIUS is promoted
        only for monitoring and alert-release semantics on the MIMIC-backed healthcare lane.
        Certificate-valid release is reported from runtime-emitted healthcare CVA on this row.
  \item Shift-awareness guarantees are bounded to the drift rates observed during
        validation; adversarial robustness under intentional telemetry attacks has
        not been evaluated.
\end{itemize}

\subsection{Paper Organisation}

\Cref{sec:problem} formalises the OASG problem.  \Cref{sec:related} surveys related
work.  \Cref{sec:architecture} details the \dcs{} kernel and typed interfaces.
\Cref{sec:contracts} covers the contract stack and governance surface.
\Cref{sec:theory} presents the theorem bridge connecting contracts to safety.
\Cref{sec:shift} introduces the shift-aware mechanism.  \Cref{sec:battery}
and~\cref{sec:av} provide the deepest per-domain empirical sections for BESS and AV,
while \Cref{sec:crossdomain} synthesises the active three-domain lane, including the
bounded Healthcare row.
\Cref{sec:reproducibility} documents the reproducibility infrastructure.
\Cref{sec:limitations} discusses limitations.  \Cref{sec:conclusion} concludes.
