% === Section 13: Conclusion ===
\section{Conclusion}\label{sec:conclusion}

We presented ORIUS, a universal runtime safety layer for physical AI systems
operating under degraded observation.  The core contribution is the \dcs{}
kernel---Detect, Calibrate, Constrain, Shield, Certify---that closes the
Observation--Action Safety Gap for any upstream controller through typed
contracts and domain-agnostic governance.

\subsection{Summary of Results}

We validated ORIUS on a three-domain defended lane with intentionally unequal
evidence tiers: Battery is the theorem-grade witness row, AV is an all-zip grouped
nuPlan replay row, and Healthcare is a bounded retrospective monitoring row.

\paragraph{Battery Energy Storage.}
\begin{itemize}[nosep]
  \item 4 controller variants $\times$ 3 fault scenarios = 12 configurations.
  \item 288 runtime steps per configuration.
  \item \textbf{Zero true-state violations} ($\TSVR = 0$) across all configurations.
  \item 288 safety certificates, all chain-valid, 100\% audit completeness.
\end{itemize}

\paragraph{Autonomous Vehicles.}
\begin{itemize}[nosep]
  \item Completed all-zip grouped nuPlan replay surface with 37{,}344 held-out
        runtime scenarios and 1{,}531{,}104 runtime steps.
  \item \textbf{TSVR reduced} from 0.2893 (baseline) to 0.000163 (ORIUS).
  \item 50.03\% intervention rate and 17.37\% fallback activation rate.
  \item 1{,}531{,}104 safety certificates, 99.992\% certificate-valid release rate,
        and 100\% audit completeness on the promoted row.
  \item Relative-gap shift coverage remains an explicit limitation; the row is not
        promoted as full closed-loop road deployment.
\end{itemize}

\paragraph{Healthcare Monitoring.}
\begin{itemize}[nosep]
  \item Bounded MIMIC-backed monitoring and alert-release row.
  \item Retrospective evidence only; no prospective clinical deployment,
        regulatory approval, or treatment recommendation is claimed.
\end{itemize}

\subsection{Key Insights}

\begin{enumerate}[nosep]
  \item \textbf{Typed transfer is practical}: The same kernel binary, contract stack,
        and governance surface handle a slow-timescale energy system, a bounded AV
        replay surface, and a retrospective healthcare monitoring row. Domain
        adaptation requires only implementing six typed adapter methods.
  \item \textbf{Degradation-proportional conservatism}: The inflation mechanism
        (\cref{eq:inflation}) provides a smooth trade-off between safety and
        performance, becoming more conservative precisely when observation quality
        degrades.
  \item \textbf{Governance is non-negotiable}: The certificate chain and audit trail
        are essential for regulatory compliance and post-incident analysis in
        safety-critical deployments.
  \item \textbf{Conformal prediction is a strong foundation}: The distribution-free
        coverage guarantee of conformal prediction, combined with the shift-aware
        adaptive mechanism, provides robust calibration across diverse scenarios.
\end{enumerate}

\subsection{Impact}

ORIUS demonstrates that a single runtime safety kernel can be reused across
different CPS semantics without rewriting the kernel itself. The camera-ready claim
is deliberately bounded: Battery supplies the deepest proof witness, nuPlan AV shows
bounded replay transfer, and Healthcare shows bounded retrospective monitoring.  The
next scientific step is not broader rhetoric; it is larger denominators, closed-loop
AV evidence, and prospective healthcare governance under the same certificate gate.
