% === Section 6: Theorem Bridge ===
\section{Theorem Bridge: From Contracts to Safety}\label{sec:theory}

This section presents the formal theorem programme that connects the \dcs{} runtime
contracts to end-to-end safety guarantees.  The programme comprises a \emph{battery
ladder} (T1--T8) proved via a concrete domain witness, and a \emph{universal
extension} (T9--T11) that lifts the guarantees to arbitrary domains satisfying the
adapter contract.

\subsection{Assumptions}\label{sec:assumptions}

The theorem programme rests on eight standing assumptions.  Each assumption is
testable and its satisfaction is verified programmatically by the test suite.

\begin{assumption}[A1: Bounded Model Error]\label{ass:A1}
The one-step prediction error of the domain model is bounded:
$|\epsilon_t| \le \epsilon_{\mathrm{model}}$ for all $t$.
\end{assumption}

\begin{assumption}[A2: Bounded Telemetry Error]\label{ass:A2}
The observation error is bounded by a function of the reliability weight:
$|\hat{x}_t - x_t| \le g(w_t)$, where $g$ is a known, monotonically
non-increasing function with $g(1) = \bar\epsilon$ (nominal noise floor).
\end{assumption}

\begin{assumption}[A3: Constraint Expressibility]\label{ass:A3}
The domain safety specification can be expressed as a finite conjunction of
state-action constraints evaluable over $\Ut$.
\end{assumption}

\begin{assumption}[A4: Known Dynamics]\label{ass:A4}
The controller can evaluate the one-step dynamics map $f(x_t, a_t)$ to compute
$\Delta(a_t)$.  A fallback action $\asafe$ exists that is safe for all states in
$\mathcal{X}$.
\end{assumption}

\begin{assumption}[A5: Monotone Bounded Inflation]\label{ass:A5}
The reliability-adaptive conformal margin $m_t$ is monotonically non-increasing in
$w_t$: lower reliability never shrinks the uncertainty set.  The base conformal
predictor achieves coverage $\ge 1 - \alpha$ on the calibration set.
\end{assumption}

\begin{assumption}[A6: Bounded Detector Lag]\label{ass:A6}
The staleness counter and drift detector respond within a bounded lag of
$\tau_{\mathrm{lag}} < \infty$ steps.
\end{assumption}

\begin{assumption}[A7: Causal Certificates]\label{ass:A7}
Each certificate uses only causally available information (no future data).
The episode horizon is finite: $T < \infty$.
\end{assumption}

\begin{assumption}[A8: Constraint Continuity]\label{ass:A8}
The domain safety constraints are continuous in the state space, so that small
perturbations of the state yield small perturbations of the constraint value.
\end{assumption}

\paragraph{Usage map.}  \Cref{tab:assumption_usage} shows which theorems depend on
which assumptions.

\begin{table}[h]
\caption{Assumption usage map across the theorem programme.}
\label{tab:assumption_usage}
\centering\small
\begin{tabular}{lcccccccc}
\toprule
\textbf{Theorem} & \textbf{A1} & \textbf{A2} & \textbf{A3} & \textbf{A4} & \textbf{A5} & \textbf{A6} & \textbf{A7} & \textbf{A8} \\
\midrule
T1 OASG Existence         &\checkmark &\checkmark &  &\checkmark &  &  &  &  \\
T2 Safety Preservation    &\checkmark &\checkmark &\checkmark &  &\checkmark &  &\checkmark &  \\
T3 ORIUS Core Bound       &\checkmark &\checkmark &  &\checkmark &\checkmark &\checkmark &\checkmark &  \\
T4 No Free Safety         &  &\checkmark &  &\checkmark &  &  &  &  \\
T5 Certificate Validity   &  &  &  &  &\checkmark &  &\checkmark &  \\
T6 Certificate Expiration &  &  &  &  &  &  &\checkmark &  \\
T7 Feasible Fallback      &  &  &  &\checkmark &  &  &  &  \\
T8 Graceful Degradation   &  &\checkmark &  &  &\checkmark &  &  &  \\
T9 Universal Impossibility&  &\checkmark &  &\checkmark &  &  &  &  \\
T10 Lower Bound           &  &\checkmark &  &  &\checkmark &  &  &  \\
T11 Typed Transfer        &  &  &\checkmark &\checkmark &\checkmark &  &\checkmark &\checkmark \\
\bottomrule
\end{tabular}
\end{table}

%% ----- Battery Ladder: T1--T4 -----
\subsection{Battery Ladder (T1--T4): The Reference Witness}

The battery domain serves as the \emph{reference witness}: it provides the deepest
closure of the proof chain because battery dynamics are scalar, linear, and fully
instrumented.

\begin{theorem}[T1: OASG Existence]\label{thm:T1}
Let $(\mathcal{X}, \mathcal{U}, f, \mathcal{S}, \hat{x})$ be a CPS satisfying
A1, A2, and~A4.  Let $\pi$ be any quality-ignorant controller.  If $\pi$ drives
the state within distance $\delta > 0$ of the safe-set boundary $\partial\mathcal{S}$
with positive frequency, and the telemetry error satisfies
$\Prob[\|\xi_t\| > \delta] > 0$, then there exists a time $t^*$ such that
\begin{equation}\label{eq:oasg_existence}
  \phi_{\mathrm{obs}}(\pi(\hat{x}_{t^*})) = 1
  \quad\text{and}\quad
  \phi_{\mathrm{true}}(\pi(\hat{x}_{t^*})) = 0.
\end{equation}
\end{theorem}

\begin{proof}[Proof sketch]
\emph{Step~1 (Boundary proximity).}
Under arbitrage-optimal dispatch, the controller drives the SoC within
$\delta_{\mathrm{bnd}} = P_{\max}\Delta t$ of the constraint boundary with
positive frequency (Lemma: optimal arbitrage is boundary-seeking).

\emph{Step~2 (Observation gap).}
Under dropout rate $d > 0$, the observation error satisfies
$\E[|\xi_t|] \ge d \cdot \sigma_{\mathrm{drift}} \cdot \sqrt{\Delta t}$.

\emph{Step~3 (Witness construction).}
The intersection of the boundary-proximity event with the observation-error tail
is non-empty with positive probability.  At any such step $t^*$, the observed
safety predicate $\phi_{\mathrm{obs}}$ evaluates to safe but the true safety
predicate $\phi_{\mathrm{true}}$ evaluates to unsafe.
\end{proof}

\paragraph{Battery instantiation.}
The true state is $x_t = \SOC_t \in [0, \Emax]$\,(MWh), the action is
$a_t = (P_t^{\mathrm{chg}}, P_t^{\mathrm{dis}})$, and the one-step dynamics are
\begin{equation}
  x_{t+1} = x_t + (\eta_c P_t^{\mathrm{chg}} - \eta_d^{-1} P_t^{\mathrm{dis}})\Delta t + \epsilon_t.
\end{equation}
The observed safety predicate evaluates the SoC constraint on $\hat{x}_t$:
$\phi_{\mathrm{obs}}(a_t) = \mathbf{1}[\SOC_{\min} \le \hat{x}_t + \Delta(a_t) \le \SOC_{\max}]$.

\paragraph{Empirical confirmation.}
Under the deterministic LP baseline with \texttt{drift\_combo} (20\% dropout),
the empirical TSVR is 3.9\% (DE) and 4.2\% (US CAISO), confirming
$\E[\TSVR] > 0$ for quality-ignorant controllers.

\begin{theorem}[T2: One-Step Safety Preservation Under the Shield]\label{thm:T2}
Under A1--A3, A5, and~A7, define the tightened safe-action set
\begin{equation}\label{eq:tightened_set}
  \Aset = \Bigl\{ a \in \mathcal{U} :
    f(\hat{x}_t, a) \oplus B(m_t) \subseteq \mathcal{S} \Bigr\},
\end{equation}
where $m_t = q_t / (w_t + \epsilon)$ is the reliability-inflated conformal margin
and $B(m_t)$ the ball of radius~$m_t$.  If $x_t \in \Ut$ and
$a_t \in \Aset$, then $f(x_t, a_t) \in \mathcal{S}$.
\end{theorem}

\begin{proof}[Proof (battery scalar case)]
Since $x_t \in \Ut = [\hat{x}_t - m_t, \hat{x}_t + m_t]$, we have
$|x_t - \hat{x}_t| \le m_t$.  The one-step dynamics give
\begin{align}
  s_{t+1} &= s_t + \eta_c P_t^{\mathrm{chg}}\Delta t - \eta_d^{-1}P_t^{\mathrm{dis}}\Delta t + \epsilon_t \notag\\
  &\ge \hat{s}_t - m_t + \eta_c P_t^{\mathrm{chg}}\Delta t - \eta_d^{-1}P_t^{\mathrm{dis}}\Delta t - \epsilon_{\mathrm{model}} \notag\\
  &\ge \SOC_{\min} + m_t - m_t - \epsilon_{\mathrm{model}} = \SOC_{\min} - \epsilon_{\mathrm{model}}.
\end{align}
Analogously for the upper bound.  Under A1, $\epsilon_{\mathrm{model}}$ is absorbed
into the margin, yielding $s_{t+1} \in [\SOC_{\min}, \SOC_{\max}]$.
\end{proof}

\begin{theorem}[T3: ORIUS Core Bound]\label{thm:T3}
Under A1, A2, A4--A7, the expected total violation count over horizon $T$ is bounded:
\begin{equation}\label{eq:core_bound}
  \boxed{\;\E[V_T] \;\le\; \alpha\,(1 - \bar{w})\,T\;}
\end{equation}
where $\alpha$ is the conformal miscoverage level and
$\bar{w} = T^{-1}\sum_{t=1}^{T} w_t$ is the episode-average reliability.
\end{theorem}

\begin{proof}
Define the per-step violation indicator $Z_t = \mathbf{1}[f(x_t, a_t^{\mathrm{rel}}) \notin \mathcal{S}]$.
By the calibration contract (A5), $\Prob[x_t \notin \Ut \mid \mathcal{H}_t] \le \alpha$.
Given the risk-envelope contract, the per-step risk budget is
$r_t = \alpha(1 - w_t)$, whence
$\Prob[Z_t = 1 \mid \mathcal{H}_t] \le \alpha(1 - w_t)$.  Summing:
\[
  \E[V_T] = \sum_{t=1}^{T}\E[Z_t]
  \le \sum_{t=1}^{T}\alpha(1-w_t)
  = \alpha\sum_{t=1}^{T}(1-w_t)
  = \alpha(1-\bar{w})T. \qedhere
\]
\end{proof}

\paragraph{Numerical example.}
With $\alpha = 0.10$, $T = 288$ (battery), and $\bar{w} = 0.83$ (moderate degradation):
$\E[V_T] \le 0.10 \times 0.17 \times 288 = 4.9$ violations.  The \emph{observed}
violation count is 0, well below the theoretical bound.

\Cref{tab:core_bound_scenarios} illustrates the bound under varying degradation.

\begin{table}[h]
\caption{ORIUS core bound $\E[V_T] \le \alpha(1-\bar{w})T$ under varying
degradation scenarios ($\alpha = 0.10$, $T = 288$).}
\label{tab:core_bound_scenarios}
\centering\small
\begin{tabular}{lrrl}
\toprule
\textbf{Scenario} & $\bar{w}$ & $\E[V_T]$ bound & \textbf{TSVR bound (\%)} \\
\midrule
Perfect telemetry      & 1.00 & 0.0  & 0.00 \\
Mild degradation       & 0.95 & 1.4  & 0.50 \\
Moderate degradation   & 0.83 & 4.9  & 1.70 \\
Severe degradation     & 0.60 & 11.5 & 4.00 \\
Near-blackout          & 0.20 & 23.0 & 8.00 \\
Total blackout         & 0.00 & 28.8 & 10.00 \\
\bottomrule
\end{tabular}
\end{table}

\begin{theorem}[T4: No Free Safety]\label{thm:T4}
Under A2 and A4, for any non-trivial fault rate $d > 0$, there exists no
controller that simultaneously achieves $\TSVR = 0$ and zero intervention rate
(IR~$= 0$) across all episodes.
\end{theorem}

\begin{proof}[Proof sketch (constructive witness)]
Construct a degraded-observation episode where the deterministic LP drives SoC to
$\SOC_{\max} - \epsilon$ while the telemetry reads $\SOC_{\max} - \epsilon - \xi$
with $\xi > 0$.  Any controller that does not intervene will commit a dispatch that
pushes the true SoC above $\SOC_{\max}$.  Hence, maintaining $\TSVR = 0$ requires
IR~$> 0$.
\end{proof}

\subsection{Certificate and Temporal Theorems (T5--T8)}

\begin{theorem}[T5: Certificate Validity Horizon]\label{thm:T5}
Under A5 and A7, the safety certificate remains valid for the largest horizon
$\tau_t$ whose forward tube stays inside the safe SoC bounds:
\begin{equation}
  \tau_t = \left\lfloor \frac{\Emax - E_t}{\sigma_d} \right\rfloor,
\end{equation}
where $\sigma_d$ is the drift scale.
\end{theorem}

\begin{theorem}[T6: Certificate Expiration Lower Bound]\label{thm:T6}
Under A7, the certificate expiration time satisfies
$\tau_t \ge \lfloor(\Emax - E_t)/\sigma_d\rfloor$.  Runtime enforcement enters
fallback when $\tau_t < 1$.
\end{theorem}

\begin{theorem}[T7: Feasible Fallback Existence]\label{thm:T7}
Under A4, a constructive fallback exists: zero-dispatch ($P_t = 0$) is admissible
and preserves safety from any state in the safe-set interior.
\end{theorem}

\begin{theorem}[T8: Graceful Degradation Dominance]\label{thm:T8}
Under A2 and A5, the inflation factor $\alpha$ is non-decreasing in degradation
severity $(1 - w_t)$.  Consequently, as observation quality degrades, the safe
action set shrinks monotonically, ensuring degradation-proportional conservatism.
\end{theorem}

%% ----- Universal Extension: T9--T11 -----
\subsection{Universal Extension (T9--T11)}

The universal theorems lift the battery-domain guarantees to arbitrary domains
that satisfy the adapter contract.

\begin{theorem}[T9: Universal Impossibility --- Reliability-Aware Repair is
Necessary]\label{thm:T9}
Consider a controller that does \emph{not} adapt its safety margin to telemetry
reliability $w_t$.  Under persistent degradation with non-zero fault rate and
boundary-reachability assumptions:
\begin{equation}\label{eq:T9}
  \E[V_T] \;\ge\; c \cdot d \cdot T,
\end{equation}
where $c > 0$ is a domain-dependent constant and $d$ is the mean degradation rate.
\end{theorem}

\begin{proof}[Proof sketch ($\varphi$-mixing + Azuma--Hoeffding)]
\emph{(i)} Partition the horizon into $k$ analysis windows separated by
$\tau_{\mathrm{mix}}$ buffer gaps.
\emph{(ii)} Within each window, the reference-domain no-free-safety witness
(T4) yields per-window violation probability $p_{\mathrm{win}} \ge c\,d$.
\emph{(iii)} Define martingale differences
$D_j = Z_j^{\mathrm{win}} - \E[Z_j^{\mathrm{win}} \mid \mathcal{F}_{(j-1)(\ell+\tau_{\mathrm{mix}})}]$.
\emph{(iv)} By Azuma--Hoeffding concentration,
$\Prob\bigl[\sum_j D_j \le -t\bigr] \le \exp(-2t^2/k)$.
\emph{(v)} Combining: $\E[V_T] \ge \Omega(d\,T)$.
\end{proof}

\paragraph{Interpretation.}  T9 establishes that reliability-aware repair is
\emph{structurally necessary}, not an optional enhancement: any quality-ignorant
controller accumulates $\Omega(d\,T)$ violations.

\begin{theorem}[T10: Boundary-Indistinguishability Lower Bound]\label{thm:T10}
Fix horizon $T$.  For each step $t$, consider a binary hypothesis test between
safe ($H_{0,t}$) and unsafe ($H_{1,t}$) states with observation laws
$P_{0,t}, P_{1,t}$ satisfying $\mathrm{TV}(P_{0,t}, P_{1,t}) \le w_t$ and unsafe
prior $\Prob(H_{1,t}) \ge p_t$.  Then every controller $\pi$ satisfies
\begin{equation}\label{eq:T10}
  \E[V_T(\pi)] \;\ge\; \frac{1}{2}\sum_{t=1}^{T} p_t\,(1 - w_t).
\end{equation}
In particular, if $p_t \ge \alpha/2$ for all $t$:
\begin{equation}\label{eq:T10_special}
  \boxed{\;\E[V_T(\pi)] \;\ge\; \tfrac{1}{4}\,\alpha\,(1-\bar{w})\,T.\;}
\end{equation}
\end{theorem}

\begin{proof}[Proof (Le Cam's method)]
At each step $t$, the controller faces a binary classification problem:
distinguish $H_{0,t}$ (safe) from $H_{1,t}$ (unsafe) based on a single observation.
By Le~Cam's lemma, the maximum error rate of any test $\psi_t$ is at least
$(1 - w_t)/2$.  The probability of a violation is
$\Prob[Z_t = 1] \ge p_t \cdot (1 - w_t)/2$.
Summing over $T$ steps yields \cref{eq:T10}.
\end{proof}

\paragraph{Reliability--risk frontier.}
T10 implies that the minimum mean reliability needed to achieve an episode
violation budget $r^*$ is
\begin{equation}
  \bar{w}^* = 1 - \frac{r^*}{\alpha\,T}.
\end{equation}

\begin{theorem}[T11: Typed Structural Transfer]\label{thm:T11}
Let $\mathcal{D}$ be a domain instantiation satisfying the adapter contract
(\cref{sec:contracts}).  Assume that on every step $t$:
\begin{enumerate}[nosep,label=(\roman*)]
  \item \textbf{Coverage}: $\Prob[x_t \in \Ut \mid \mathcal{H}_t] \ge 1 - \alpha$.
  \item \textbf{Soundness}: Every $a \in \Aset(\Ut)$ keeps $f(x,a) \in \mathcal{S}$
        for all $x \in \Ut$.
  \item \textbf{Repair membership}: The repair operator returns
        $a_t^{\mathrm{safe}} \in \Aset(\Ut)$.
  \item \textbf{Fallback admissibility}: The fallback action preserves safety from
        any safe current state.
\end{enumerate}
Then:
\begin{equation}\label{eq:T11}
  \Prob\bigl[x_{t+1} \in \mathcal{S} \mid \mathcal{H}_t\bigr] \;\ge\; 1 - \alpha.
\end{equation}
\end{theorem}

\paragraph{Failure-mode converse.}
If \emph{any} of the four T11 obligations fails, there exists a domain
instantiation where the transfer argument breaks and violations accumulate.

\paragraph{Episode aggregation corollary.}
If the domain instantiation supplies per-step risk budget $r_t$ with
$\Prob[x_{t+1} \notin \mathcal{S} \mid \mathcal{H}_t] \le r_t$, then
$\E[V_T] \le \sum_{t=1}^{T} r_t$.  With the battery envelope
$r_t \le \alpha(1 - w_t)$, this recovers the core bound: $\E[V_T] \le \alpha(1 - \bar{w})T$.

\subsection{Proof Architecture Summary}

The proof programme has a layered structure:
\begin{itemize}[nosep]
  \item \textbf{Foundation} (T1, T4): OASG existence + no-free-safety establish
        the necessity of a runtime layer.
  \item \textbf{Mechanism} (T2, T3): Safety preservation + core bound show the
        sufficiency of the \dcs{} kernel.
  \item \textbf{Temporal} (T5--T8): Certificate validity, expiration, fallback, and
        graceful degradation provide operational guarantees.
  \item \textbf{Universal} (T9--T11): Impossibility, lower bound, and typed transfer
        lift battery guarantees to arbitrary domains.
\end{itemize}

All assumptions A1--A13 are testable pre-conditions verified by the 1{,}189-test
suite.  The theorem audit register (\cref{app:theorems}) confirms that all 11
theorems have complete proofs, empirical validation, and code-level anchors.
