% === Section 11: Reproducibility ===
\section{Reproducibility}\label{sec:reproducibility}

All results reported in this paper are fully reproducible from locked artifacts.
This section documents the reproducibility infrastructure.

\subsection{Artifact Bundle}

The complete artifact bundle is organised under
\texttt{artifacts/canonical\_runs/} and includes:

\begin{itemize}[nosep]
  \item Trained LightGBM models (AV) and the promoted three-domain runtime artifacts.
  \item Battery deep-lane checkpoints only as supplemental diagnostic assets; they do
        not carry the headline claim that a learned model improves the promoted lane.
  \item Processed feature files (Parquet format).
  \item Conformal calibration quantiles.
  \item Runtime trace CSVs (battery: 288 rows; AV: 18{,}697 rows covering
        baseline + ORIUS).
  \item Certificate chains (JSON).
  \item Report figures (PNG) and tables (CSV).
  \item Combined manifest (\texttt{combined\_manifest.csv}).
\end{itemize}

\subsection{Dependency Locking}

Python dependencies are locked in two files:
\begin{itemize}[nosep]
  \item \texttt{requirements.txt}: direct dependencies with version pins.
  \item \texttt{requirements.lock.txt}: fully resolved transitive closure with
        SHA-256 hashes.
\end{itemize}

The runtime environment is Python 3.11+ with NumPy, Pandas, LightGBM, and
Matplotlib as primary dependencies.  No GPU is required.

\subsection{Single-Command Pipelines}

Both domains can be reproduced with a single command:

\paragraph{Battery.}
\begin{verbatim}
make battery-pipeline
\end{verbatim}

\paragraph{AV.}
\begin{verbatim}
zsh scripts/run_full_av.sh
\end{verbatim}

The AV script executes five steps: (1) training and manifest building,
(2) DC3S runtime on the full corpus, (3) report generation, (4) combined manifest
rebuild, and (5) test suite execution.

\subsection{Test Suite}

The test suite comprises 1{,}189 tests covering:
\begin{itemize}[nosep]
  \item Contract invariant tests (monotone inflation, weight bounds, interval
        non-degeneracy).
  \item Kernel round-trip tests (observation $\to$ certificate).
  \item Adapter-specific tests (battery dispatch, AV prediction).
  \item Governance surface tests (chain validity, audit completeness).
  \item Integration tests (full pipeline end-to-end).
\end{itemize}

All 1{,}189 tests pass; 6 are skipped (environment-dependent tests not applicable
to the current host).

\subsection{SHA-256 Hashes}

Key artifact hashes are recorded in the combined manifest for tamper detection.
The manifest is regenerated on each pipeline run and includes:
\begin{itemize}[nosep]
  \item Domain name, status, report path, artifact path, and corpus identifier.
  \item File hashes for models, data files, and configuration.
\end{itemize}

\subsection{Computational Requirements}

\Cref{tab:compute} summarises the computational cost.

\begin{table}[h]
\caption{Computational requirements for full pipeline reproduction.}
\label{tab:compute}
\centering\small
\begin{tabular}{lrr}
\toprule
\textbf{Stage} & \textbf{Battery} & \textbf{AV} \\
\midrule
Training         & $<$10~s     & $\sim$45~s \\
Runtime          & $<$30~s     & $\sim$55~min \\
Report gen.      & $<$10~s     & $\sim$30~s \\
Test suite       & \multicolumn{2}{c}{$\sim$141~s (incl.\ adapter stubs)} \\
\midrule
\textbf{Total}   & $<$1~min    & $\sim$58~min \\
\bottomrule
\end{tabular}
\end{table}

All benchmarks measured on Apple M-series (single core, 16~GB RAM).

\subsection{LaTeX Compilation}

The manuscript itself is compiled via:
\begin{verbatim}
make ieee-detailed-compile
\end{verbatim}
which runs \texttt{latexmk} with \texttt{pdflatex} and \texttt{bibtex}.
