\section{Domain Instantiations under One Runtime Contract}
\label{sec:ieee-domains}

The key test of a universal safety layer is not whether every domain shares the same
dynamics or controller. It is whether the same runtime contract survives domain change
without losing semantic meaning. ORIUS evaluates that question across six rows.

\begin{table*}[t]
\centering
\caption{Six ORIUS domain rows under a shared runtime contract.}
\label{tab:ieee-domain-rows}
\small
\begin{tabularx}{\textwidth}{@{}p{0.16\textwidth}p{0.20\textwidth}p{0.17\textwidth}p{0.16\textwidth}p{0.23\textwidth}@{}}
\toprule
\textbf{Domain} & \textbf{Degraded-observation hazard} & \textbf{Safety object} & \textbf{Current tier} & \textbf{Exact closure boundary}\\
\midrule
Battery energy storage & Dispatch legal on observed state but unsafe on true state of charge or reserve surface & State-of-charge and dispatch envelope preservation & Witness row & Deepest closure surface in the current stack.\\
Autonomous vehicles & Action proposed on stale or optimistic headway / collision state & TTC and entry-barrier preservation under repair & Defended bounded row & Bounded to the current TTC plus predictive-entry-barrier contract.\\
Industrial process control & Control actuation released on degraded plant measurements & Power-temperature or operating-envelope preservation & Defended bounded row & Bounded to the current plant-family and replay surface.\\
Healthcare monitoring & Clinical alerting or intervention released on degraded vital telemetry & Vital-alert envelope and bounded intervention discipline & Defended bounded row & Bounded to the current monitoring and intervention contract.\\
Navigation and guidance & Corridor control executed on degraded localization or map consistency & Feasible corridor / obstacle envelope preservation & Gated row & Real-data defended replay chain is still missing.\\
Aerospace control & Envelope management executed on degraded flight-state surface & Flight-energy / envelope preservation under repair & Gated row & Real multi-flight runtime validation surface is still missing.\\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Autonomous vehicles}
The AV row is defended under a bounded safety object: time-to-collision and predictive
entry-barrier preservation when the perception or state-estimation surface degrades. The
value of the row is that ORIUS is not merely filtering a scalar dispatch command. It is
repairing an action inside a tightly constrained interaction envelope while preserving a
bounded fallback semantics. The row is defended, but still explicitly limited to the
current contract rather than advertised as full autonomy closure.

\subsection{Industrial process control}
The industrial row demonstrates that ORIUS also travels into operating regimes with
thermodynamic or plant-family constraints rather than geometric avoidance alone.
Degraded observation here leads to envelope violation, not just tracking error. The row
therefore helps show that ORIUS is about admissible actuation under uncertainty, not
about one family of robotic motion tasks.

\subsection{Healthcare monitoring}
Healthcare matters because degraded observation can suppress intervention rather than
only create excessive actuation. ORIUS still fits that setting because the runtime layer
is defined over admissible decisions, fallback policies, and certificate semantics, not
only over physical force commands. The defended healthcare row is bounded to the current
monitoring and alerting contract, but it proves that the runtime grammar is not limited
to mechanical control loops.

\subsection{Navigation and aerospace}
Navigation and aerospace are the two honest blockers. Navigation remains structurally
compatible with the ORIUS contract but lacks the defended real-data replay chain needed
to promote the row. Aerospace carries partial degradation and replay evidence, but not
the real multi-flight runtime surface needed for a defended bounded row. These two rows
are therefore part of the universal argument and part of the scientific boundary at the
same time.
