\section{Battery as the Witness Domain}
\label{sec:ieee-battery}

Battery remains the witness domain because it carries the deepest theorem-to-code-to-
artifact closure in the present ORIUS stack. That status is methodological, not
conceptual. The paper is not arguing that batteries are the center of Physical AI. It
is using the battery row to show what a fully specified ORIUS instantiation looks like
when the theorem chain, benchmark chain, fallback chain, and certificate chain are all
present in one place.

\subsection{Hazard and safety object}
The battery hazard is simple to state and representative of the larger ORIUS problem.
Dispatch can appear legal on the observed state of charge, reserve, or availability
surface while already violating the true physical envelope. The safety object is
therefore zero true-state violation of the dispatch and state-of-charge envelope under
degraded observation.

\subsection{Runtime instantiation}
The battery adapter uses reliability scoring, uncertainty inflation over feasible state
of charge, action clipping or safe-hold when the admissible power interval collapses,
and certificate-gated dispatch release. This makes the witness row useful for more than
numerical performance. It shows the full semantics of repair, fallback, certificate
expiry, and audit continuity under a domain where the runtime surface is already mature.

\subsection{Why the battery row matters}
The witness row earns its role because it exposes all three layers of the ORIUS claim:
theoretical motivation, benchmarked repair, and governed runtime release. That closure
depth is what the weaker rows are measured against.

\begin{table}[t]
\centering
\caption{Why battery remains the witness row in the IEEE draft.}
\label{tab:ieee-battery-witness}
\small
\begin{tabularx}{\columnwidth}{@{}p{0.30\columnwidth}X@{}}
\toprule
\textbf{Surface} & \textbf{Witness-row status}\\
\midrule
Formal bridge & Deepest theorem-to-artifact chain in the current ORIUS stack.\\
Replay evidence & Most complete degraded-observation replay, stress, and runtime traces.\\
Fallback semantics & Mature safe-hold and bounded repair behavior with certificate traces.\\
Governance & Strongest CertOS lifecycle and audit continuity evidence.\\
Role in this paper & Anchors the strongest empirical row without allowing the manuscript to become battery-centric.\\
\bottomrule
\end{tabularx}
\end{table}

This is the right role for battery in a universal-first paper. It anchors the strongest
evidence surface while leaving the conceptual center of the manuscript with the runtime
layer itself.
