\section{Related Work and Why Prior Families Are Not Enough}
\label{sec:ieee-related-work}

ORIUS does not claim that prior literatures ignored safety. The claim is more precise:
the existing families solve adjacent pieces of the problem, but none by itself closes
the runtime bridge from degraded observation to repaired actuation and auditable
release.

\subsection{Conformal and distribution-free uncertainty}
Distribution-free predictive inference gives a principled way to wrap learned
predictions with finite-sample uncertainty sets
\cite{vovk2005algorithmic,lei2018distribution,romano2019conformalized,gibbs2021adaptive,angelopoulos2023gentle,barber2023beyond,zaffran2022adaptive,kiyani2025decision}.
ORIUS inherits that finite-sample uncertainty machinery because the runtime needs a
mechanism for inflating the observation-consistent state set when trust falls. But
conformal methods alone do not decide tightened admissible actions, repaired actuation,
fallback burden, or release certificate semantics after that inflation. They are
uncertainty tools, not release-boundary runtimes.

Recent decision-theoretic work makes this boundary sharper: prediction sets are useful
to risk-averse agents because they shape downstream actions, not because they are a
complete runtime policy by themselves. ORIUS adopts that action-facing interpretation
and then adds the missing release machinery: reliability scoring, admissible-action
tightening, repair or fallback, and certificate emission.

\subsection{Runtime monitoring, anomaly detection, and supervisory veto}
Runtime monitors, anomaly detectors, and cyber-physical fault analyses provide alarms,
scores, and veto conditions when telemetry or execution behavior becomes suspicious
\cite{havelund2004runtime,bartocci2018introduction,basseville1993detection,page1954continuous,page1955test,hinkley1971inference,rajkumar2010cps,pasqualetti2013attack,teixeira2015secure,liu2008isolation,breunig2000lof}.
ORIUS inherits the idea that degraded observation must be recognized before execution.
It differs because alarms and vetoes alone do not compute tightened admissible actions,
repaired actuation, fallback burden, or certificate-bearing release governance.

\subsection{Simplex architectures and runtime assurance}
Simplex architectures, supervisory switching, runtime assurance, and shields provide a
strong language for overriding a nominal controller and releasing a trusted fallback
\cite{sha1998simplex,sha2001using,desai2019soter,bloem2015shield,leucker2009brief,kwon2025adaptive}.
ORIUS inherits the idea that the nominal controller is not the final authority. The gap
is that many simplex-style schemes start after the candidate action already exists and
assume that the observed state is coherent enough to supervise directly. ORIUS starts
one layer earlier by treating degraded observation itself as the thing that changes what
may still be released, which forces the runtime to reason jointly about reliability,
uncertainty inflation, repair geometry, fallback, and certificate semantics.

\subsection{Domain benchmarks used as evidence boundaries}
The three defended rows rely on established domain evidence surfaces, but ORIUS does
not treat those sources as interchangeable. The AV row is now bounded to nuPlan replay
and planning-benchmark semantics \cite{caesar2021nuplan,karnchanachari2024nuplan,peng2025nuplanr};
the healthcare row is bounded to retrospective MIMIC-derived monitoring evidence
\cite{johnson2023mimiciv,physionet2024mimiciv31,huang2025mimicsepsis}. These sources
support reviewer-visible boundaries: nuPlan gives a real driving-log replay surface for
the current AV row, while MIMIC-IV and MIMIC-Sepsis frame healthcare as retrospective
monitoring research rather than live clinical deployment.

\subsection{Safety filters, barrier methods, and robust control}
Barrier functions, robust MPC, predictive safety filters, and reachable-set control
make it possible to turn uncertainty into admissible action sets
\cite{ames2019control,ames2014cbf,wabersich2021predictive,fisac2019general,mayne2005robust,langson2004tubes,rawlings2017mpc,camacho2013model}.
ORIUS inherits that safe-set geometry. What it adds is not a replacement for plant-local
control design, but a typed release-boundary layer that can host those methods under one
cross-domain contract. The key scientific move is to make the runtime objects reusable
even when the underlying safety geometry differs across domains.

\begin{table*}[t]
\centering
\caption{Why adjacent method families are necessary but insufficient for the ORIUS object.}
\label{tab:ieee-family-gap}
\small
\begin{tabularx}{\textwidth}{@{}p{0.18\textwidth}p{0.24\textwidth}p{0.50\textwidth}@{}}
\toprule
\textbf{Method family} & \textbf{What it contributes} & \textbf{What remains missing for OASG}\\
\midrule
Conformal / distribution-free inference & Reliability-aware uncertainty around learned predictions & Does not specify tightened admissible actions, repaired actuation, fallback burden accounting, or certificate semantics once observation degrades.\\
Runtime monitoring / anomaly detection & Detecting degraded telemetry or veto conditions around candidate execution & Stops at alarms or vetoes unless paired with tightened action computation, repaired release, and release governance.\\
Simplex / runtime assurance & Supervisory override and safe fallback around a nominal controller & Usually assumes the observed state is already coherent enough to supervise and does not define reliability-conditioned tightening or cross-domain contract reuse.\\
Safety filters / robust control & Plant-local mapping from uncertainty into safe action geometry & Does not by itself define degraded-observation release governance, certificate lifecycle, or a reusable typed contract across domains.\\
\bottomrule
\end{tabularx}
\end{table*}

The paper's claim is therefore not that ORIUS supersedes these literatures. It claims
that Physical AI needs a reusable runtime layer that binds them together around a
single action-semantics problem: once degraded observation enters the loop, the system
must decide what may still be safely done, not only whether an anomaly has occurred.
