\section{Cross-Domain Synthesis}
\label{sec:ieee-synthesis}

The scientific value of ORIUS is not that every row already looks equally complete. It
is that one runtime grammar survives across very different physical systems while the
paper remains honest about which rows are defended and which are gated. That distinction
is what turns ORIUS from a portability story into a candidate field-level safety layer.

\begin{table*}[t]
\centering
\caption{What ORIUS standardizes across domains, and what remains domain-specific.}
\label{tab:ieee-universal-vs-domain}
\small
\begin{tabularx}{\textwidth}{@{}p{0.20\textwidth}p{0.32\textwidth}p{0.40\textwidth}@{}}
\toprule
\textbf{Layer} & \textbf{Shared across all rows} & \textbf{Domain-specific content}\\
\midrule
Observation semantics & Reliability assessment, degraded-observation detection, uncertainty widening trigger & Sensor modalities, freshness signals, and consistency features.\\
Safety geometry & Tightened admissible action set, repair-or-fallback release logic & State constraints, action constraints, safety-object geometry.\\
Benchmarking & True-state violation, intervention, fallback, certificate validity, latency, domain metrics & Domain-specific margin definitions and auxiliary metrics.\\
Governance & CertOS lifecycle, certificate integrity, audit continuity & Policy-specific fallback payloads and application-specific certificate fields.\\
\bottomrule
\end{tabularx}
\end{table*}

This synthesis yields a sharper universality statement than a transfer-learning claim.
ORIUS is not saying that one policy transfers across all six rows. It is saying that
Physical AI systems share a runtime failure mode once observation degrades, and that the
same kernel can mediate that failure mode while respecting domain-specific safety
geometry. That is why the architecture can be universal even when the evidence remains
tiered.

The paper therefore keeps the parity gate visible. Battery is the witness row. AV,
industrial, and healthcare are defended bounded rows. Navigation and aerospace are
visible blockers rather than quiet placeholders. This is not defensive writing. It is
the mechanism that lets the paper make a large architectural claim without pretending
that universality of contract automatically implies equality of closure.
