\section{Limitations and Claim Boundary}
\label{sec:ieee-limitations}

The ORIUS argument is strongest when the claim boundary remains explicit. This draft
does not claim universal optimality, equal-theorem-depth parity, universal hardware
closure, or flat deployment readiness across all supported domains. It claims one
universal runtime safety layer, one reusable action-semantics grammar for degraded
observation, and one governed evidence ladder that determines how much rhetoric each
domain row may currently carry.

That distinction matters because the architecture already travels farther than the
tracked evidence. Battery is the witness row. Autonomous vehicles, industrial control,
and healthcare monitoring are defended bounded rows under explicit contracts. Navigation
and aerospace remain the two blocker rows for any equal-domain closure claim. The
flagship manuscript therefore uses those domains to show the reach of the runtime
contract, while the parity matrix and the private claim-delta ledger preserve the
boundary between current repo truth and visionary universal framing.

The bounded-universal position is still scientifically valuable. It makes ORIUS a real
cross-domain safety layer today, while giving future closure work a precise target:
real-data navigation replay and a non-placeholder aerospace runtime surface with
material post-repair improvement on the governing safety object. Those are evidence
obligations, not conceptual holes in the runtime grammar itself.  The adversarial OQE
mode partially addresses the reliability-spoofing limitation, but full game-theoretic
completeness across all attack strategies remains an open evidence obligation.
