\section{Theory Bridge and Claim Discipline}
\label{sec:ieee-theory}

The theorem surface of ORIUS is intentionally narrower than the rhetoric of a fully
closed universal program. The paper uses theory to justify the shape of the runtime
layer, not to imply that every domain has already inherited equal proof depth.

For readability, the main paper should present the theory as four flagship
theorem roles plus supporting definitions, propositions, and corollaries rather
than as an eleven-theorem list: (i) observation-aware safety gap,
(ii) observation necessity, (iii) ORIUS covered release preservation, and
(iv) typed transfer with covered observation-ambiguity optimality.  Temporal
certificate/fallback discipline remains essential operational support, but it
should not be packaged as extra novelty inflation.  The permanent T1--T11 labels
remain audit identifiers, not a requirement that every supporting surface be
promoted as a main novelty theorem.

The bridge has three parts. First, if admissibility depends on the true state and the
observation channel can deviate materially from that state, then OASG is a genuine
safety event rather than a notation trick. Second, if lower reliability produces a
monotone inflation of the observation-consistent state set, then the admissible action
set must become smaller or remain unchanged. Third, if the runtime only releases actions
from that tightened set, or falls back when no such action exists, then the remaining
true-state violation risk is governed by the soundness of the uncertainty set and the
repair map rather than by the optimism of the nominal controller.

Temporal certificates and fallback add a fourth operational layer: T5 is a
certificate-horizon definition, while T6 and T7\_Battery are supporting
certificate/fallback propositions.  Together they state that certificates are
time-bounded safety objects and that the battery runtime either holds safely,
recovers by boundary-aware safe landing, or fails closed.  The T6 support
surface is confidence-aware and tracks a closed-form first-passage lower bound
on the remaining certificate life; the runtime enforces the resulting
$\tau_t \ge 1$ check at each step.

\begin{table}[t]
\centering
\caption{Theory-to-evidence bridge used in the IEEE draft.}
\label{tab:ieee-theory-bridge}
\small
\begin{tabularx}{\columnwidth}{@{}p{0.26\columnwidth}X@{}}
\toprule
\textbf{Claim layer} & \textbf{What the paper is willing to claim}\\
\midrule
Structural & Degraded observation can create action-semantic failure even when the nominal policy is internally coherent.\\
Contractual & A typed runtime layer can convert reliability loss into tightened admissible action sets, repair, fallback, and certificate semantics.\\
Empirical & Battery is the witness row; AV and healthcare are defended bounded rows in the active submission lane.\\
Non-claim & The paper does not claim equal theorem depth, equal deployment readiness, promoted T9/T10 status, or flat closure beyond the defended three-domain lane.\\
\bottomrule
\end{tabularx}
\end{table}

This distinction is important because ORIUS aims to be scientifically ambitious without
being careless. The architecture is intentionally written as a universal safety layer
for Physical AI. The evidence, however, is tiered. That tiering is not a weakness in
the argument; it is the mechanism that keeps theorem language, benchmark language, and
runtime language aligned with tracked artifacts.

The practical consequence is straightforward. A domain does not become a defended row
because it looks structurally compatible with the adapter contract. It becomes defended
only after the data surface, replay surface, repair semantics, fallback semantics,
certificate traces, and benchmark outputs all clear the governed promotion gate. That is
why the universal claim in this paper is architectural and bounded rather than flatly
equalized across domains already. For T9/T10, the generated promotion package is the
status authority and currently keeps both rows non-promoted until Battery, AV, and
Healthcare all discharge the relevant constants, assumptions, proof, and artifact
gates.
