\chapter{Certificate Half-Life and Communication Blackout}
\label{ch:cert-half-life-blackout}

This chapter records the battery blackout study exactly as it is now locked
in the repository.  The current result is more conservative than an earlier
drafted half-life narrative: the frozen safe action collapses to a zero-power
hold, so the battery remains safe throughout the tested blackout horizons.
That is still useful operationally, because it shows what the present
certificate logic does under severe communication loss.

\section{Why Blackout Stress Matters}

Battery controllers do not always lose safety because of a bad optimisation
decision.  They can also lose safety because communication disappears after a
safe action has already been issued.  In a blackout, the operational question
is not ``what action is optimal now?'' but rather ``how long can the last
safe action be trusted before the plant must settle into a conservative
holding mode?''

\section{Current Experimental Protocol}

The locked script \texttt{scripts/run\_blackout\_study.py} performs the
following battery-domain stress test:

\begin{enumerate}
  \item Generate a \texttt{drift\_combo} episode with the DC3S loop active.
  \item Advance the controller to a freeze point where a safe action has been
    certified.
  \item Hold the certified action fixed for blackout horizons of
    1, 4, 12, 24, and 48 hours.
  \item Evaluate the true-state battery trajectory for violations during the
    blackout.
\end{enumerate}

The resulting artifact is \texttt{reports/blackout/blackout\_study.csv}.

\section{Locked Results}

Table~\ref{tab:blackout-study} reproduces the actual locked study.

\begin{table}[ht]
\centering
\caption{Locked battery blackout study from \texttt{blackout\_study.csv}.}
\label{tab:blackout-study}
\begin{tabular}{lrrr}
\toprule
\textbf{Blackout (h)} & \textbf{TSVR (\%)} & \textbf{Sev.\ P95 (MWh)} & \textbf{Coverage} \\
\midrule
  0  & 0.0 & 0.00 &   0\% \\
  1  & 0.0 & 0.00 & 100\% \\
  4  & 0.0 & 0.00 & 100\% \\
 12  & 0.0 & 0.00 & 100\% \\
 24  & 0.0 & 0.00 & 100\% \\
 48  & 0.0 & 0.00 & 100\% \\
\bottomrule
\end{tabular}
\end{table}

The battery-operational interpretation is that the frozen certified action in
this run is effectively a safe hold: no charge, no discharge, no SOC
boundary crossing.  The certificate therefore behaves less like a decaying
aggressive dispatch and more like a safe-landing action that remains benign
throughout the tested blackout window.

\section{What the Current Result Does and Does Not Show}

This locked result establishes a bounded claim:

\begin{itemize}
  \item The present battery implementation can enter a conservative frozen
    action that remains safe for at least 48 blackout hours in the tested run.
  \item The current study does \emph{not} yet identify a finite empirical
    half-life, because no safety breakdown occurred within the measured
    horizon.
  \item The study therefore supports the safe-landing narrative of
    Chapter~\ref{ch:graceful-degradation} more strongly than a
    calibrated certificate-expiration law.
\end{itemize}

\section{Operational Consequence}

For the battery domain, this is still an important control result.  When
telemetry disappears entirely, the present DC3S stack has an auditable path
to a benign hold action rather than blindly persisting with a stale,
high-impact dispatch.  In asset terms, that means the blackout policy favors
state preservation over speculative useful work once the measurement channel
goes dark.

\section{Summary}

The blackout chapter now matches the locked evidence.  The current battery
package shows a 48-hour safe hold under frozen certification, with zero
true-state violations throughout the measured horizon.  That is a strong
battery-safe-landing result, but it should not be overstated as a fully
identified certificate half-life law until future runs produce a meaningful
decay surface.
