\chapter{Safety-Case Bundle Index}
\label{app:safety-case-bundle-index}

This appendix maps the dissertation to a safety-case style evidence bundle without making
any compliance claim. The goal is simple: a reviewer should be able to see which hazard,
which argument, which evidence packet, and which non-claim boundary belong together. The
appendix is therefore an evidence-structure aid, not a sector-specific certification claim
\cite{ul4600_2023,nist2023ai_rmf,iso26262_2018,iec61508_2010}.

\begin{longtable}{p{0.18\textwidth}p{0.26\textwidth}p{0.34\textwidth}} 
\caption{Safety-case bundle index for the ORIUS dissertation.}\label{tab:safety-case-bundle-index}\\
\toprule
\textbf{Bundle} & \textbf{Dissertation role} & \textbf{Primary evidence surfaces}\\
\midrule
\endfirsthead
\toprule
\textbf{Bundle} & \textbf{Dissertation role} & \textbf{Primary evidence surfaces}\\
\midrule
\endhead
Hazard definition & Defines degraded observation and OASG as the governing safety problem. & Chapters~\ref{ch:introduction-thesis-claims}--\ref{ch:formal-problem-formulation}, universal claim matrix, assumption register.\\
Theorem bundle & States what is proved under explicit assumptions. & Chapter~\ref{ch:mathematical-foundations}, Appendix~\ref{app:proofs}, theorem dependency table, integrated theorem gate figure.\\
Benchmark bundle & Defines replay, dual trajectories, metrics, and fault families. & Chapter~\ref{ch:benchmark-reproducibility-protocol}, Appendix~\ref{app:benchmark-specs}, domain closure matrix, submission scorecard.\\
Witness bundle & Shows the deepest theorem-to-artifact row. & Chapters~\ref{ch:witness-domain-implementation}--\ref{ch:witness-results-failure-analysis}, battery tables and figures, release manifest.\\
Cross-domain bundle & Shows structural universality and bounded transfer. & Chapter~\ref{ch:universal-orius-across-domains}, submission scorecard, runtime/governance matrices.\\
Governance bundle & Shows certificate lifecycle, runtime budgets, fallback, and audit continuity. & Chapter~\ref{ch:certos-runtime-assurance}, governance lifecycle matrix, runtime budget matrix, review-facing governance derivatives.\\
Boundary bundle & States what remains gated, roadmap, or non-promotional. & Chapter~\ref{ch:governance-reproducibility-limitations-conclusion}, deployment validation scope, gap matrix, exact non-claim prose.\\
\bottomrule
\end{longtable}

\section{How to read the bundle index}
The bundle index is intentionally orthogonal to the chapter order. A chapter can contribute
to more than one bundle, and a bundle can span multiple generated artifacts. That structure
mirrors real safety-argument practice more closely than a chapter-by-chapter checklist: one
claim family may rely simultaneously on a theorem statement, a replay trace, a runtime
lifecycle table, and an explicit non-claim boundary.

\section{Interpretive boundary}
The dissertation uses this bundle index only as an evidence-structure aid. It aligns the
manuscript with disciplined safety-argument practice and lifecycle-oriented risk governance
without implying certification or sector-specific regulatory compliance. Its role is to keep
the argument inspectable, not to relabel the dissertation as a compliance submission.
