\chapter{Related Work and the Novelty Gap}
\label{ch:related-work-novelty-gap}

ORIUS enters a crowded intellectual landscape, but the relevant literatures are crowded in a
revealing way. Each contributes a necessary piece of the final argument: uncertainty
quantification explains what can be said about predictive trust, runtime assurance explains
why intervention must remain separate from nominal control, safe-control methods explain how
constraints can reshape action, fault-detection work explains how degradation is recognized,
and governance frameworks explain how claims and evidence must be bounded. The novelty gap
appears only when those families are read together. They are rich in local technique yet
still leave the release-boundary object of degraded observation split across separate
traditions.

\section{Conformal prediction and adaptive calibration}
Conformal prediction and related distribution-free uncertainty methods provide one of the
few disciplined ways to wrap learned models with coverage-aware uncertainty without
assuming a perfectly specified error law
\cite{vovk2005algorithmic,lei2018distribution,romano2019conformalized,gibbs2021adaptive,angelopoulos2023gentle,barber2023beyond}.
Conformalized quantile regression and adaptive conformal variants extend that logic to
heteroskedastic and nonstationary settings, which makes them especially relevant once a
physical system moves through regime changes or degraded telemetry states
\cite{romano2019conformalized,gibbs2021adaptive,zaffran2022adaptive,stankeviciute2021conformal,xu2021dynamic,sesia2021comparison,bates2021distribution}.

What remains missing is the release-boundary question that follows from calibrated
uncertainty. Coverage guarantees do not by themselves decide how the admissible action set
should change, which fallback is legitimate, or what evidence must accompany release once
the observation surface becomes unreliable. ORIUS therefore treats calibration as an input
to runtime safety rather than as the full safety object.

\section{Runtime assurance and Simplex-style supervision}
Runtime assurance architectures, supervisory controllers, shielding, and runtime
verification supply the language of intervention, veto, fallback, and monitored execution
\cite{sha2001using,sha1998simplex,desai2019soter,bloem2015shield,bartocci2018introduction,havelund2004runtime,leucker2009brief,schierman2015run,bak2011runtime}.
That line of work is indispensable because it treats safe release as a separate problem from
nominal policy design.

Yet most supervisory schemes enter after the nominal controller has already proposed an
action under a state estimate that is treated as sufficiently meaningful. Crucially,
classic Simplex architectures hard-code the boundary between unverified and verified
controllers based on static state-space triggers, implicitly assuming the physical state
observation itself remains perfectly reliable. ORIUS addresses an earlier, more insidious
fault line: once observation quality degrades invisibly (the OASG), state-triggered
Simplex fails because it trusts a corrupted monitor. The DC³S assurance
layer must reason jointly about epistemic reliability, dynamically widened state
uncertainty, repaired action, fallback, and certificate semantics. A supervisor without
that observation-side, dynamically calibrated logic is still missing the central OASG
object.

\section{Safety filters, shields, and robust safe control}
Control barrier functions, predictive safety filters, robust MPC, and tube-based control
translate model uncertainty into admissible control sets and corrective intervention rules
\cite{ames2019control,ames2014cbf,wabersich2021predictive,nguyen2016exponential,wang2017safety,fisac2019general,mayne2005robust,langson2004tubes,rawlings2017mpc,ben2009robust,camacho2013model}.
These methods are the closest classical relatives of ORIUS because they formalize the
difference between nominal control intent and safe control execution.

The remaining gap is that this family usually assumes the uncertainty set and safety object
have already been defined in a domain-specific, often static or highly conservative way
(e.g., assuming a fixed worst-case bounded disturbance $w \in W$). ORIUS does not replace
that work with a universal controller. Instead, by architecturally composing Conformal
Quantile Regression (CQR) directly with Robust Optimization at runtime, DC³S dynamically
inflates and deflates the robust feasible set based on active telemetry drift. It
eliminates the crippling conservativeness of standalone robust control while providing
coverage-backed safety guarantees that adapt as sensors fail. The novelty is not the
existence of safe control or CQR alone; it is the runtime composition that dynamically
bridges data-driven epistemic uncertainty directly into physical constraint-tightening,
making safe control portable across domains.

\section{Anomaly, drift, and degraded-observation detection}
Fault diagnosis, change detection, cyber-physical security, and anomaly monitoring provide
the vocabulary for stale, delayed, reordered, dropped, or corrupted observations
\cite{rajkumar2010cps,derler2012modeling,baheti2011cyberphysical,pasqualetti2013attack,teixeira2015secure,basseville1993detection,page1954continuous,page1955test,hinkley1971inference,liu2008isolation,breunig2000lof}.
This literature is essential because degraded observation must first be recognized before it
can influence action release.

Detection alone does not specify how action semantics should change after trust degrades. A
drift alarm can say that something is wrong without saying how uncertainty inflates, how the
safe set tightens, what repair is permitted, or what certificate makes the next step
auditable. ORIUS uses detection as the opening move in a runtime safety sequence, not as
the full response.

\section{Domain-specific cyber-physical safety systems}
Energy systems, autonomous mobility stacks, healthcare monitoring, and other
cyber-physical systems already contain rich safety practice in their own local forms
\cite{rajkumar2010cps,baheti2011cyberphysical,pasqualetti2013attack,teixeira2015secure,sha2001using,ames2019control}.
This practice matters because any claim of universality that ignores domain structure is
empty from the start.

What they usually do not expose is a common contract through which degraded-observation
replay, repaired action, fallback semantics, certificate validity, and latency can be
compared across domains. ORIUS addresses that missing cross-domain runtime object rather
than trying to erase domain structure.

\section{Safety-case governance and lifecycle discipline}
The dissertation also inherits an important lesson from safety-case and lifecycle practice:
strong safety claims require structured argument, explicit assumptions, traceable evidence,
and disciplined lifecycle boundaries. ORIUS does not claim compliance with sector-specific
standards, but its claim-lock pipeline, artifact provenance, and certificate-governance
surfaces are best understood in conversation with claim-based safety arguments and
lifecycle-oriented risk management
\cite{ul4600_2023,nist2023ai_rmf,iso26262_2018,iec61508_2010}.

Governance frameworks say how evidence should be organized and how claims should be bounded;
they do not themselves provide the runtime object that converts degraded observation into
repaired or denied release. ORIUS borrows the discipline of explicit evidence structure and
lifecycle boundaries, but its technical novelty remains the runtime layer that sits between
observation degradation and actuation.

\section{Benchmark, replay, and reproducibility gaps}
There is also a measurement gap that cuts across the technical literatures above. Many papers
report prediction error, control regret, collision rate, or intervention count, but do not
record the dual-trajectory evidence needed to tell whether an action was only observed-safe or
was truly safe on the latent physical state. That omission matters because the OASG is
invisible unless the benchmark keeps both trajectories in view at once.

Without replayable fault schedules, governed intervention traces, and a locked mapping from
headline claims to artifacts, a physical-AI safety result is difficult to audit and easy to
overstate. ORIUS therefore treats measurement discipline as part of the contribution rather
than as packaging around the contribution. The benchmark contract, claim matrix, and manifest
surfaces are necessary because they keep the monograph grounded in true-state safety rather
than nominal software coherence.

\section{Why ORIUS is not another forecasting or perception paper}
The literature also invites a narrower but common misunderstanding: that ORIUS should be read as
one more attempt to beat a baseline predictor, one more perception stack, or one more
domain-specific robust controller. The dissertation rejects that framing directly. In the
strongest witness row, forecasting quality matters because it affects calibrated uncertainty and
therefore the release boundary, but the main scientific object is still the runtime layer that
decides what action remains admissible once the observation surface becomes unreliable.

\paragraph{Why this matters for the novelty claim.}
This distinction prevents ORIUS from collapsing into a model-comparison paper. A new forecaster
or perception backbone may improve nominal error while leaving the release-boundary question
unchanged. ORIUS is novel only if it remains centered on degraded observation, admissible
repair, governed fallback, and certificate semantics. That is why the related-work chapter is
organized by missing runtime objects rather than by model families alone.

\section{The novelty gap ORIUS addresses}
The literature becomes most powerful when these families are read together, but it also
reveals a persistent separation of levels. Uncertainty methods stop at calibrated sets.
Assurance methods stop at supervisory intervention. Safe-control methods stop at
plant-specific repair geometry. Detection methods stop at identifying degraded telemetry.
Domain systems stop at application-local safety practice. What is missing between them is a
single typed runtime composition that starts with observation reliability and ends with
repaired action, fallback, certificate semantics, and cross-domain replay under one common
contract.

The exact flagship novelty sentence defended by the dissertation is:
\begin{quote}
\emph{ORIUS identifies OASG as the degraded-observation release hazard and provides a reliability-aware runtime safety layer across Battery, AV, and Healthcare.}
\end{quote}
That sentence is intentionally narrower than a claim of a new conformal method, a new
robust-optimization primitive, a new universal controller, or a new
conditional-coverage theorem. The promoted ML surface is grouped calibration by
reliability bucket plus bounded runtime-safety deltas on the active three-domain lane.

That missing composition is ORIUS's novelty claim. The dissertation does not argue that any
one ORIUS component is unprecedented in isolation. It argues that the combination of
observation reliability, action repair, certificate semantics, and universal replay has not
previously been stabilized as one reusable runtime safety layer for physical AI. ORIUS is
therefore not ``better forecasting,'' ``better perception,'' or ``one more controller.'' It
is a release-boundary discipline for degraded-observation safety.

The closest plausible rival to ORIUS is not any single paper in these families. It is an
imagined composition of them: adaptive conformal intervals for uncertainty, a safety filter
or Simplex supervisor for intervention, anomaly detection for degraded telemetry, and
runtime verification for logging and policy enforcement. Even that composite stack still
stops short of ORIUS's central object. It lacks a typed release contract that binds
observed-state and true-state semantics, repaired action, fallback, certificate lifecycle,
and cross-domain replay under one explicit interface. That final integration step is the
novelty claim the dissertation defends.

\paragraph{Boundary of this chapter.}
This chapter establishes a literature gap, not a theorem or a parity result. Its role is to
show why uncertainty, supervision, safe control, detection, and governance still leave the
release-boundary object fragmented. The next chapters must still prove and validate that
ORIUS closes enough of that fragmentation to justify a universal runtime safety-layer claim
under bounded evidence.
