\chapter{Battery as the Witness Domain}
\label{ch:battery-bridge}
\label{ch:battery-to-orius}

Battery appears first among the six domain chapters because it carries the deepest
theorem-to-code-to-artifact lineage in the present book, not because ORIUS is an
energy-specific framework.  This chapter uses the same universal template as the other
domain chapters and then points into the longer battery evidence block that follows.

\section{Domain problem}
Grid-scale storage dispatch must remain physically feasible even when the telemetry channel
that reports state of charge, power, or availability becomes stale, delayed, missing, or
otherwise degraded.

\section{Degraded-observation hazard}
Battery makes the universal ORIUS hazard concrete: a dispatch action can look legal on the
observed state while already violating the true state of charge or reserve envelope of the
underlying asset.

\section{ORIUS instantiation}
The battery adapter binds reliability scoring, uncertainty inflation, action repair,
certificate issuance, and benchmark replay to an explicit dispatch feasibility surface.

\section{Safety object}
The governing safety object is zero true-state violation of the dispatch and state-of-charge
envelope under degraded observation.

\section{Dataset and training surface}
The battery row is backed by the deepest locked data, replay, theorem, and artifact stack
in the repo, including the detailed calibration, stress, latency, and certificate-horizon
chapters that follow immediately after this bridge chapter.

\section{Replay and evidence surface}
The detailed battery chapters provide the reference OASG witness, the zero-TSVR repair
surface, the stress-test analysis, conditional-coverage audits, graceful-degradation study,
bounded composition case, and adversarial probing evidence that make the battery row the
deepest current reference surface in the monograph.

\section{Fallback and runtime behavior}
Battery uses certificate-gated safe-hold, dispatch clipping, and bounded expiry semantics so
the runtime can preserve the physical envelope even when the nominal optimizer remains
aggressive.

\section{Limitations and exact non-claims}
Battery is the deepest theorem-grade row, but it is not the conceptual center of the book
and it is not allowed to silently stand in for equal-domain closure.  Its role is to anchor
the proof depth while the rest of the monograph tests how much of that contract survives
domain transfer.
