\chapter{Compositional Safety}
\label{ch:compositional-safety}

\section{Why local safety does not automatically compose}
One of the easiest mistakes in safety argumentation is to assume that individually safe
controllers remain jointly safe when they share resources or constraints. ORIUS treats
composition as a bounded extension precisely because this implication fails in general.
Local certificates can remain valid while shared budgets, timing windows, or coupled plant
dynamics still create system-level failure.

\section{A simple counterexample}
Consider two locally admissible controllers that each satisfy their own state and action
constraints. If both consume a shared feeder budget, corridor occupancy, or actuator
authority reserve, their simultaneous release can still violate the system-level constraint
even though neither local controller is individually unsafe. This is the compositional
version of OASG: local legality can diverge from system-level safety when the shared
constraint is not represented in the local release object.

\section{Shared constraints as the right composition surface}
Composition is therefore framed around shared constraints rather than around an immediate
claim of universal multi-agent closure. Shared power budgets, corridor occupancy, common
actuator authority, or coupled timing envelopes are the right first-class objects because
they say where local certificates stop being sufficient. ORIUS can then lift the
certificate object into a negotiated setting without pretending that single-agent repair
theorems automatically solve the multi-agent case.

\IfFileExists{paper/assets/tables/generated/tbl_multi_agent.tex}{%
\input{paper/assets/tables/generated/tbl_multi_agent}%
}{%
\input{assets/tables/generated/tbl_multi_agent}%
}

\section{Multi-agent certificates and negotiation}
The composition story in ORIUS is therefore narrow but meaningful. A local certificate says
that an action is safe given one agent's uncertainty set and fallback semantics. A
compositional certificate must additionally encode the shared constraint on which multiple
agents depend. This turns negotiation and fairness from policy afterthoughts into runtime
objects. If the shared constraint cannot be satisfied jointly, the runtime must decide which
agents may continue, which must degrade, and which fallbacks remain legitimate.

\section{Why composition remains bounded in the dissertation}
Composition matters scientifically because it shows that the ORIUS grammar does not end at
single-agent control. At the same time, it remains outside the strongest universal claim.
The present dissertation does not defend an equal-domain composition theorem across all six
rows. It shows that shared constraints can be represented through the same runtime object
and that local certificates alone are not sufficient once coupling enters the system.

\section{Compositional takeaway}
What this chapter contributes is a bounded extension, not a new center of gravity for the
thesis. ORIUS can lift degraded-observation safety into a shared-constraint setting by
making system-level resources explicit in the release object. It does not claim that local
safety composes automatically, and it does not claim that every current domain row already
clears a defended multi-agent parity gate.
