\chapter{Conclusion}
\label{ch:monograph-conclusion}

ORIUS is presented in this monograph as a universal safety layer for physical AI under
degraded observation.  The universal contribution is architectural, semantic, and
governance-centered: one typed runtime contract, one benchmark discipline, one
certificate lifecycle, and six first-class domain chapters that all instantiate the
same degraded-observation question.

The manuscript makes four durable claims.  First, degraded observation is not a
nuisance variable; it is a universal action-semantics hazard for physical AI.  Second,
a reusable Detect--Calibrate--Constrain--Shield--Certify kernel can expose that hazard
at runtime, translate it into admissible repair and fallback semantics, and leave an
auditable certificate trail behind every protected action.  Third, the same benchmark
schema and governance contract can be reused across six domains without collapsing
them into one plant-specific story.  Fourth, universality becomes more credible when
it is defended by a visible parity gate than when it is asserted by analogy.

That last point is what makes the current ORIUS package scientifically robust. Battery remains the deepest witness row; autonomous vehicles, industrial control, and healthcare monitoring are defended bounded rows; navigation and aerospace remain explicitly gated rows. The resulting claim is not smaller than a flat universal
slogan.  It is stronger.  It says that ORIUS already closes a real cross-domain
runtime-safety problem while making the boundary of unfinished evidence impossible to
hide.

The field implication is equally important.  Physical AI safety is often written as a
choice between better models, better nominal control, or better post hoc oversight.
ORIUS argues for a fourth object: a first-class runtime safety layer that sits between
nominal intelligence and actuation, reasons directly about degraded observation, and
forces repair, fallback, and certification into the same protected loop.  If that
object becomes standard, physical AI can be reviewed and deployed with clearer failure
semantics than most current system stacks expose.

The evidence boundary remains visible. Navigation is promoted only when a locked real-data train/validate/replay row exists and clears replay, soundness, fallback, and runtime gates. Aerospace is promoted only when provider-approved multi-flight telemetry closes the governed parity lane with material post-repair gain on the defended flight safety object. Architecture universality is preserved; equal-domain universality is claimed only after every domain clears the same evidence, calibration, runtime, and governance gates. It claims one universal
architecture, one reusable runtime grammar, one governed evidence ladder, and a
credible program for promoting additional domains without weakening the scientific
contract.
