
\chapter{From Battery Adapter to Universal ORIUS}
\label{ch:battery-to-orius}

The preceding chapters proved a concrete result: the DC3S shield
eliminates true-state SOC violations for a grid-scale battery under
realistic telemetry faults.  This chapter asks whether that result is an
accident of battery physics or a structural consequence of the ORIUS
architecture.  We argue the latter.  Every pipeline component operates on
abstractions with natural counterparts in other cyber-physical domains.
The battery adapter is one instantiation; the ORIUS principle is the
invariant.

\section{Why the Battery Result Is Structurally General}

The DC3S pipeline never inspects the internal dynamics of the battery
plant.  It consumes five objects:

\begin{enumerate}
  \item An observation vector $z_t$ (telemetry readings at time $t$).
  \item A scalar reliability score $w_t \in [w_{\min}, 1]$ summarising
    telemetry quality.
  \item A conformal prediction region $C_t^{\text{RAC}}$ over the true state,
    inflated by $1/w_t$.
  \item A feasible action set $\mathcal{A}_t$ defined by domain constraints.
  \item A safe-action projection $a_t^{\text{safe}}$ that maps any candidate
    action into $\mathcal{A}_t$ under the worst-case realisation of
    $C_t^{\text{RAC}}$.
\end{enumerate}

None of these objects mention SOC, voltage, or Coulomb counting.  They are
\emph{domain-general}: any CPS providing equivalent definitions can reuse
the ORIUS pipeline without modification.

\section{Mapping Battery Objects to the Universal Adapter}

Table~\ref{tab:domain-mapping} makes the mapping explicit.

\begin{table}[ht]
\centering
\caption{Battery-specific objects and their domain-general counterparts.}
\label{tab:domain-mapping}
\begin{tabular}{lll}
\toprule
\textbf{ORIUS symbol} & \textbf{Battery instantiation} & \textbf{General meaning} \\
\midrule
$z_t$                  & SCADA packet (SOC, V, I, load) & Raw telemetry vector \\
$w_t$                  & OQE composite score            & Observation reliability \\
$C_t^{\text{RAC}}$     & SOC conformal interval          & State uncertainty region \\
$\mathcal{A}_t$        & Charge/discharge power bounds   & Constraint-feasible actions \\
$a_t^{\text{safe}}$    & SOC-preserving dispatch          & Safety-projected action \\
\bottomrule
\end{tabular}
\end{table}

\section{What Changes by Domain}

Porting ORIUS to a new domain requires replacing four domain-specific
components:

\begin{description}
  \item[Plant dynamics.] The battery uses a linear SOC integrator;
    a robotic arm requires nonlinear rigid-body dynamics.  The pipeline
    is agnostic---it consumes the predicted state set, not the equations.
  \item[State definition.] SOC is a scalar in $[0,1]$.  An autonomous
    vehicle has a six-degree-of-freedom pose.  The conformal region
    $C_t^{\text{RAC}}$ generalises to a set in $\mathbb{R}^n$.
  \item[Constraint structure.] Battery constraints are box constraints
    on SOC and power.  Other domains may have polytopic, conic, or
    nonlinear constraints.  The safe-action filter requires only that
    projection onto the feasible set is computationally tractable.
  \item[Fault taxonomy.] Dropout, delay, drift, and stale-sensor faults
    are battery-specific.  Each domain has its own fault families---GPS
    multipath for vehicles, encoder drift for robots---but the OQE
    scoring architecture applies unchanged.
\end{description}

\section{What Remains Invariant Across Domains}

Four structural invariants transfer without modification:

\begin{enumerate}
  \item \textbf{OQE architecture.}  The feature vector changes; the
    weighted scoring function and calibration procedure do not.
  \item \textbf{RAC-Cert inflation.}  The conformal prediction interval
    is domain-agnostic.  Reliability-adaptive width scaling
    $\hat{q}_\alpha / w_t$ applies to any residual distribution.
  \item \textbf{Repair projection.}  The safe-action filter projects
    onto the feasible set under worst-case uncertainty.  The projection
    operator changes with the constraint geometry; the principle of
    worst-case projection does not.
  \item \textbf{Certification logic.}  The certificate records
    $(w_t, C_t^{\text{RAC}}, a_t^{\text{safe}}, \text{margin})$
    regardless of what the symbols denote physically.
\end{enumerate}

\section{How to Port ORIUS to a New Domain}

A practitioner deploys ORIUS on a new plant by implementing a
\texttt{DomainAdapter}---the analogue of \texttt{BatteryDomainAdapter}.
The adapter provides:

\begin{enumerate}
  \item A telemetry parser that produces $z_t$.
  \item An OQE feature extractor for the domain's fault families.
  \item A plant model (or learned surrogate) for conformal residual
    calibration.
  \item A constraint specification and projection oracle.
\end{enumerate}

The ORIUS core---OQE scoring, RAC inflation, certificate issuance---remains
a shared library, tested once and reused rather than reimplemented per domain.

\section{Chapter Summary}

The battery result is structurally general because the DC3S pipeline
operates on domain-abstract objects: observations, reliability scores,
conformal regions, feasible sets, and projected actions.  What changes
across domains is the plant-side adapter; what remains invariant is the
safety architecture.  Porting ORIUS to a new domain reduces to
implementing a \texttt{DomainAdapter}---a bounded engineering task, not a
research contribution---provided the new domain satisfies the regularity
conditions of Chapter~\ref{ch:problem-formulation}.
