
\chapter{What This Thesis Proves and What It Leaves Open}
\label{ch:what-proves}

A dissertation that claims a universal principle and validates it across multiple
domains must be explicit about the boundary between proven and projected.
This chapter draws that boundary for the ORIUS framework.

\section{The Full Formal Apparatus}

The formal surface of this dissertation is larger than the T1--T8 label
suggests.  The integrated audit (Table~\ref{tab:integrated-theorem-surface-summary})
records \textbf{67 formal items}: 18~theorems, 5~lemmas, 7~propositions,
10~corollaries, 11~definitions, and 16~standing assumptions.  All 18
theorems pass the integrated theorem gate with locked code anchors and
artifact traceability.

The foundational theorems precede the formal DC3S ladder:

\begin{description}
  \item[\textbf{S1} (Existence of the Illusion under Dropout).]
    Established in Chapter~\ref{ch:safety-illusion},
    S1 proves that under any positive dropout rate a deterministic
    controller can satisfy its observed constraint while violating the
    true physical envelope.  This is the phenomenon that motivates
    the entire framework.
  \item[\textbf{S2} (Informal Safety Guarantee of DC3S).]
    Also in Chapter~\ref{ch:safety-illusion}, S2 gives
    the informal correctness claim: DC3S closes the gap S1 identifies
    by inflating the feasible action set by the reliability-adjusted
    conformal quantile.
\end{description}

The eight principal theorems T1--T8 then formalise, quantify, and
extend this guarantee across one-step, episode-level, and multi-step
temporal horizons.

\section{What Is Established Across the Tiered Runtime Package}

The following results are established by reproducible experiment on the locked
CPSBench harness (seed~42, $N = 2000$ fault-stressed steps, unified fault
schedule: dropout~15\%, spike~8\%, stale~10\%) together with the battery
reference proofs and the locked tier tables:

\begin{enumerate}
  \item \textbf{OASG existence.}  The Observation-Action Safety Gap is
    measurably non-zero across the locked runtime rows under the shared fault
    schedule.
  \item \textbf{DC3S effectiveness.}  The DC3S shield achieves
    0.0\% TSVR across the three proof-validated domains (energy management,
    industrial process control, and medical monitoring) under the locked
    three-seed evaluation protocol.
  \item \textbf{Zero true-state violation rate.}  Energy management, industrial
    process control, and medical monitoring each achieve 0.0\% TSVR under DC3S
    in the strongest evidence rows.  The autonomous vehicle and aerospace domains
    are classified as proof-candidate and experimental surfaces respectively; their
    TSVR remains at baseline under the locked protocol (2.78\% and 9.72\%).
  \item \textbf{Formal safety bound.}  Theorem~\ref{thm:safety-preservation}
    guarantees that the true-state violation probability is at most $\alpha$
    with confidence $1 - \delta$, under exchangeability of the calibration
    residuals and bounded observation error.  The bound is formalized on the
    battery reference row and instantiated across the tiered domain maps via
    the shared domain-adapter construction.
  \item \textbf{Bounded intervention cost.}  The shield's repair rate remains
    below 100\% across the five forecasting-capable runtime rows (mean~$\bar{w}_t \in
    [0.66, 0.81]$), confirming that DC3S does not collapse all actions to a
    constant fallback.
  \item \textbf{Sub-millisecond overhead.}  Pipeline latency (P95~$<$~0.1~ms)
    is consistent across the five forecasting-capable runtime rows, confirming runtime
    suitability for fast control-loop applications.
\end{enumerate}

Navigation remains simulation-backed portability evidence outside this
forecasting-surface summary.

\section{How Papers~1--6 Divide the Thesis Burden}

The thesis now states its program structure explicitly rather than leaving it
implicit in imported long-form blocks:

\begin{enumerate}
  \item \textbf{Paper~1} is the complete proof witness.  It carries the
    battery OASG result and the T1--T8 ladder.
  \item \textbf{Paper~2} is a bounded temporal theorem layer.  It adds
    certificate horizon, expiration, and blackout safe-hold semantics.
  \item \textbf{Paper~3} is a bounded empirical fallback layer.  It adds the
    canonical graceful-fallback benchmark under prolonged blindness.
  \item \textbf{Paper~4} is a benchmark-contract layer.  It formalizes
    ORIUS-Bench as the replayable evaluation discipline of the thesis.
  \item \textbf{Paper~5} is a bounded composition counterexample layer.  It
    shows where local certificates fail to compose and where coordination
    restores fleet-level safety.
  \item \textbf{Paper~6} is a bounded runtime-governance layer.  It adds
    lifecycle semantics, audit continuity, fallback, and recovery.
\end{enumerate}

This program-first structure is maximal in thesis prominence but still
conservative in proof burden: Papers~2--6 are real defended contributions, yet
none is promoted to the same theorem class as Paper~1 unless the linked code,
tests, and artifacts support that stronger claim.

\section{What Is Fully Validated in Battery (Reference Domain)}

The battery domain provides the complete theorem-to-code-to-artifact lineage:

\begin{enumerate}
  \item \textbf{OASG existence.}  Measurably non-zero at dropout rates
    as low as~5\%.  Chapter~\ref{ch:main-results} reports TSVR of
    3.9--4.2\% for the deterministic LP baseline at 20\% dropout.
  \item \textbf{DC3S effectiveness.}  Exactly 0\% TSVR across all tested
    dropout rates $d \in [0, 0.30]$, both grid regions, and all fault combinations.
  \item \textbf{Full formal proof surface.}  Theorems T1--T8 are stated,
    proved, and connected to code artifacts for the battery domain.
    Chapter~16--20 each carry domain-instantiation subsections extending
    each theorem to the four non-battery domains.
  \item \textbf{Intervention cost bounded.}  Repair rate remains below
    3.1\% under \texttt{drift\_combo} and below 20\% even at 30\% dropout.
\end{enumerate}

\section{What Is Partially Proved}

Several claims are supported by evidence but carry caveats:

\begin{enumerate}
  \item \textbf{Conditional coverage.}  CQR group coverage
    (Chapter~\ref{ch:conditional-coverage}) is validated on the four
    tested fault families.  Coverage under \emph{out-of-distribution}
    faults not in the calibration set has not been verified.
  \item \textbf{Aging effects.}  Aging-aware recalibration
    (Chapter~\ref{ch:temporal-extensions}) is demonstrated on synthetic
    degradation curves.  Multi-year field aging data is not available.
  \item \textbf{Certificate half-life.}  The blackout study
    (Chapter~\ref{ch:cert-half-life-blackout}) establishes a bounded
    safe-hold result; longer horizons (weeks, months) remain extrapolated.
  \item \textbf{Transfer across regions.}  The US--DE transfer stress
    test (Chapter~\ref{ch:transfer-generalisation}) shows positive
    results for two grid regions.  Grids with fundamentally different
    market structures are untested.
\end{enumerate}

\section{Why Multi-Domain Validation Strengthens the Core Claim}

The universal validation is not scope-widening---it is confirmation that
the core claim (DC3S closes the OASG) is not an artefact of battery physics.
Each additional domain that passes the same promotion gate (locked telemetry,
verified training surface, SIL trace pass, material TSVR reduction) provides
independent evidence that the OASG is a structural phenomenon in any CPS with
degraded telemetry, not a battery-specific measurement anomaly.

Battery provides the canonical instantiation with full theorem lineage.
Industrial and healthcare provide proof-validated empirical depth under the
shared universal replay harness.  AV contributes proof-candidate evidence,
aerospace contributes experimental evidence, and navigation provides
simulation-backed portability evidence.  Together, the runtime rows provide
broad evidence of the OASG phenomenon under the ORIUS principle, while the
final claim boundary still follows the locked tier matrix rather than a flat
reading of every row.

\section{What Remains Outside the Current Evidence}

\begin{enumerate}
  \item \textbf{Field deployment.}  No domain has been validated on production
    hardware in a live safety-critical loop.
  \item \textbf{Heterogeneous multi-asset composition.}  Fleet-level shielding
    is validated for a two-battery shared-transformer case; cross-domain
    composition (battery + AV + medical device) is an open problem.
  \item \textbf{Adversarial completeness.}  The active-probing audit detects
    the tested spoofing strategy.  Completeness against all adversarial
    strategies is not claimed.
  \item \textbf{Non-box constraint geometry.}  All current domain adapters
    use box or interval constraints.  Polytopic and conic constraint
    structures are architectural claims, not yet empirically validated.
\end{enumerate}

\section{Chapter Summary}

This thesis contributes a 67-item formal apparatus (18 theorems, 5 lemmas,
7 propositions, 10 corollaries, 11 definitions, 16 assumptions) grounded
in 25 chapter-level traceability rows connecting each formal result to its
code anchor and artifact evidence.  The ORIUS principle is established on a
tiered runtime package: the OASG exists across the locked runtime rows,
DC3S achieves zero TSVR in the battery reference row and the two
proof-validated peers, AV remains proof-candidate, aerospace remains
experimental, and navigation adds simulation-backed portability evidence.
Being explicit about what remains outside the battery reference row and
proof-validated peer rows protects the core multi-domain result from
credibility it does not yet claim.
