
\chapter{Conclusion}
\label{ch:conclusion}

\section{The Central Message}

Physical AI controllers routinely make decisions based on observations they
have no reason to trust.  A telemetry packet arrives late, or not at all,
or carrying a stale reading---and the controller evaluates its safety
constraints against a state the plant may never have occupied.  The
resulting Observation-Action Safety Gap is invisible to any evaluation
protocol that checks constraints only on the observed trajectory.

This dissertation has shown, in the battery dispatch domain, that the gap
is real, measurable, and dangerous.  More importantly, it has shown that
the gap is \emph{closable}.  The ORIUS principle---make observation
reliability part of the control state, inflate uncertainty in proportion to
unreliability, and project actions onto the tightened feasible set---eliminates
true-state violations without replacing the upstream optimiser or requiring
a new plant model.

Under degraded observation, safety cannot be taken for granted.  ORIUS makes
observation reliability a first-class quantity in the control loop, not an
afterthought logged for post-hoc analysis.

\section{What the Battery Domain Taught Us}

Grid-scale battery dispatch was chosen as the first proving ground because
it combines hard state constraints, high action-to-state sensitivity, and
network-mediated telemetry.  The experimental campaign yielded several
lessons that extend beyond batteries:

\begin{enumerate}
  \item \textbf{Hidden violations are common.}  Under 20\% dropout, a
    deterministic LP violates true-state SOC bounds at 3.9--4.2\% of
    steps while reporting full feasibility.  The controller believes it
    is safe; the plant disagrees.
  \item \textbf{Conformal prediction is the right uncertainty language.}
    Distribution-free coverage guarantees allow the safety layer to operate
    without a parametric model of telemetry faults.  The CQR intervals
    adapt to the empirical residual distribution, including heavy tails
    induced by sensor drift.
  \item \textbf{Reliability-adaptive inflation is the key mechanism.}
    A conformal interval of fixed width is either too wide when telemetry
    is good (excessive conservatism) or too narrow when telemetry is
    degraded (missed violations).  Scaling width by $1/w_t$ tracks the
    actual risk, concentrating intervention in the episodes that need it.
  \item \textbf{The cost of safety is modest.}  Intervention rates of
    2.8--3.1\% under the canonical fault scenario translate to fewer than
    three corrections per hundred dispatch intervals---a negligible
    operational burden for the elimination of all hidden violations.
\end{enumerate}

\section{Why Observation Reliability Belongs Inside Control}

Classical control treats observation noise as an exogenous disturbance to
be filtered, not a quantity the controller should reason about.  This
separation is justified when the noise model is accurate and stationary.
In modern CPS, neither condition holds: telemetry faults are non-stationary,
packet-level, and correlated with system load.  Treating reliability as
external means the controller cannot tighten its actions when observation
quality degrades, and it cannot relax them when quality is high.

ORIUS brings reliability inside the loop.  The OQE score $w_t$ is updated
at every step from raw telemetry features.  It feeds directly into the
uncertainty inflator, which sets the width of the safe-action region.  The
controller's conservatism is therefore \emph{proportional to its ignorance}:
when it knows little, it acts cautiously; when it knows much, it acts
freely.  This proportionality is the architectural insight of the
dissertation, and it is domain-independent.

\section{What ORIUS Can Become}

ORIUS is no longer only a blueprint for future generalisation.  In the
program-first dissertation, the battery result remains the reference proof
witness, while Papers~2--6 already make temporal validity, graceful
fallback, benchmark discipline, bounded composition, and bounded runtime
governance explicit parts of the thesis contribution.

What remains open is narrower than older roadmap language suggested.  The
frontier is not ``whether ORIUS extends beyond battery'' in the abstract; it
is whether specific open rows and deployment surfaces can be promoted beyond
their current bounded status.  Autonomous vehicles still face a
constraint-class mismatch, navigation still lacks the defended real-data row,
and aerospace still lacks a stronger promotion contract.  The runtime itself
still requires field deployment work beyond the current bounded governance
semantics.  Those are real open problems, but they now sit on top of a
substantive program rather than ahead of it.

\section{Final Remarks}

This dissertation began with a question: what happens when a controller
cannot trust its own eyes?  The answer, demonstrated rigorously in one
domain and argued structurally for others, is that safety must be
\emph{earned} at every time step by measuring observation quality,
quantifying the resulting uncertainty, and restricting actions to those
that remain safe under the worst case.

The battery domain was the right place to start, and the present dissertation
already shows how ORIUS extends beyond that start through Papers~2--6 without
pretending those extensions carry equal proof closure.  That is the right
thesis boundary: one full witness, one explicit research program, and one
honest frontier.
