\chapter{Compositional Safety for Battery Fleets}
\label{ch:compositional-safety-fleets}

This chapter keeps the thesis inside the battery domain while pushing beyond
the single-asset case.  The current locked extension is a two-battery shared
transformer study, executed through the canonical multi-agent scenario and
packaged through
\path{reports/fleet/protocol_compare.csv},
\path{reports/fleet/fleet_composition_two_battery.csv} and
\path{reports/fleet/two_battery_composition_metrics.csv}.

\section{Why Composition Matters}

Two individually safe batteries can still create a fleet-level problem if
they respond to local conditions without coordinating at the point of
interconnection.  In battery terms, the risk is not only a local SOC breach;
it is also aggregate export or curtailment pressure on the shared transformer.

\section{Battery-Fleet View}

The fleet extension keeps the ORIUS battery logic local to each unit:

\begin{enumerate}
  \item Each battery produces its own candidate safe action under degraded
    telemetry.
  \item The shared transformer then enforces a fleet-level envelope on the
    combined dispatch.
  \item Any excess is absorbed as coordinated curtailment rather than as an
    unsafe aggregate command.
\end{enumerate}

This is the battery-operational meaning of composition: preserve local safety
while making sure the shared export path never becomes the new hidden gap.

\section{Locked Two-Battery Composition Surface}

The fleet wrapper now uses one executable source of truth:
\texttt{src/orius/multi\_agent/scenarios.py}.  The chapter-facing script
\texttt{scripts/run\_two\_battery\_fleet.py} runs that scenario, keeps the
step log under the historical fleet filename for compatibility, and derives
the summary metrics directly from the protocol comparison.  It no longer mixes
the chapter claims with the separate curtailment-only wrapper path.

The resulting bounded summary reports:

\begin{itemize}
  \item 24 independent joint-violation steps,
  \item 0 centralized joint-violation steps,
  \item 0 distributed joint-violation steps, and
  \item 1920\,MWh of useful work under coordination.
\end{itemize}

These values come directly from
\texttt{reports/fleet/two\_battery\_composition\_metrics.csv}.

\begin{table}[ht]
\centering
\caption{Locked fleet-composition metrics from the current two-battery study.}
\label{tab:fleet-results}
\begin{tabular}{lrr}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Battery consequence} \\
\midrule
Independent joint-violation steps & 24 & Local certificates do not compose at the shared feeder \\
Centralized joint-violation steps & 0 & Coordinated ORIUS repair closes the two-battery gap here \\
Distributed joint-violation steps & 0 & Negotiated coordination matches centralized closure here \\
Coordinated useful work (MWh) & 1920 & Safety repair does not force total shutdown in this bounded case \\
\bottomrule
\end{tabular}
\end{table}

\section{Multi-Agent Protocol Comparison}

Table~\ref{tab:multi_agent} reports the non-composition counterexample
quantitatively: two batteries each proposing 60~MW of discharge into a
shared 80~MW feeder.  The independent protocol produces 24 joint
violations; ORIUS-coordinated (centralized) produces zero, while
preserving the same useful work output.

\IfFileExists{paper/assets/tables/generated/tbl_multi_agent.tex}{%
\input{paper/assets/tables/generated/tbl_multi_agent}%
}{%
\input{assets/tables/generated/tbl_multi_agent}%
}

\section{Interpretation}

The important battery result is not that the fleet chapter is ``finished'' as
a full theorem-backed composition theory.  It is that the repo now contains a
locked, chapter-facing composition run showing exactly where local
certificate composition fails and how coordination removes those feeder
violations without reducing useful work in the present two-battery stress
case.  That closes a practical gap between single-battery safety and fleet
operation without pretending that all composition questions are solved.

\section{What Remains Open}

The present fleet package is intentionally narrow:

\begin{itemize}
  \item It is a two-battery study, not a heterogeneous multi-asset proof.
  \item It reports one shared-feeder counterexample and its coordinated repair,
    not a full fairness or degradation-allocation frontier.
  \item It motivates the quality-score gossip and coordinated fallback ideas,
    but does not yet prove their optimality.
\end{itemize}

\section{Summary}

The battery thesis now includes a real fleet-composition artifact rather than
only a conceptual extension.  The locked two-battery study shows a concrete
non-composition counterexample, zero-violation coordinated repair, and
preserved useful work on the same executable scenario surface.  For the
battery book, that is the correct claim: composition has moved from
future-work rhetoric into an auditable battery extension, while still
remaining narrower than a full universal coordination theory.
