\chapter{What the Monograph Proves and What It Does Not}
\label{ch:what-battery-thesis-proves}

This chapter draws a strict line between the central theorem contribution of
the dissertation and the surrounding evidence package.  The main value of the
work is not that it claims equal proof closure across six domains; it is that
it isolates one degraded-observation safety kernel, proves it completely on a
battery reference surface, extends it with necessity and lower-bound results,
and then instantiates it across tiered cross-domain evidence rows with an
explicit claim boundary.

\section{Central Theorem Contribution}

The monograph establishes the following theorem program:

\begin{enumerate}
  \item \textbf{The observation--reality gap is the control object.}  The
    observation-action safety gap is not a battery-only accident.  It arises
    whenever a controller proposes actions from degraded telemetry while
    safety is evaluated on the true plant state.  T1 identifies this gap on
    the reference surface, and T9 shows that a quality-ignorant controller
    cannot remove the resulting violation scaling by tuning alone.

  \item \textbf{Battery is the full proof witness.}  The battery chapters
    provide the only complete reference proof ladder: gap existence, one-step
    safety preservation, the episode-level risk envelope, no-free-safety,
    certificate horizon results, fallback existence, and graceful-degradation
    dominance.  This is the governance-critical proof spine of the
    dissertation.

  \item \textbf{The universal extension is formal, not rhetorical.}  T10
    proves that the degraded-observation risk envelope has the correct
    first-order scaling law, and T11 states the exact structural obligations
    under which the battery argument transfers to a new domain.  The
    dissertation therefore claims a domain-agnostic kernel for safe control
    under degraded observation, not a solved theory of generic universal safe
    control.

  \item \textbf{The safety bound is quantitative and verifiable.}
    Theorem~\ref{thm:core-bound} on the battery reference surface establishes
    $\mathbb{E}[V] \le \alpha(1-\bar{w})T$ under Assumptions A1--A7.  The
    T10 lower bound shows the same dependence is minimax-rate tight up to a
    universal constant.  The bound is therefore both operational and
    theoretically non-accidental.

  \item \textbf{Cross-domain evidence is tiered and explicit.}  Industrial
    process control and healthcare are defended instantiations of the same
    kernel.  Autonomous vehicles remain proof-candidate because of the
    constraint-class mismatch analysed in the monograph's universality
    chapter.  Navigation remains
    shadow-synthetic portability evidence, and aerospace remains
    experimental.  These rows support the transfer claim at different levels;
    they do not carry the same proof burden as battery.

  \item \textbf{Traceability is end-to-end.}  Each major theorem group is
    tied to code anchors and locked artifacts.  The integrated theorem surface
    register (Appendix~\ref{app:integrated-theorem-surface-register}) maps
    claim to code to artifact for every strongest-claim result.
\end{enumerate}

\section{What Is Established Empirically}

The empirical package supports the theorem program in a tiered way:
\begin{enumerate}
  \item the battery row is the reference witness where the theorem ladder and
    executable artifacts are both deepest;
  \item industrial and healthcare defend the transfer story with locked
    replay, SIL pass, and material TSVR reduction;
  \item AV, navigation, and aerospace show the practical boundary of the
    kernel and help locate where additional control design or stronger data is
    still required.
\end{enumerate}

\section{How Papers~2--6 Fit the Monograph Boundary}

The monograph keeps Paper~1 as the only full proof witness, but it does not
hide the rest of the program in footnotes.  Papers~2--6 are explicit and
first-class in the dissertation under bounded roles:

\begin{enumerate}
  \item Paper~2 is the bounded temporal extension.
  \item Paper~3 is the bounded graceful-fallback benchmark surface.
  \item Paper~4 is the evaluation-contract layer.
  \item Paper~5 is the bounded composition counterexample layer.
  \item Paper~6 is the bounded runtime-governance layer.
\end{enumerate}

This is the correct monograph boundary.  The program is broad and coherent,
but the proof burden is stratified rather than flattened.

\section{Theorem-to-Empirical Correspondence}

Table~\ref{tab:theorem-empirical-correspondence} maps the four core theorems
to their empirical confirmations across domains.

\begin{table}[htbp]
\centering
\caption{Core theorem ladder and empirical confirmation across the tiered six-row runtime package.}
\label{tab:theorem-empirical-correspondence}
\begin{tabular}{p{3.8cm}p{5.5cm}p{4.5cm}}
\toprule
\textbf{Theorem} & \textbf{Prediction} & \textbf{Empirical confirmation} \\
\midrule
T1 -- OASG Existence
  & TSVR $> 0$ under quality-ignorant controller
  & 2.78\%--21.53\% baseline TSVR across the locked non-battery runtime rows \\
\addlinespace
T2 -- Safety Preservation
  & $\Pr[\phi_{\text{true}}(a_t) = 1] \ge 1-\alpha$ per step
  & TSVR $= 0$\% across the three proof-validated domains (energy management, industrial, healthcare) \\
\addlinespace
T3 -- Core Bound
  & $\mathbb{E}[V] \le \alpha(1-\bar{w})T$
  & Runtime TSVRs stay within the conservative battery-anchored envelope \\
\addlinespace
T4 -- No-Free-Safety
  & Any safe-in-observation-only policy has TSVR $> 0$
  & Rule-based baselines confirm non-zero TSVR on the locked runtime rows \\
\bottomrule
\end{tabular}
\end{table}

\section{What the Result Means}

The ORIUS result is strong because it ties every major theorem object to an
operational consequence.  Hidden violations are not abstract pathologies; they
correspond to unsafe physical outcomes such as damaging battery discharge,
vehicle over-speed, process power exceedance, ICU alarm suppression failure,
or aircraft envelope breach.  Reliability-aware interval inflation is not
only a calibration device; it widens the safe-action margin in proportion to
degradation severity.  Action repair is not merely a projection lemma; it is
the mechanism that prevents unsafe actuation when the observation channel is
too optimistic.  Certificates are not audit theatre; they make the causal
assumptions behind each repaired action explicit.

\section{What Is Not Established}

The monograph does \emph{not} establish the following:
\begin{enumerate}
  \item a generic universal stability theory for all safe-control problems;
  \item equal proof closure across the six evaluated domains;
  \item production-readiness for every adapter without further engineering and
    validation;
  \item adversarial completeness beyond the bounded degradation and replay
    assumptions stated in Chapter~\ref{ch:assumptions};
  \item a replacement for full CBF, robust MPC, or viability theory in their
    native settings.
\end{enumerate}

\section{Scope of the Final Claim}

The final claim is precisely bounded:

\begin{quote}
Under degraded telemetry, observed safety can diverge from true safety.
ORIUS addresses this gap with a domain-agnostic repair-and-certify kernel
built on reliability-aware uncertainty inflation, safe-action repair, and
causal runtime certificates.  In this monograph, battery is the complete
formal reference lineage; industrial and healthcare are defended
instantiations; AV remains proof-candidate; aerospace remains experimental;
and navigation remains shadow-synthetic portability evidence.
\end{quote}

This claim is supported by code, artifacts, and chapter-level alignment.  It
does not extend to production deployment, adversarial completeness, or
heterogeneous multi-domain composition without further engineering validation
(Chapter~\ref{ch:outside-current-evidence}).

\section{Summary}

The dissertation now ends at the right level of precision.  Its main theorem
contribution is a domain-agnostic degraded-observation safety kernel with one
full proof witness, one necessity theorem, one lower-bound theorem, and one
explicit transfer theorem.  The battery domain remains the complete proof
surface.  Industrial and healthcare defend the same kernel at a lower proof
burden.  AV, navigation, and aerospace map the current frontier of the claim
rather than inflating it.  That is monograph-grade scope: narrow enough to be
defensible, broad enough to matter.
