\chapter{What Remains Outside the Current Evidence}
\label{ch:outside-current-evidence}

The ORIUS framework is stronger when it is explicit about what has \emph{not}
been shown.  This chapter records the outer boundary of the current evidence
package so the thesis does not borrow credibility from work that has not yet
been executed.

\section{Not Claimed}

The current package does not claim:

\begin{enumerate}
  \item deployment safety in any production system without domain-specific
    engineering validation and regulatory approval,
  \item full physical hardware validation on a live safety-critical stack
    (battery, aircraft, ICU monitor, or vehicle),
  \item adversarial completeness against every possible spoofing or
    Byzantine sensor strategy,
  \item heterogeneous multi-asset composition guarantees across domain
    boundaries (e.g., battery + AV + medical device in a shared fleet),
  \item a final certificate-expiration law beyond the conservative blackout
    hold behavior locked in Chapter~\ref{ch:cert-half-life-blackout}, or
  \item coverage guarantees for fault families outside the calibration set
    (out-of-distribution faults, novel sensor failure modes).
\end{enumerate}

\section{Still Partial}

Several thesis-hardening layers exist in meaningful but still bounded form:

\begin{itemize}
  \item Aging and asset preservation are represented through proxy tables and
    calibration design rather than long-horizon field degradation data.
  \item Fleet composition is implemented for the two-battery shared-transformer
    case, not for a broader heterogeneous fleet.
  \item The HIL chapter has an auditable runtime rehearsal, but not a full
    external-bench package with independent device timing and relay shutdown.
  \item Navigation domain validation is simulation-backed rather than
    field-telemetry-backed.
\end{itemize}

\section{Per-Domain Next Closure Obligations}

Each validated domain leaves one explicit next closure obligation for the
broader ORIUS program.

\begin{description}
  \item[Battery.] Adaptive conformal recalibration under non-stationary
    energy demand (e.g., demand-response events, grid frequency events).
    Online decay-rate estimation for the certificate half-life model to
    replace the current fixed decay parameter.

  \item[Autonomous Vehicles.] Multi-vehicle interaction under shared
    telemetry infrastructure: extending OQE scoring to handle V2X-fused
    observations and validating DC3S repair in cooperative platooning
    scenarios where one vehicle's action affects another's safety margin.

  \item[Industrial Process Control.] Multi-process constraint coupling:
    when multiple actuators share a power budget, the safe-action projection
    must solve a coupled feasibility problem.  Extending the RAC-Cert
    inflation to handle joint constraint surfaces is the key technical step.

  \item[Medical Monitoring (ICU Vitals).] Multi-patient fleet monitoring
    with shared calibration surface: OQE reliability scores should account
    for cross-patient covariate shift.  Alarm-fatigue mitigation via
    certificate-gated alert suppression is a safety-critical extension.

  \item[Aerospace.] Multi-aircraft separation assurance integrated with
    ATC intent data.  The current single-aircraft envelope model does not
    account for traffic advisory corrections; extending the feasible action
    set to encode separation minima would turn ORIUS into a conflict-resolution
    advisory layer on top of TCAS.

  \item[Navigation.] Transition from simulation-validated to
    field-telemetry-backed evidence.  GPS-denied scenarios (underground
    parking, dense urban canyons) and LiDAR-based OQE feature extraction
    are the next data-collection targets.
\end{description}

\section{Precise Failure Boundaries}
\label{sec:precise-failure-boundaries}

A mature safety argument requires not only what the system proves, but
\emph{exactly where it fails}.  The following boundaries are not vague
caveats but computable thresholds derived from the framework structure.

\paragraph{1. Feasibility exhaustion.}
As $w_t \to 0$, the margin $m_t = q_t / (w_t + \varepsilon) \to \infty$,
causing the tightened action set $\mathcal{A}_t$ to shrink toward the
empty set.  DC3S detects this condition and invokes the safe-landing
fallback policy (Chapter~\ref{ch:graceful-degradation}).  The feasibility
exhaustion threshold is domain-specific: for energy management, $w_t < 0.05$
triggers zero-feasibility at nominal SOC mid-range ($\mathrm{SOC} \approx 0.5$).
Domains with wider constraint sets (navigation, industrial) tolerate lower
$w_t$ before feasibility is exhausted.  Assumption~A3 (feasible safe repair)
is formally satisfied whenever $\mathcal{A}_t \ne \emptyset$; the
safe-landing protocol activates when this condition is violated.

\paragraph{2. OQE self-compromise.}
The OQE scores telemetry quality using anomaly detection and inter-arrival
timing.  If the OQE's own input channel is manipulated --- for example, an
adversary injects a valid-looking packet with a false reliability signal
--- the shield adapts to the wrong $w_t$.  Active probing
partially addresses this by detecting injected patterns, but it does not
constitute a complete adversarial hardness proof.  The OQE's trust model is
that the channel is Byzantine-fault-free; adversarial completeness under a
Dolev--Yao attacker remains outside the current evidence package.

\paragraph{3. Calibration distributional shift.}
The conformal guarantee holds under approximate exchangeability between the
calibration set and the deployment distribution.  A sustained distributional
shift that invalidates the calibration residuals will cause the empirical
$\mathrm{PICP}_{90}$ to drop below $1 - \alpha$.  The finite-sample
correction $\varepsilon_n = O(1/\sqrt{n})$ applies within-distribution
only; out-of-distribution shift is not covered.  The operational monitoring
protocol (Chapter~\ref{ch:deployment-path-verification}) specifies how to
detect and respond to this condition by triggering conformal recalibration
on a fresh holdout split.

\paragraph{4. Temporal fault correlation.}
The Core Bound $\mathbb{E}[V] \le \alpha(1-\bar{w})T$ uses linearity of
expectation and assumes step-level independence conditioned on $w_t$.
Under temporally correlated faults --- for example, a 30-minute sensor
dropout storm --- per-step violation indicators may cluster in time.  The
expectation bound remains valid because linearity of expectation does not
require independence; however, a PAC-style concentration bound under
temporal correlation would require a Hoeffding--Azuma argument adding a
term of order $\sqrt{T \ln(1/\delta)}$.  The current finite-sample
correction handles i.i.d.\ calibration residuals and does not account
for within-storm correlation structure.

\section{Why This Boundary Matters}

The ORIUS framework is meant to be hard to challenge because it is internally
coherent and multi-domain consistent, not because it quietly widens its scope.
By naming what remains outside the current evidence, the thesis protects the
core multi-domain result from being diluted by larger claims the present
evidence package does not yet support.  Correctness of the safety bound and
equivalence of domain performance are the locked results; scalability to
production, adversarial completeness, and heterogeneous composition are the
open research agenda that the framework is designed to grow into.

\section{Summary}

The current evidence package is deep across the five locked-telemetry domains,
but is intentionally bounded in scope.  Energy management, autonomous vehicles,
industrial process control, medical monitoring, and aerospace each contribute
replay-backed empirical evidence under the shared fault schedule; navigation
adds simulation-backed portability evidence rather than field-telemetry-backed
validation.  None contributes deployment-grade certification.  That is the
correct scientific posture for a framework-contribution thesis at this stage.
