\chapter{Deployment Path and Verification Discipline}
\label{ch:deployment-path-verification}

The ORIUS framework closes with a deployment path that is operationally
serious but evidence-bounded.  The goal is not to announce production
deployment.  The goal is to specify how deployment would proceed for any
domain without breaking the claim discipline established in this thesis.

\section{Universal Deployment Readiness Checklist}

Before deploying DC3S in any domain, the following conditions must be
verified:

\begin{enumerate}
  \item \textbf{Domain adapter registration.}  A concrete \texttt{DomainAdapter}
    subclass implementing all five pipeline stages (ingest, OQE, uncertainty,
    tighten, repair, certify) must be registered in the domain registry.
    All adapter unit tests must pass.

  \item \textbf{Calibration surface validation.}  A domain-specific calibration
    split must achieve $\mathrm{PICP}_{90} \ge 0.90$.  The calibration
    residuals must satisfy approximate exchangeability with the deployment
    telemetry distribution.

  \item \textbf{OQE threshold configuration.}  The reliability score $w_t$
    thresholds (alert level, blackout level) must be calibrated against
    domain-specific telemetry fault rates.  Stale, spike, and dropout
    fault injection tests must confirm OQE detection lag within the
    Assumption~A6 bound.

  \item \textbf{Constraint and feasibility verification.}  The safety predicate
    $\phi(x_t)$, the constraint set $\mathcal{U}$, and the margin function
    $m_t = q_t / (w_t + \varepsilon)$ must be reviewed by domain experts.
    Assumption~A3 (feasible safe repair) must be confirmed: $\mathcal{A}_t$
    must be non-empty whenever the domain is in a safe physical state.

  \item \textbf{End-to-end benchmark replay.}  A CPSBench replay with separate
    truth and observed trajectories must be executed over at least 1000
    telemetry steps under the target fault schedule.  The empirical TSVR
    must satisfy the Core Bound ($\mathbb{E}[V] \le \alpha(1-\bar{w})T$).

  \item \textbf{Artifact manifest traceability.}  Every output artifact
    (tables, figures, certificates) must be registered in the artifact
    manifest with its generating script, run parameters, and version hash.
    Re-running the named script must reproduce the artifact within numerical
    tolerance.
\end{enumerate}

\section{Universal Deployment Ladder}

For any CPS domain, the deployment ladder is:

\begin{enumerate}
  \item \textbf{Benchmark replay}---observed versus true trajectory separation,
    TSVR measurement, Core Bound verification.
  \item \textbf{Software-in-loop (SIL) rehearsal}---DC3S running against a
    simulated or replayed telemetry stream, with full certificate emission
    and manifest logging.
  \item \textbf{Hardware-in-loop (HIL) validation}---real telemetry transport,
    physical actuator interface, relay behavior under injected faults.
  \item \textbf{Supervised pilot operation}---bounded fallback policies,
    certificate-gated alert suppression, human-in-the-loop audit.
  \item \textbf{Wider deployment}---only after all prior layers remain
    manifest-traceable and no regression in TSVR or certificate integrity
    has been observed.
\end{enumerate}

The energy management domain is currently at Stage~1--3.  No domain in
this thesis has reached Stage~4.  Stages~4--5 require domain-specific
regulatory engagement beyond the scope of this framework contribution.

\section{Operational Monitoring}

Once deployed in any live tier, DC3S requires continuous monitoring of:

\begin{itemize}
  \item \textbf{OQE trend.}  A sustained drop in the average reliability
    score $\bar{w}_t$ below the alert threshold ($w < 0.60$ for energy
    management, domain-specific for others) should trigger a calibration
    review and possible shield recalibration.

  \item \textbf{Intervention rate drift.}  An intervention rate (IR) that
    exceeds the baseline by more than 10 percentage points may signal
    distribution shift in the input telemetry.  The conformal quantile
    $q_t$ should be re-estimated on a fresh calibration split.

  \item \textbf{Certificate half-life.}  The runtime certificate validity
    window should be monitored using the decay model in
    Chapter~\ref{ch:cert-half-life-blackout}.  Certificates expiring faster
    than expected indicate elevated fault rates or OQE degradation.

  \item \textbf{TSVR monitoring.}  Where ground-truth state can be accessed
    post-hoc (e.g., from SCADA historian data or sensor reconciliation),
    the empirical TSVR should be computed at each reporting period and
    compared to the Core Bound prediction.
\end{itemize}

\section{Rollback Protocol}

If any monitoring indicator exceeds its alarm threshold, the following
rollback protocol applies:

\begin{enumerate}
  \item \textbf{Certificate hold.}  Suspend certificate emission and fall
    back to the safe-landing policy (Chapter~\ref{ch:graceful-degradation}).
  \item \textbf{Fault diagnosis.}  Identify the root cause of the degradation
    (telemetry faults, model drift, constraint boundary change) using the
    OQE diagnostic flags.
  \item \textbf{Calibration refresh.}  Re-run conformal calibration on a
    fresh holdout split reflecting current operating conditions.
  \item \textbf{Re-verification.}  Execute a full SIL pass under the refreshed
    calibration before resuming certificate emission.
\end{enumerate}

\section{Verification Discipline}

Every stage of the deployment ladder must preserve four invariants:

\begin{itemize}
  \item a named script path that can regenerate each output artifact,
  \item a run-scoped artifact with stable, version-controlled location,
  \item a manifest entry tying the artifact to the corresponding claim, and
  \item an explicit statement of what the result does \emph{not} prove.
\end{itemize}

This discipline is not only an engineering convenience.  It is a
prerequisite for any safety-critical certification process.  A domain
that cannot reproduce its validation artifacts on demand cannot make a
defensible safety argument to a regulatory body.

\section{Summary}

The ORIUS deployment path is incremental and auditable.  Each new layer
inherits the theorem-to-code-to-artifact traceability that governs the
current evidence package.  Widening to production---within any domain---is
a matter of climbing the deployment ladder step by step, with each rung
verified against the Core Bound and logged in the artifact manifest.
