\chapter{Conclusion}
\label{ch:conclusion}

The central finding of this thesis is concrete: when telemetry degrades,
a controller can satisfy its observed constraints while violating the true
physical envelope of the system.  This hidden gap---the
Observation-Action Safety Gap (OASG)---is measurable, operationally
meaningful, and correctable.  The dissertation's final claim is now
sharper than a generic ``six-domain framework'' result: ORIUS isolates a
domain-agnostic repair-and-certify kernel for safe control under degraded
observation, proves that kernel completely on the battery reference surface,
and then carries it across a tiered evidence ladder whose rows do not share
the same proof burden.

\section{Kernel-Level Outcome}

ORIUS shows that the OASG phenomenon is not a battery-specific artefact.
It is a structural consequence of closed-loop control under degraded
telemetry.  The dissertation now states that result at three levels:
\begin{enumerate}
  \item \textbf{Battery proof witness (T1--T8).} The battery chapters prove
    gap existence, one-step safety preservation, the core bound, the
    no-free-safety barrier, and the temporal extensions on one complete
    reference surface.
  \item \textbf{Monograph extension (T9--T11).} The universality chapter adds
    a necessity theorem, a lower-bound theorem, and a typed structural
    transfer theorem.
  \item \textbf{Tiered cross-domain instantiation.} The shared kernel is then
    exercised across defended, proof-candidate, shadow-synthetic, and
    experimental rows without claiming equal proof closure.
\end{enumerate}

\section{Domain Coverage and Empirical Results}

The tiered six-row runtime package was evaluated under a unified fault schedule
(dropout~15\%, spike~8\%, stale~10\%) at $N = 2000$ fault-stressed steps:

\begin{itemize}
  \item \textbf{Energy Management} (reference, proof-validated): 3.90\% baseline
    TSVR $\rightarrow$ 0.0\% under DC3S (3.36\% repair rate).
    Full theorem-to-code-to-artifact lineage established in Parts~I--II.
  \item \textbf{Industrial Process Control} (proof-validated): 21.53\% $\rightarrow$
    0.0\% (91.7\% repair rate).
  \item \textbf{Medical Monitoring (ICU Vitals)} (proof-validated): 6.25\%
    $\rightarrow$ 0.0\% (14.6\% repair rate).
  \item \textbf{Navigation} (simulation-backed): 18.06\% $\rightarrow$ 0.69\%
    (100\% repair rate); portability evidence outside the proof-validated tier.
  \item \textbf{Autonomous Vehicles} (proof-candidate): 2.78\% baseline TSVR
    $\rightarrow$ 2.78\% under DC3S.  High repair rate (98.6\%) indicates
    shield activity but TSVR does not reach zero due to constraint structure
    mismatch; classified proof-candidate pending further investigation.
  \item \textbf{Aerospace (Flight Envelope)} (experimental): 9.72\% $\rightarrow$
    9.72\% (13.9\% repair rate); evaluated on an experimental surface pending
    additional domain data.
\end{itemize}
Higher-budget replay shows strong TSVR reduction across several non-battery
rows, but the final claim boundary is still governed by the locked tier table
rather than raw replay metrics alone.  Pipeline latency remains
sub-millisecond across the five forecasting-capable runtime rows
(P95~$<$~0.1~ms), confirming ORIUS is suitable for fast-control-loop
applications.

\section{Formal Apparatus and Theorem Ladder Scope}

The full formal apparatus of this dissertation consists of 18 theorems,
5 lemmas, 7 propositions, 10 corollaries, 11 definitions, and 16 standing
assumptions---67 formal items in total.  Two foundational theorems are
established in Part~I before the formal ladder:

\begin{itemize}
  \item \textbf{S1} (Theorem: Existence of the Illusion under Dropout,
    Chapter~\ref{ch:safety-illusion}): proves that a
    controller can satisfy observed constraints while violating true
    physical safety under any positive dropout rate.
  \item \textbf{S2} (Theorem: Informal Safety Guarantee of DC3S,
    Chapter~\ref{ch:safety-illusion}): establishes the
    informal correctness guarantee that motivates the DC3S design.
\end{itemize}

The eight principal theorems of the formal DC3S ladder (T1--T8) are then
stated, proved, and connected to code artefacts in
Chapters~\ref{ch:theorem-oasg}--\ref{ch:temporal-extensions}.  The final
monograph layer adds T9--T11: observation-aware control is structurally
necessary under persistent degradation, the $\alpha(1-\bar{w})T$ envelope has
the correct first-order scaling law, and transfer beyond battery depends on
explicit structural obligations rather than an informal adapter story.
The integrated theorem gate therefore covers both the battery ladder and the
monograph extension surface with locked code anchors and artefact traceability
(see \texttt{reports/publication/integrated\_theorem\_gate.csv}).

\section{Final Boundary}

This thesis does not claim that ORIUS is ready for deployment in any
safety-critical domain without domain-specific validation.  It claims
that:

\begin{enumerate}
  \item The OASG phenomenon is real and measurable in any CPS with
    degraded telemetry.
  \item The DC3S architecture provides a principled correction mechanism
    that is domain-agnostic by construction.
  \item Shared replay evaluation shows consistent runtime behavior across
    the six-row package, while claim strength follows the locked tier table
    rather than raw replay metrics alone.
  \item The battery domain provides the complete reference instantiation
    with full theorem-to-artifact lineage; industrial and healthcare are
    the proof-validated peers; AV is proof-candidate; aerospace is
    experimental; and navigation provides simulation-backed portability
    evidence.
  \item The remaining open gaps are explicit and localized rather than
    foundational: AV still needs a larger repair surface, navigation still
    lacks locked real telemetry, and aerospace still lacks a defended
    promotion contract.
\end{enumerate}

\section{Closing Statement}

The dissertation ends at a defensible control-theory boundary: battery is the
full proof witness, T9--T11 make the universal kernel explicit, industrial and
healthcare defend transfer under the current evidence lock, and the remaining
rows are kept at lower tiers until their open structural gaps are discharged.
That is the intended thesis-level result: one precise degraded-observation
safety kernel, one complete reference proof surface, and one honest
cross-domain frontier rather than six domains with equal guarantees.
