\chapter{A Domain-Agnostic Kernel for Safe Control Under Degraded Observation}
\label{ch:universality-completeness}

This chapter is the control-theory extension of the dissertation.  Its
purpose is not to add more runtime rows; it is to isolate the
domain-agnostic kernel that the rest of the monograph instantiates.  The
central claim is deliberately narrow: under degraded observation, safe
control is governed by an observation-consistent state set, a
reliability-conditioned tightening law, a repair map, and a causal
certificate surface.  Battery remains the only full proof reference
surface.  Industrial and healthcare remain defended instantiations.
Autonomous vehicles, navigation, and aerospace remain lower-tier evidence.

The point of the chapter is to make that claim precise enough that a reader
can separate three things cleanly: what the reference-row proof already establishes,
what the universal kernel adds, and where the current transfer boundary still
stops.  The implementation follows the same boundary.  The canonical
production kernel realizes the typed repair-and-certify contract, while the
formal arguments below isolate the control objects that must exist in any
domain instantiation.

% ----------------------------------------------------------------
\section{Universal Problem Statement and Information Pattern}
\label{sec:universal-kernel-model}
% ----------------------------------------------------------------

We consider a partially observed, safety-constrained control system on a
discrete horizon $t = 0,1,\dots,T-1$:
\[
  x_{t+1} = f(x_t, a_t, \omega_t),
  \qquad
  y_t = h(x_t, \nu_t),
  \qquad
  w_t = \psi(y_{0:t}),
\]
where $x_t \in \mathcal{X} \subseteq \mathbb{R}^n$ is the true plant state,
$a_t \in \mathcal{A} \subseteq \mathbb{R}^m$ is the applied control,
$y_t$ is the telemetry channel output, $\omega_t$ is admissible plant
disturbance, $\nu_t$ is telemetry corruption, and $w_t \in [0,1]$ is the
reliability score extracted from the observation history.  The safe state set
is
\[
  \mathcal{S} = \{x \in \mathcal{X} : \phi(x) \ge 0\},
\]
and the violation count is
\[
  V_T = \sum_{t=0}^{T-1} \mathbf{1}[x_t \notin \mathcal{S}].
\]

The candidate controller sees only the observation history
$\mathcal{H}_t = \sigma(y_{0:t})$ and proposes
$a_t^{\mathrm{cand}} = \pi_t(\mathcal{H}_t)$.  The ORIUS kernel then builds
an observation-consistent state set $U_t$, derives a tightened safe-action
set $\mathcal{A}_t^{\mathrm{safe}}(U_t)$, repairs the candidate action into
$a_t^{\mathrm{safe}}$, and emits a certificate summarizing the causal data
used at step $t$.

\begin{definition}[Degraded-observation safe-control problem]
\label{def:degraded-observation-safe-control}
Given only the observation history $\mathcal{H}_t$, the reliability score
$w_t$, and a nominal candidate action $a_t^{\mathrm{cand}}$, construct a
measurable repair-and-certify map such that:
\begin{enumerate}
  \item the repaired action depends only on causal information available at
    time $t$;
  \item the repaired action preserves safety for all states consistent with
    the degraded observation, up to the stated miscoverage level $\alpha$;
  \item the resulting episode-level violation count admits an explicit bound
    in terms of $\alpha$, the reliability process, and the horizon.
\end{enumerate}
\end{definition}

For the control-theory reader, this chapter answers three questions beyond the
reference proof ladder:

\begin{table}[htbp]
\centering
\caption{Theorem questions answered by the monograph extension.}
\label{tab:universal-theorem-questions}
\begin{tabular}{p{1.5cm}p{4.7cm}p{5.1cm}}
\toprule
\textbf{Theorem} & \textbf{Control-theory question} & \textbf{Role in the monograph} \\
\midrule
T9 & Is reliability-aware repair structurally necessary, or merely useful? & Necessity theorem for degraded-observation control \\
\addlinespace
T10 & Does the ORIUS envelope have the right scaling law, or is it loose? & Lower-bound theorem for degraded-observation risk \\
\addlinespace
T11 & Which structural obligations let the reference proof pattern transfer? & Template transfer theorem for new domains \\
\bottomrule
\end{tabular}
\end{table}

The dependency order is strict.  T1--T8 establish the reference proof
ladder.  T9 proves that ignoring reliability breaks safety structurally.
T10 shows that the $\alpha(1-\bar{w})T$ scaling cannot be removed by a better
controller alone.  T11 then identifies the obligations under which the same
repair-and-bound structure transfers across domains.

% ----------------------------------------------------------------
\section{Observation-Aware Control Is Structurally Necessary (T9)}
\label{sec:t9-universal-impossibility}
% ----------------------------------------------------------------

Theorem~\ref{thm:no-free-safety} shows, on the reference proof surface, that
quality-ignorant safety logic can fail even when observed-state evaluation
looks safe.  The broader universality claim is that this pathology is not a
single-domain accident; it is a structural consequence of degraded observation.

\begin{theorem}[Universal Impossibility, T9]
\label{thm:universal-impossibility}
Consider a closed-loop controller that does not adapt its safety margin to the
telemetry reliability process $w_t$.  Under persistent observation degradation
with non-zero fault rate and under the standing boundary-reachability
assumptions used throughout this dissertation, there exists a constant
$c > 0$ such that
\[
  \mathbb{E}\!\left[V_T\right] \;\ge\; c \, d \, T,
\]
where $d$ is the mean degradation rate and $V_T$ is the true-state violation
count over the horizon.
\end{theorem}

\begin{proof}[Proof (martingale--$\varphi$-mixing argument)]
The proof proceeds in three steps.

\medskip
\noindent\textbf{Step~1: $\varphi$-mixing fault process.}
By Assumption~A6$'$ (Chapter~\ref{ch:assumptions}), the fault injection
process $\{F_t\}$ is $\varphi$-mixing with coefficient $\varphi(k)\to 0$
as $k\to\infty$.  Concretely, $\varphi(k) \le C_\phi\,\rho^k$ for
constants $C_\phi > 0$ and $\rho \in (0,1)$, so a buffer gap of
$\tau_{\mathrm{mix}} = \lceil \log(1/\varepsilon)\,/\,|\log\rho| \rceil$
steps makes any two windows $\varepsilon$-approximately independent in
total variation.

\medskip
\noindent\textbf{Step~2: Separated windows and per-window risk.}
Partition $[1,T]$ into $k$ analysis windows of length $\ell$ separated
by buffer gaps of length $\tau_{\mathrm{mix}}$, choosing
$\ell = \lfloor T\,/\,(2k) \rfloor$ so that $k \le T\,/\,
(2(\ell + \tau_{\mathrm{mix}}))$.  The reference-domain
no-free-safety witness (Theorem~\ref{thm:no-free-safety}) is applied
inside each window: if the controller does not adapt its safety margin
to the OQE signal $w_t$, the observation gap drives the true state
across the constraint boundary while the observed state appears feasible.
The per-window violation probability satisfies
$p_{\mathrm{win}} \ge c\,d$ for a positive constant $c > 0$ that depends
on the boundary-reachability conditions of Assumption~A4.

\medskip
\noindent\textbf{Step~3: Martingale concentration.}
Define $D_j = Z_j^{\mathrm{win}} - \mathbb{E}\!\bigl[Z_j^{\mathrm{win}}
\mid \mathcal{F}_{(j-1)(\ell+\tau_{\mathrm{mix}})}\bigr]$, where
$Z_j^{\mathrm{win}} \in \{0,1\}$ indicates a violation in window $j$.
The sequence $\{D_j\}$ is a martingale difference sequence with
$|D_j|\le 1$.  By the Azuma--Hoeffding inequality,
\[
  \Pr\!\Bigl[\,\textstyle\sum_{j=1}^k D_j \le -t\,\Bigr]
  \;\le\; \exp\!\Bigl(-\tfrac{2t^2}{k}\Bigr).
\]
Since $\mathbb{E}\bigl[\sum_j Z_j^{\mathrm{win}}\bigr] \ge k\,c\,d$,
setting $t = k\,c\,d/2$ gives
\[
  \Pr\!\Bigl[\,V_T \;\le\; \tfrac{k\,c\,d}{2}\,\Bigr]
  \;\le\;
  \exp\!\Bigl(-\tfrac{k\,c^2\,d^2}{2}\Bigr)
  \;=\;
  \exp\!\Bigl(-\Omega(dT)\Bigr).
\]
Therefore, with probability at least $1 - e^{-\Omega(dT)}$, the
violation count satisfies $V_T \ge \Omega(dT)$.  Taking expectations
confirms $\mathbb{E}[V_T]\ge\Omega(dT)$, completing the proof.
\end{proof}

\begin{remark}
The martingale argument is strictly stronger than a Borel--Cantelli
appeal: Borel--Cantelli requires pairwise independence across windows,
which does not hold when the true state carries over between adjacent
windows.  The $\varphi$-mixing approach replaces exact independence with
approximate independence quantified by $\varphi(\tau_{\mathrm{mix}})$;
the buffer gaps absorb the coupling cost without changing the
$\Omega(dT)$ scaling.
\end{remark}

T9 is the necessity theorem for the universal kernel: whenever safety depends
on a corrupted observation channel, a controller that ignores reliability
cannot remove the violation scaling by algorithmic tuning alone.  The defended
evidence remains tiered.  Battery anchors the proof lineage directly;
industrial and healthcare support the same runtime story under defended
replay; the weaker rows remain portability evidence rather than
theorem-validating peers.

% ----------------------------------------------------------------
\section{Boundary-Indistinguishability Lower Bound (T10)}
\label{sec:t10-lower-bound}
% ----------------------------------------------------------------

The core ORIUS bound from the reference domain is
\[
  \mathbb{E}\!\left[V_T\right] \;\le\; \alpha(1-\bar{w})T,
\]
where $\alpha$ is the miscoverage level and $\bar{w}$ is the mean reliability.
The monograph extension asks whether that envelope is merely conservative or
whether it captures the right scaling law.

\begin{theorem}[Boundary-indistinguishability lower bound, T10]
\label{thm:information-theoretic-lower-bound}
Fix a horizon $T$.  For each step $t$, assume there exists a boundary-testing
subproblem with two latent hypotheses $H_{0,t}$ (safe side of the boundary)
and $H_{1,t}$ (unsafe side of the boundary) such that:
\begin{enumerate}
  \item the controller acts only through the degraded observation channel;
  \item the observation laws $P_{0,t}$ and $P_{1,t}$ satisfy
    $\mathrm{TV}(P_{0,t},P_{1,t}) \le w_t$;
  \item the unsafe-side prior satisfies $\Prob(H_{1,t}) \ge p_t$ for some
    boundary-mass sequence $p_t \in [0,1]$.
\end{enumerate}
Then every controller $\pi$ satisfies
\[
  \mathbb{E}\!\left[V_T(\pi)\right]
  \;\ge\;
  \frac{1}{2}\sum_{t=1}^T p_t(1-w_t).
\]
In particular, if $p_t \ge \alpha/2$ for all $t$, then
\[
  \mathbb{E}\!\left[V_T(\pi)\right]
  \;\ge\;
  \tfrac{1}{4}\,\alpha(1-\bar{w})T.
\]
\end{theorem}

\begin{proof}
Fix a step $t$ and consider the controller's binary decision problem on the
safe/unsafe boundary subproblem.  By Le~Cam's lemma,
\[
  \max\!\bigl(\Prob[\psi_t\text{ errs}\mid H_{0,t}],\;
              \Prob[\psi_t\text{ errs}\mid H_{1,t}]\bigr)
  \;\ge\;
  \frac{1-\mathrm{TV}(P_{0,t},P_{1,t})}{2}
  \;\ge\;
  \frac{1-w_t}{2}.
\]
Whenever $H_{1,t}$ holds and the controller errs, the step contributes one
true-state violation, so
\[
  \Prob[Z_t = 1]
  \;\ge\;
  \Prob(H_{1,t}) \cdot \Prob[\psi_t\text{ errs}\mid H_{1,t}]
  \;\ge\;
  p_t \cdot \frac{1-w_t}{2}.
\]
Summing across the horizon and applying linearity of expectation yields
\[
  \mathbb{E}[V_T(\pi)]
  \;\ge\;
  \sum_{t=1}^T \Prob[Z_t = 1]
  \;\ge\;
  \frac{1}{2}\sum_{t=1}^T p_t(1-w_t).
\]
The special case $p_t \ge \alpha/2$ then implies the stated
$\tfrac{1}{4}\alpha(1-\bar{w})T$ frontier.  No stronger conclusion is claimed
without a domain-specific argument that justifies the boundary-mass sequence
$p_t$.
\end{proof}

\begin{remark}
T10 is scoped but substantive.  It is an information-theoretic lower bound
for any controller that only sees the degraded observation channel under the
explicit boundary-mass and indistinguishability hypotheses above.  It does
not claim a globally sharp minimax constant for every observation model, but
it does show that the $\alpha(1-\bar{w})T$ scaling cannot be dismissed as a
pure artifact of one repair heuristic.
\end{remark}

The resulting \emph{reliability-risk frontier} is useful even when read
conservatively:
\[
  \bar{w}^{\star}
  \;=\;
  1 - \frac{r^{\star}}{\alpha T},
\]
for a target episode-level violation budget $r^{\star}$.  The canonical
runtime helper for the conservative envelope lives in
\path{src/orius/universal_theory/risk_bounds.py}.  The richer frontier analysis
used in this chapter is implemented in
\path{src/orius/universal/reliability_safety_frontier.py}.  Together these
helpers make the central design message operational: degraded observation
imposes a quantitative infrastructure requirement, not just a qualitative
warning.

% ----------------------------------------------------------------
\section{Typed Structural Transfer and Failure-Mode Converse (T11)}
\label{subsec:adapter-contract}
% ----------------------------------------------------------------

The final dissertation claim no longer rests on a vague notion of ``adapter
portability.''  The reusable objects are now explicit both in the writing and
in code.

\begin{definition}[Universal adapter contract]
\label{def:adapter-contract}
A domain instantiation is compatible with the degraded-observation transfer
claim if it supplies:
\begin{enumerate}
  \item an observation packet and parsed state;
  \item a reliability assessment $w_t \in [0,1]$;
  \item an observation-consistent state set $U_t$;
  \item a domain safety specification and tightened safe-action set
    $\mathcal{A}_t$;
  \item a repair operator returning a safe action or admissible fallback;
  \item a causal certificate recording the repaired action, uncertainty state,
    reliability state, and assumption tags.
\end{enumerate}
In the simplified reference harness used for theorem miniatures, these objects
collapse to four adapter operations:
\texttt{observe()}, \texttt{uncertainty\_set()}, \texttt{repair()}, and
\texttt{fallback()}.
\end{definition}

The canonical typed implementation is provided by
\path{src/orius/universal_theory/contracts.py} and
\path{src/orius/universal_theory/kernel.py}.  The smaller reference harness in
\path{src/orius/universal/contract.py} packages the same idea into a compact
machine-checkable contract with five invariants:
\begin{description}
  \item[\textbf{Inv 1} Repair membership.] The repaired action lies in the
    tightened safe-action set.
  \item[\textbf{Inv 2} Monotone tightening.] Lower reliability never enlarges
    the admissible set.
  \item[\textbf{Inv 3} Reliability boundedness.] Reliability remains in
    $[0,1]$.
  \item[\textbf{Inv 4} Calibration consistency.] The inflation law preserves
    the base conformal surface at full reliability.
  \item[\textbf{Inv 5} Failure completeness.] Total observation failure
    collapses the admissible action set to the fallback regime.
\end{description}

\begin{theorem}[Typed structural transfer theorem, T11]
\label{thm:universality-completeness}
Let $\mathcal{D}$ be a domain instantiation for the degraded-observation
safe-control problem in Definition~\ref{def:degraded-observation-safe-control}.
Assume that, for every step $t$:
\begin{enumerate}
  \item the constructed observation-consistent state set satisfies
    $\Prob[x_t \in U_t \mid \mathcal{H}_t] \ge 1-\alpha$;
  \item the tightened safe-action set is sound for that uncertainty set, in
    the sense that every action in $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ keeps
    the next state inside $\mathcal{S}$ for all $x \in U_t$ and all
    admissible disturbance;
  \item the repair map returns an element of
    $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ whenever that set is non-empty;
  \item when the tightened safe-action set is empty, the fallback action is
    admissible and preserves safety from any current safe state.
\end{enumerate}
Then the repaired action produced by the ORIUS kernel satisfies the same
one-step safety-preservation statement as the reference theorem:
\[
  \Prob[x_{t+1}\in\mathcal{S}\mid \mathcal{H}_t] \ge 1-\alpha.
\]
\noindent In this theorem the four listed obligations are the entire transfer
surface.  No broader episode-level claim is inherited unless the new domain
also supplies an explicit predictable per-step risk budget as in
Corollary~\ref{cor:t11-aggregation}.
\end{theorem}

\begin{proof}
Fix a step $t$ and condition on the observation history $\mathcal{H}_t$.
On the event $E_t = \{x_t \in U_t\}$, Assumption~(2) implies that every action
in $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ preserves membership in the safe set
at the next step.  If the tightened safe-action set is non-empty, then
Assumption~(3) ensures that the repaired action returned by the kernel lies in
that set, hence preserves safety on $E_t$.  If the tightened safe-action set
is empty, then Assumption~(4) ensures that the fallback action preserves
safety from any current safe state.  Therefore,
\[
  \Prob[x_{t+1}\in\mathcal{S} \mid \mathcal{H}_t]
  \;\ge\;
  \Prob[E_t \mid \mathcal{H}_t]
  \;\ge\;
  1-\alpha,
\]
which is the one-step safety-preservation statement.
\end{proof}

\begin{corollary}[Episode aggregation under explicit per-step budgets]
\label{cor:t11-aggregation}
If, in addition, the instantiated domain supplies a predictable per-step risk
budget sequence $r_t$ satisfying
\[
  \Prob[x_{t+1}\notin\mathcal{S}\mid \mathcal{H}_t] \le r_t
\]
for all $t$, then
\[
  \mathbb{E}[V_T] \le \sum_{t=1}^T r_t.
\]
When a particular domain instantiation justifies the reference-row envelope
$r_t \le \alpha(1-w_t)$, the episode bound
$\mathbb{E}[V_T] \le \alpha(1-\bar{w})T$ follows immediately.
\end{corollary}

\begin{proof}
Let $Z_t=\mathbf{1}\{x_{t+1}\notin\mathcal{S}\}$.  The assumed bound implies
\[
  \mathbb{E}[Z_t \mid \mathcal{H}_t] \le r_t.
\]
Taking expectations and summing over $t$ gives
\[
  \mathbb{E}[V_T]
  =
  \sum_{t=1}^T \mathbb{E}[Z_t]
  \le
  \sum_{t=1}^T r_t.
\]
The reference-row envelope is the special case $r_t \le \alpha(1-w_t)$.
\end{proof}

\begin{proposition}[Failure of any transfer obligation breaks the reference proof pattern]
\label{prop:t11-counterexamples}
If any one of the four T11 obligations fails, then there exists a domain
instantiation for which the reference transfer argument breaks.
\end{proposition}

\begin{proof}
If obligation~(1) fails, the latent state may lie outside $U_t$ with
probability larger than $\alpha$, so the repair step can be safe only for the
wrong uncertainty set and the one-step safety argument no longer applies.
If obligation~(2) fails, an action may belong to the purported tightened set
while still allowing a successor outside $\mathcal{S}$ for some state in
$U_t$, so soundness is lost directly.  If obligation~(3) fails, the repair map
can return an action outside the tightened safe set even when a safe repaired
action exists, which again breaks the one-step argument.  If obligation~(4)
fails, the kernel can reach a state where the tightened safe-action set is
empty and no admissible fallback exists, so the transfer proof has no safe
release path.  In each case the failure is structural rather than rhetorical:
the missing obligation removes one of the exact proof steps used in the T11
forward direction.
\end{proof}

T11 is therefore a typed structural transfer theorem paired with an explicit
failure-mode converse, not a blanket declaration that every evaluated row
already satisfies those obligations.  The energy-management reference row provides the full reference
proof.  Other domains enter the monograph only when their instantiation
evidence is strong enough to support the corresponding one-step obligations,
and the stronger episode-level claim appears only when a domain-specific
per-step risk budget is justified explicitly.

\subsection{Machine-Checkable Contract Surfaces}

Two levels of executable checking now exist in the repository.

\begin{enumerate}
  \item \textbf{Production structural contract.}
    \path{src/orius/universal_theory/contracts.py} exposes the typed
    \texttt{DomainInstantiation} protocol together with a lightweight
    structural verifier.  The focused regression test
    \path{tests/test_universal_theory.py} checks canonical battery,
    industrial, healthcare, and vehicle adapters against that surface and then
    exercises the new universal step kernel end-to-end.
  \item \textbf{Reference invariant harness.}
    \path{src/orius/universal/contract.py} contains a more explicit invariant
    checker used for theorem miniatures and reductions.  The focused tests
    \path{tests/test_universal_contract.py} and
    \path{tests/test_unification.py} deliberately construct compliant and
    broken adapters to show which invariants matter.
\end{enumerate}

The simplified usage pattern is:

\begin{verbatim}
from orius.universal.contract import ContractVerifier

verifier = ContractVerifier(adapter, alpha=0.05)
verifier.check(z_t=np.array([0.0]), q_t=0.1)
\end{verbatim}

That harness is intentionally modest.  It does not replace the defended
domain-specific artifacts.  It makes the transfer theorem executable: the
reader can see the exact invariants, violate them, and observe the resulting
contract failure.

% ----------------------------------------------------------------
\section{Unification: ORIUS Subsumes Prior Frameworks}
\label{sec:unification}
% ----------------------------------------------------------------

\begin{theorem}[Unification]
\label{thm:unification}
The following prior safe-control families can be represented as special cases
of the ORIUS degraded-observation kernel:
\begin{enumerate}
  \item \textbf{Control Barrier Functions (CBF).} A fixed reliability policy
    $w_t \equiv 1$ together with barrier-defined action repair.
  \item \textbf{Robust MPC.} A constant reliability policy induced by a fixed
    tube radius together with projection onto the tube-tightened feasible set.
  \item \textbf{Safe RL.} A learned or posterior-derived reliability policy and
    a learned uncertainty surface coupled to a safety repair operator.
\end{enumerate}
ORIUS strictly generalises these reductions by allowing the reliability signal
to vary with the actual observation channel and by attaching an explicit
certificate surface to each repaired action.
\end{theorem}

\begin{proof}[Proof sketch]
The repository contains executable reference reductions in
\path{src/orius/universal/unification.py}.  \texttt{CBFAsORIUS} fixes
$w_t \equiv 1$ and uses barrier-style projection.  \texttt{RobustMPCAsORIUS}
fixes $w_t$ to the normalised tube radius and uses tube-style tightening.
These are not production replacements for full CBF or MPC solvers; they are
small reduction witnesses showing that ORIUS can express both within one
degraded-observation contract.  Safe RL fits the same pattern once posterior
concentration or uncertainty drives the reliability term.
\end{proof}

The unification is summarised in Table~\ref{tab:unification-comparison}.

\begin{table}[htbp]
\centering
\caption{Existing safe-control families as ORIUS reductions.}
\label{tab:unification-comparison}
\begin{tabular}{p{2.8cm}p{2.8cm}p{3.5cm}p{3.8cm}}
\toprule
\textbf{Framework} & \textbf{Reliability policy} & \textbf{Uncertainty / repair surface} & \textbf{ORIUS generalises by} \\
\midrule
CBF & $w_t \equiv 1.0$ & Barrier-defined safe set and projection & Adapting safety margins when telemetry degrades \\
\addlinespace
Robust MPC & $w_t \equiv c$ & Fixed tube tightening and projection & Using time-varying reliability instead of one constant tube level \\
\addlinespace
Safe RL & $w_t$ from posterior concentration & Learned uncertainty plus safety repair & Replacing model-specific confidence with an explicit degraded-observation contract \\
\addlinespace
\textbf{ORIUS / DC3S} & dynamic $w_t$ from OQE & Observation-consistent state set, repair, fallback, and certificate & \textit{The reusable degraded-observation kernel} \\
\bottomrule
\end{tabular}
\end{table}

% ----------------------------------------------------------------
\section{Constraint-Class Mismatch: When 1-D Action Repair Is Insufficient}
\label{sec:constraint-class-mismatch}
% ----------------------------------------------------------------

The DC3S shield projects each candidate action onto a tightened interval
derived from the conformal uncertainty set.  This is sufficient whenever the
domain safety constraint is a box---or more generally, whenever it decomposes
as a monotone function of a single action coordinate.  The following
proposition shows that velocity-dependent constraints, such as the headway
constraint in the autonomous-vehicle domain, require a strictly larger repair
surface.

\begin{proposition}[Constraint-class mismatch barrier]
\label{prop:constraint-class-mismatch}
Let the true safety constraint be the velocity-dependent headway condition
\[
  \phi(x_t)
  \;=\;
  \bigl[\,\mathrm{gap}_t \;\ge\; d_{\min} + \tau\,v_t\,\bigr],
  \quad d_{\min},\tau > 0,
\]
where $x_t = (\mathrm{gap}_t, v_t)$ and the gap obeys
$\mathrm{gap}_{t+1} = \mathrm{gap}_t + \Delta t\,(v_{\mathrm{lead}}
- v_t - a_t\,\Delta t)$.  Suppose the DC3S shield projects the
candidate acceleration into a one-dimensional safe set
\[
  \mathcal{A}_t
  \;=\;
  \bigl[\,a_{\min},\;
    (b_t - v_t)/\Delta t\,\bigr],
  \quad b_t = \mathrm{gap}_t\,/\,(\Delta t + \tau) \in \mathbb{R},
\]
and suppose that there exist states $(\mathrm{gap}_t, v_t)$ with
$v_t > (\mathrm{gap}_t - d_{\min})\,/\,\tau$---that is, $x_t\notin\mathcal{S}$
already at the start of the step.
Then:
\begin{enumerate}
  \item No action $a_t \in [a_{\min}, a_{\max}]$ moves $x_{t+1}$ into
    $\mathcal{S}$ in one step, because the constraint violation is
    determined by \emph{current} $(v_t, \mathrm{gap}_t)$, not by the
    next-step speed the shield controls.
  \item The shield's one-dimensional repair is therefore infeasible for
    current-step constraint satisfaction: $\phi(x_t)$ depends on the
    pre-action state, not on the post-action state.
  \item DC3S's T3 upper bound $\mathbb{E}[V_T]\le\alpha(1-\bar{w})T$
    remains valid for the \emph{next-step} constraint (the set the shield
    can actually control), but it does not apply to the \emph{current-step}
    plant-level constraint that the AV evaluation harness measures.
\end{enumerate}
Consequently, the AV domain incurs a structural proof gap: the shield repairs
future-step feasibility, but the plant reports current-step violations.
\end{proposition}

\begin{proof}
If $v_t > (\mathrm{gap}_t - d_{\min})\,/\,\tau$, then
$\mathrm{gap}_t < d_{\min} + \tau v_t$, so $\phi(x_t) = \mathrm{False}$
regardless of $a_t$.  The action $a_t$ affects $v_{t+1} = v_t + a_t\,\Delta t$
and $\mathrm{gap}_{t+1}$, but not the evaluation $\phi(x_t)$.
The shield's admissible set $\mathcal{A}_t$ therefore provides no mechanism
to avoid the step-$t$ violation; it only constrains step-$(t+1)$.
\end{proof}

\begin{corollary}[AV promotion routes from the original headway mismatch]
\label{cor:av-promotion}
Discharging the original autonomous-vehicle headway mismatch requires at
least one of the following:
\begin{enumerate}[label=(\alph*)]
  \item \emph{Augmented state repair}: the shield solves a two-dimensional
    quadratic program over $(a, v)$ that explicitly enforces
    $\mathrm{gap}_t \ge d_{\min} + \tau v_t$ as a current-state feasibility
    constraint, activating a full-brake fallback when the state is already
    outside~$\mathcal{S}$.
  \item \emph{TTC reformulation}: replace the headway constraint with a
    time-to-collision lower bound, $\mathrm{TTC}_t \ge t_{\min}$, whose
    gradient with respect to $a_t$ is non-zero and admits a valid
    one-dimensional repair.
  \item \emph{Predictive entry barrier}: add a one-step lookahead check
    that refuses to enter states from which $\phi(x_{t+1}) = \mathrm{False}$
    is unavoidable, reducing the observed TSVR to the T3 envelope.
\end{enumerate}
Without one of these routes, the AV residual is a \emph{structural} residual,
not a conformal-coverage residual.
\end{corollary}

\paragraph{Current closure route.}
This dissertation adopts the TTC reformulation together with the predictive
entry barrier as the canonical AV promotion path.  The two-dimensional
$(a,v)$ repair remains a valid future route, but it is not the route claimed
here.  Under the bounded TTC contract, the closure matrix now records the AV
row as proof-validated: locked replay reduces TSVR to zero on the canonical
validation harness, the conservative entry-barrier soundness cases clear, and
the CertOS lifecycle executes on the same adapter surface.  This promotion is
deliberately narrow.  It does not promote every richer headway, lane-change,
or multi-agent vehicle setting to equal proof depth.

% ----------------------------------------------------------------
\section{Chapter Summary}
\label{sec:ch37-summary}
% ----------------------------------------------------------------

This chapter closes the monograph at the correct control-theory level.

\begin{enumerate}
  \item \textbf{T9} frames observation-aware safety logic as structurally
    necessary under persistent degradation.
  \item \textbf{T10} turns mean reliability into a quantitative
    infrastructure-to-risk frontier.
  \item \textbf{T11} packages universality as a typed kernel and adapter
    template, not as a claim that every domain already has battery-level proof
    closure.
  \item \textbf{The unification theorem} explains how prior safe-control
    families sit inside the same degraded-observation framework.
\end{enumerate}

The central theorem contribution is now precise: the safe-control object under
degraded telemetry is not the nominal state alone but a reliability-indexed
observation-consistent state set together with a repair map and a causal
certificate surface.  Battery supplies the full proof witness.  T9 and T10
show that degraded observation creates a necessity and a lower bound.  T11
shows how the same kernel transfers when a new domain discharges the explicit
structural obligations.  This is a defensible control-theory claim; it is not
an assertion that all evaluated domains already enjoy equal proof closure.
