\section{Cross-Domain Synthesis}
\label{sec:prof-cross-domain}

ORIUS is best read as a universal runtime safety layer for Physical AI, not as a
single-domain controller that happens to generalize.  The common object across the
monograph is the degraded-observation hazard: the controller sees an observation
surface, but safety is judged on the true state or the observation-consistent
state set.  That is why the same runtime grammar recurs across battery storage,
autonomous vehicles, industrial process control, healthcare monitoring, navigation,
and aerospace.

\begin{table*}[t]
\centering
\caption{Current cross-domain status in the tracked ORIUS parity surface.}
\label{tab:prof-cross-domain-status}
\small
\begin{tabularx}{\textwidth}{@{}p{0.18\textwidth}p{0.16\textwidth}p{0.28\textwidth}p{0.28\textwidth}@{}}
\toprule
\textbf{Domain} & \textbf{Tier} & \textbf{What the repo currently shows} & \textbf{Why it matters for the universal claim}\\
\midrule
Battery energy storage & reference & Deepest theorem-to-code-to-artifact witness; tracked as the calibration row, not the conceptual center. & Establishes the strongest witness surface for the runtime grammar.\\
Autonomous vehicles & proof\_validated & Locked replay, typed kernel checks, and bounded TTC plus predictive-entry-barrier semantics clear the current gate. & Demonstrates that the universal contract travels beyond energy while remaining bounded to the present vehicle contract.\\
Industrial process control & proof\_validated & Real telemetry, replay closure, and material post-repair improvement clear the defended row. & Shows the runtime grammar on a non-energy plant with strong empirical closure.\\
Healthcare monitoring & proof\_validated & Locked ICU-vital telemetry and bounded intervention semantics clear the current gate. & Shows the contract can govern monitoring and intervention without actuation-heavy assumptions.\\
Navigation and guidance & shadow\_synthetic & Adapter and replay harness are present, but the real-data closure row is still blocked. & Supports portability, not parity; it should remain explicitly qualified.\\
Aerospace control & experimental & Placeholder or supplemental flight surfaces exist, but the defended parity gate is not closed. & Marks the outer boundary of the current program and must not be presented as defended closure.\\
\bottomrule
\end{tabularx}
\end{table*}

The synthesis that follows from this status is narrower and stronger than a generic
transfer-learning claim.  ORIUS standardizes the runtime objects that matter for
safety under degraded observation:
reliability assessment, uncertainty widening, admissible-set tightening, repair or
fallback, certificate emission, and auditable lifecycle logging.  What varies across
domains is the safety geometry, the telemetry parser, and the domain-specific repair
policy.  What stays invariant is the safety-layer grammar.

That distinction is the key cross-domain result.  Battery remains the witness row, but
it does not monopolize the theory.  Industrial and healthcare are defended bounded
rows with real replay evidence.  Autonomous vehicles are defended under the current
bounded TTC contract.  Navigation and aerospace are still informative because they
show where the parity gate has not yet been earned.

The professor-facing takeaway is therefore precise: ORIUS already supports a
universal safety architecture, but the evidence ladder is still tiered.  The
architecture is universal in contract; the closure depth is not yet universal in
proof.  That is exactly the right level of honesty for an R1-grade monograph.
