\section{Related Work and the Insufficiency of Prior Families}
\label{sec:prof-related-work}

ORIUS does not claim that prior work ignored uncertainty or safety. The claim is
narrower and stronger: the major adjacent families solve pieces of the problem, but not
the full runtime bridge from degraded observation to repaired actuation and auditable
release.

\subsection{Conformal prediction and calibration are necessary, but not enough}
Distribution-free and conformal methods provide finite-sample uncertainty surfaces for
learned predictors \cite{vovk2005algorithmic,lei2018distribution,romano2019conformalized,gibbs2021adaptive,zaffran2022adaptive,angelopoulos2023gentle,barber2023beyond}.
These tools are indispensable for ORIUS because the runtime must widen the
observation-consistent set when reliability drops. But they do not, by themselves,
specify which action remains legal after that widening. A prediction interval is not a
certificate of actuation legality. ORIUS therefore treats calibration as the input to a
safety layer rather than as the safety layer itself.

\subsection{Runtime assurance and Simplex-style supervision do not close the observation gap}
Runtime assurance, supervisory control, and Simplex-style architectures
\cite{sha1998simplex,sha2001using,desai2019soter,bloem2015shield,bartocci2018introduction,leucker2009brief,havelund2004runtime}
show that nominal intelligence should not be the final authority over actuation. ORIUS
inherits that insight directly. The gap is that many supervisory designs begin after a
candidate action already exists and assume that the state estimate is coherent enough to
be monitored. ORIUS starts earlier: it treats the observation surface itself as unsafe,
which forces the runtime to reason jointly about reliability, uncertainty inflation,
repair geometry, fallback, and certificate validity.

\subsection{Safety filters and constrained control are mathematically strong but plant-local}
Barrier functions, robust MPC, tube MPC, predictive safety filters, and reachability-
based controllers can turn model uncertainty into admissible action sets
\cite{ames2014cbf,ames2019control,wabersich2021predictive,fisac2019general,mayne2005robust,langson2004tubes,camacho2013model,rawlings2017mpc}.
ORIUS does not compete with that mathematics. It packages those ideas behind a typed
runtime contract that can host multiple geometries across multiple plants. The
scientific move is not a new universal controller; it is one reusable way to make
degraded observation alter actuation semantics.

\subsection{Anomaly, drift, and fault detection stop too early}
Change detection, drift detection, anomaly scoring, and CPS fault analysis provide the
language for recognizing when telemetry should no longer be trusted
\cite{basseville1993detection,page1954continuous,page1955test,hinkley1971inference,breunig2000lof,liu2008isolation,pasqualetti2013attack,teixeira2015secure}.
Yet many systems stop at an alarm or a reliability score. ORIUS treats that as an
incomplete runtime response. Once trust has fallen, the system still needs to decide
which actions remain legal, which actions must be repaired, when fallback is required,
and how that decision is certified.

\subsection{Domain-specific safety work rarely becomes a universal runtime grammar}
Energy dispatch, autonomous driving, clinical monitoring, and other physical-AI safety
settings each contain strong plant-level ideas, but most remain
tied to one domain's state geometry, one benchmark, and one intervention semantics. The
result is local excellence without a shared runtime object. ORIUS positions itself as a
cross-domain safety-layer grammar that can sit above those domain-specific methods while
keeping the data and proof posture explicitly tiered.

\begin{table*}[t]
\centering
\small
\caption{Why prior families remain necessary but insufficient for the ORIUS object.}
\label{tab:prof-family-gap}
\begin{tabularx}{\textwidth}{@{}p{0.17\textwidth}p{0.25\textwidth}p{0.50\textwidth}@{}}
\toprule
\textbf{Method family} & \textbf{What it contributes} & \textbf{Why it is insufficient for OASG} \\
\midrule
Conformal and adaptive calibration & Distribution-free uncertainty and shift-aware coverage control & Does not decide repaired actuation, fallback legality, or certificate release after trust degrades. \\
Runtime assurance and shielding & Supervisory interception and safety-controller fallback & Often assumes the observation surface is coherent enough to supervise directly. \\
Safety filters and robust constrained control & Strong admissible-set mathematics under uncertainty & Usually remains plant-local and does not define a reusable benchmark/governance contract. \\
Anomaly, drift, and fault detection & Detects that telemetry has become unreliable & Stops at alarms or scores unless paired with repair, fallback, and audit semantics. \\
Domain-specific CPS safety stacks & High-quality local solutions in one plant family & Rarely expose a universal runtime grammar that transfers across domains without changing the scientific object. \\
\bottomrule
\end{tabularx}
\end{table*}

The ORIUS thesis is therefore integrative but not vague. The universal object is not
``safe AI'' in the abstract. It is the typed runtime layer that binds reliability,
uncertainty, admissible actions, repaired release, and certificate governance under one
action-semantics problem.
