\section{Runtime, Latency, and Governance Evidence}
\label{app:prof-runtime-evidence}

This appendix section collects the evidence surfaces that justify the
runtime contract in the main IEEE draft.  The purpose is not to flood the
reader with every locked artifact in the repo.  It is to show that the
adapter, benchmark, and CertOS claims are backed by measured runtime
latency, explicit replay tables, and a visible governance gate.

\subsection{Latency and Footprint}
The latency story is straightforward.  The runtime chapter
(\texttt{chapters/ch22\_latency\_systems\_footprint.tex}) measures the DC3S
pipeline with \texttt{time.perf\_counter\_ns()} and reports sub-millisecond
step time on the locked battery dispatch surface.  The key component
latencies are the reliability scorer, the drift updater, the uncertainty
set builder, and the action repair stage.

\begin{table}[t]
\centering
\caption{Per-component DC3S latency summary from the locked runtime study.}
\label{tab:prof-latency-summary}
\begin{tabular}{lrrrrr}
\toprule
\textbf{Component} & \textbf{Mean} & \textbf{Median} & \textbf{P95} & \textbf{P99} & \textbf{Max} \\
\midrule
Reliability scoring (OQE)   & 0.026 & 0.021 & 0.043 & 0.130 & 0.712 \\
Drift update (Page-Hinkley) & 0.000 & 0.000 & 0.001 & 0.001 & 0.160 \\
Uncertainty set build (RUI) & 0.012 & 0.011 & 0.011 & 0.047 & 0.724 \\
Action repair (SAF/Shield)  & 0.002 & 0.002 & 0.002 & 0.003 & 0.210 \\
\midrule
\textbf{Full DC3S step}     & \textbf{0.042} & \textbf{0.037} & \textbf{0.041} & \textbf{0.193} & \textbf{2.416} \\
\bottomrule
\end{tabular}
\end{table}

The same section also provides the best compact footprint statement for a
professor-facing draft: the full step fits comfortably inside a 30-second
dispatch cycle, and the runtime memory proxy remains dominated by the
forecast and residual buffers rather than by certificate bookkeeping.
That makes the runtime layer operationally cheap enough to be believable
as a Physical AI safety wrapper.

\subsection{Runtime Budget Across Domains}
The cross-domain runtime-budget matrix in
\texttt{reports/publication/tbl\_orius\_runtime\_budget\_matrix.tex} is the
best evidence that the runtime contract is not battery-specific.  It shows
that the six-domain program preserves sub-millisecond P95 latency while
keeping certificate rates at 100 percent on all rows.

\begin{table*}[t]
\centering
\scriptsize
\caption{Cross-domain runtime budget from the locked publication matrix.}
\label{tab:prof-runtime-budget}
\begin{tabular}{p{2.1cm}p{1.6cm}p{1.7cm}p{1.5cm}p{1.6cm}p{1.5cm}p{1.7cm}p{1.2cm}p{1.8cm}p{1.7cm}}
\toprule
\textbf{Domain} & \textbf{P95 latency} & \textbf{Reliability} & \textbf{Repair rate} & \textbf{Cert rate} & \textbf{Error rate} & \textbf{Fallback mode} & \textbf{CertOS} & \textbf{Runtime depth} & \textbf{Limit} \\
\midrule
Battery Energy Storage & 0.1193 ms & witness\_reference & n/a & 100.0\% & 0.0\% & safe\_hold & evaluated & reference\_runtime & bounded deployment items remain \\
Autonomous Vehicles & 0.185 ms & 0.783 & 100.0\% & 100.0\% & 0.0\% & full\_brake & evaluated & proof\_validated & promoted \\
Industrial Process Control & 0.152 ms & 0.655 & 91.7\% & 100.0\% & 0.0\% & power\_cap & evaluated & proof\_validated & chain complete \\
Medical and Healthcare Monitoring & 0.137 ms & 0.817 & 16.7\% & 100.0\% & 0.0\% & max\_alert & evaluated & proof\_validated & chain complete \\
Navigation and Guidance & 0.145 ms & 0.749 & 100.0\% & 100.0\% & 0.0\% & hold\_position & gated & shadow\_synthetic & real data gap \\
Aerospace Control & 0.168 ms & 0.706 & 18.8\% & 100.0\% & 0.0\% & envelope\_hold & gated & experimental & placeholder \\
\bottomrule
\end{tabular}
\end{table*}

This table should stay close to the runtime section in the professor draft
because it directly supports the claim that the universal runtime layer is
lightweight across domains while still respecting the evidence tier.

\subsection{Governance Evidence}
The governance story is best captured by the locked lifecycle matrix in
\texttt{reports/publication/tbl\_orius\_governance\_lifecycle\_matrix.tex} and
the domain-closure matrix in \texttt{reports/publication/orius\_domain\_closure\_matrix.csv}.
Together they show that CertOS is not a decorative abstraction: the same
issue/validate/expire/fallback logic is observed in the defended rows, while
navigation and aerospace remain gated by explicit blockers.

\begin{table*}[t]
\centering
\scriptsize
\caption{Cross-domain CertOS lifecycle evidence from the locked publication matrix.}
\label{tab:prof-governance-lifecycle}
\begin{tabular}{p{2.1cm}p{2.2cm}p{2.0cm}p{1.5cm}p{2.0cm}p{2.1cm}p{1.2cm}p{1.6cm}}
\toprule
\textbf{Domain} & \textbf{Lifecycle surface} & \textbf{Hash chain} & \textbf{CertOS} & \textbf{Shared constraint} & \textbf{Audit continuity} & \textbf{Complete} & \textbf{Limit} \\
\midrule
Battery Energy Storage & issue/validate/expire/fallback all seen & ok / invariants pass & evaluated & shared feeder capacity & tracked runtime audit & 97.0\% & bounded to current task surface \\
Autonomous Vehicles & issue/validate/expire/fallback all seen & ok / invariants pass & evaluated & gated & tracked runtime audit & 93.0\% & promoted \\
Industrial Process Control & issue/validate/expire/fallback all seen & ok / invariants pass & evaluated & shared power budget & tracked runtime audit & 94.0\% & bounded to current task surface \\
Medical and Healthcare Monitoring & issue/validate/expire/fallback all seen & ok / invariants pass & evaluated & gated & tracked runtime audit & 93.0\% & chain complete \\
Navigation and Guidance & gated or not yet supported & gated & gated & gated & bounded or gated & 58.0\% & real data gap \\
Aerospace Control & gated or not yet supported & gated & gated & gated & bounded or gated & 55.0\% & placeholder \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table}[t]
\centering
\caption{CertOS runtime summary from the battery lifecycle demo.}
\label{tab:prof-certos-summary}
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total steps & 96 \\
Valid steps & 47 \\
Degraded steps & 16 \\
Fallback steps & 33 \\
Runtime interventions & 33 \\
Audit entries & 96 \\
Audit completeness & 100.0\% \\
Step latency & 0.59 ms \\
\bottomrule
\end{tabular}
\end{table}

The appendix should end with the governance boundary spelled out plainly:
the runtime layer is fully visible, the certificates are auditable, and the
fallback path is explicit, but the claim remains tiered by domain.  Battery
is the reference witness; industrial and healthcare are defended bounded
rows; navigation and aerospace remain gated.  That is enough for a strong
IEEE draft, and it is more defensible than flattening the evidence into one
universal-number story.
