\section{Universal Contracts: Adapter, Benchmark, and CertOS Governance}
\label{sec:prof-runtime-contracts}

ORIUS is organized around three contracts: the adapter contract, the
benchmark contract, and the CertOS governance contract. Together they
explain how degraded observation is converted into a typed runtime safety
layer rather than a battery-only control recipe.

\subsection{Adapter contract}
The adapter is the only place where domain-specific telemetry and
constraints are allowed to enter the universal runtime. The repository and
its supporting design documents make this boundary explicit:
\texttt{src/orius/certos/runtime.py} provides a domain-neutral governance
surface, \texttt{docs/UNIVERSAL\_BENCHMARK\_SPEC.md} defines the common
step schema, and \texttt{docs/UNIVERSAL\_GOVERNANCE\_SPEC.md} requires a
domain governance policy instead of battery-shaped runtime state.

In practical terms, each domain adapter must expose a true-state violation
predicate, an observed-state satisfiability predicate, an optional margin
computation, a fallback action, and certificate-compatible action repair.
That is the minimum needed to make ORIUS portable across battery energy
storage, autonomous vehicles, healthcare monitoring, and future additional
domains without changing the meaning of safety at the kernel level.

The adapter can therefore be read as a typed semantic projection
\[
  \mathfrak{D}_t:
  (o_t,\mathcal{H}_t,a_t^{\mathrm{cand}})
  \mapsto
  (w_t,U_t,\mathcal{A}^{\mathrm{safe}}_t,\mathcal{R}_t,\mathcal{F}_t,c_t),
\]
where telemetry and nominal actions are translated into the universal runtime objects
that ORIUS knows how to govern. This is one of the most important clarifications in the
paper: universality does not mean a universal plant model. It means one typed safety
projection from domain observation to governed actuation.

\subsection{Benchmark contract}
ORIUS-Bench is the manuscript-facing replay contract. The universal step
record keeps the same fields across domains:
\texttt{true\_constraint\_violated},
\texttt{observed\_constraint\_satisfied},
\texttt{true\_margin},
\texttt{observed\_margin},
\texttt{intervened},
\texttt{fallback\_used},
\texttt{certificate\_valid},
\texttt{latency\_us}, and
\texttt{domain\_metrics}. The metrics family is equally fixed: TSVR, OASG,
certificate validity accuracy, intervention rate, audit completeness,
latency percentiles, and coverage diagnostics when conformal uncertainty is
active.

The benchmark contract matters because it decides what evidence counts. A step is not
judged by whether the nominal controller felt confident. It is judged by whether the
true constraint was violated, whether the observed state looked safe, whether ORIUS
intervened, whether fallback was used, whether the certificate remained valid, and what
latency budget was consumed. This prevents the evaluation from collapsing into
prediction accuracy or task reward while still remaining agnostic to the nominal model.

This contract prevents the argument from collapsing into a single battery
result table. It also keeps the evidence boundary honest: the same schema
can be reused across the defended three-domain lane and future additional
domains, but the rows do not carry the same proof depth.

\subsection{CertOS governance contract}
CertOS is the runtime governance layer.  Its role is deliberately narrow:
do not release an action unless the runtime can either validate a current
certificate or replace the action with an explicit fallback.  The policy
surface is domain-neutral and consists of five hooks:
\texttt{is\_actuation(action)},
\texttt{fallback\_action(constraints, state)},
\texttt{horizon\_update(certificate, reliability, drift, constraints, validity\_horizon)},
\texttt{required\_certificate\_fields()}, and
\texttt{integrity\_projection(action)}.

The invariants are correspondingly small and auditable: no actuation
without a valid or fallback certificate path, an unbroken certificate hash
chain, and fallback only when validity collapses. That is the right level
of abstraction for a professor-facing draft that wants to argue for a
fundamental safety layer.

This is also where ORIUS separates itself from many safety wrappers that are difficult
to compare scientifically. If a runtime layer cannot explain what is released, under
what certificate, for how long that certificate remains valid, and how fallback is
recorded, then it is hard to distinguish principled safety governance from an
implementation-specific patch. CertOS is deliberately narrow so that those questions are
answerable.

\begin{algorithm}[t]
\caption{CertOS Certificate Release and Lifecycle Update}
\label{alg:prof-certos}
\begin{algorithmic}
\Require repaired action \(a_t^{\mathrm{safe}}\), certificate candidate \(c_t\), reliability \(w_t\), constraints \(\kappa_t\)
\State project \(a_t^{\mathrm{safe}}\) through \texttt{integrity\_projection}
\If{\texttt{required\_certificate\_fields()} missing}
  \State mark certificate invalid and request fallback
\EndIf
\State validate lifecycle state and audit hash continuity
\If{certificate valid}
  \State compute horizon update using \texttt{horizon\_update}
  \State append validated certificate to audit ledger
  \State release projected action
\Else
  \State replace with \texttt{fallback\_action} and append degraded certificate
  \State release fallback only if lifecycle invariants still hold
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{table}[t]
\centering
\small
\caption{The three public ORIUS contracts used throughout the professor draft.}
\label{tab:prof-contract-stack}
\begin{tabularx}{\columnwidth}{@{}p{0.19\columnwidth}X@{}}
\toprule
\textbf{Layer} & \textbf{Contract obligation} \\
\midrule
Adapter & Convert degraded telemetry into reliability, uncertainty, safe-action, repair, fallback, and certificate objects. \\
Benchmark & Keep one true-state safety schema and one metric family across all domains. \\
CertOS & Release only validated or explicit fallback actions with audit continuity. \\
\bottomrule
\end{tabularx}
\end{table}

The claim boundary is therefore straightforward: ORIUS is a universal
runtime safety contract for Physical AI under degraded observation, but the
evidence remains tiered by domain and is not presented as equal-depth proof
across all rows.

That methodological discipline matters as much as the API itself. It prevents the paper
from reading like a battery controller with extra documentation. By freezing the
adapter boundary, the benchmark record, and the governance hooks, ORIUS defines one
reproducible scientific object whose meaning does not change from one domain chapter to
the next.
