\section{Battery Witness Domain}
\label{sec:prof-battery-results}

The battery row is the strongest witness in ORIUS because it closes the entire chain
from assumptions and theorems to controller code, replay artifacts, latency budgets,
and certificate traces. It is therefore the right place to show the runtime grammar in
its deepest form without allowing the paper to collapse back into a battery-only
manuscript.

\subsection{Problem, state, action, and safety object}
The witness task is storage dispatch under uncertain load, renewable injection, price
signals, and telemetry faults. The true state includes at least state of charge,
operational bounds, and dispatch context; the candidate action is charge/discharge
power. The governing safety object is not cost. It is hidden state violation under
degraded observation, measured on the true state rather than on controller-visible
legality.

\subsection{Data and replay surface}
The battery evidence is built on the locked CPSBench-derived replay surface with
168-hour episodes, fixed seeds, and fault families that include dropout, stale sensing,
delay, drift, and mixed degradation. This is what makes the battery row useful as a
witness: it is a repeatable degraded-observation environment rather than a one-off case
study.

\subsection{Witness result}
The decisive battery result is that ORIUS repair-aware controllers eliminate true-state
violations under the locked protocol while non-repair baselines do not. The deep OQE
variant is not presented as a miracle predictor; it is a quality signal routed into a
repair-and-certify layer that prevents hidden violations.

\input{generated/orius_professor_battery_witness_table.tex}

The deeper trace, ablation, and raw-sequence evidence remains important, but it belongs
in Appendix~B. In the main paper the battery row serves one purpose: it proves that the
runtime safety layer is not merely architectural. It changes the released-action
semantics and removes true-state failures on the strongest witness surface currently
available in the repo.
