\section{Introduction and Contribution Statement}
\label{sec:prof-introduction}

Physical AI systems do not act on truth. They act on telemetry, inferred state, delayed
measurements, missing packets, stale buffers, and learned surrogates of the world. The
resulting engineering problem is deeper than forecast error or planner suboptimality:
the software stack can remain internally coherent while the physical system becomes
externally unsafe. Batteries may violate hidden state-of-charge limits, vehicles may
enter unsafe headway regimes on stale perception, and clinical monitoring pipelines
may suppress timely intervention.

ORIUS is built around the claim that these failures share one universal object:
degraded observation changes the semantics of what may still be safely done. The field
therefore needs a runtime layer whose job is not to outperform every planner, but to
mediate the boundary between degraded observation and physical actuation. ORIUS,
Observation--Reality Integrity for Universal Safety, is proposed as that layer. It
detects reliability loss, calibrates the observation-consistent state set, constrains
the admissible action set, repairs or replaces unsafe candidate actions, and certifies
release through an auditable lifecycle. The framework is written around the universal
runtime claim rather than a witness-row-specific framing. Battery is retained as the
deepest witness row, not as the conceptual center of the claim.

This manuscript positions ORIUS against a precise gap in the literature. Distribution-
free inference can quantify uncertainty
\cite{vovk2005algorithmic,lei2018distribution,romano2019conformalized,gibbs2021adaptive,angelopoulos2023gentle,barber2023beyond}.
Runtime assurance and shields can intercept unsafe behavior
\cite{sha1998simplex,sha2001using,desai2019soter,bloem2015shield,bartocci2018introduction,leucker2009brief}.
Safety filters and constrained control can map uncertainty into admissible action sets
\cite{ames2019control,ames2014cbf,wabersich2021predictive,fisac2019general,mayne2005robust,rawlings2017mpc}.
Fault, anomaly, and drift detection can identify when observation should no longer be
trusted \cite{basseville1993detection,page1954continuous,page1955test,hinkley1971inference,pasqualetti2013attack,teixeira2015secure}.
What none of these families alone provides is a reusable runtime grammar that links
degraded observation to repaired actuation and auditable certificate release across
multiple physical domains.

The paper therefore makes four contributions.
\begin{enumerate}[leftmargin=1.25em,itemsep=0.2em]
  \item It defines the observation--action safety gap (OASG) as the right universal
  hazard object for Physical AI under degraded observation.
  \item It presents the Detect--Calibrate--Constrain--Shield--Certify kernel as a typed
  runtime safety layer with explicit formulas, pseudocode, and certificate semantics.
  \item It unifies domain evaluation through one adapter contract, one benchmark schema,
  and one CertOS governance surface.
  \item It argues for a bounded-universal evidence posture: battery is the witness row;
  AV and healthcare are defended bounded rows.
\end{enumerate}

\subsection{What ORIUS is and is not}
ORIUS is not a new nominal controller, a new world model, or a universal planner.
It does not ask every domain to share the same action space, dynamics model, or
learning stack. Instead it introduces a narrow but rigorous intervention layer at the
moment where candidate actuation is about to be released. That distinction matters.
Many physical AI projects already contain estimators, predictors, planners, and local
guards. What they usually lack is a reusable semantics for what happens when the
controller-visible state is no longer trustworthy. ORIUS supplies that semantics.

Equally important, ORIUS is not an excuse to hide ad hoc emergency logic behind a new
name. Every intervention mode in the framework must be typed and auditable: keep,
repair, fallback, degrade, or reject. The release decision is therefore promoted from
implementation detail to scientific object. That is the conceptual step that lets the
paper talk about batteries, vehicles, and healthcare monitoring within one formal vocabulary.

\subsection{Design principles}
The framework is organized around four design principles.
\begin{enumerate}[leftmargin=1.25em,itemsep=0.2em]
  \item \textbf{Post-nominality:} upstream intelligence remains replaceable; ORIUS is
  evaluated at the release boundary rather than by forecasting or planning quality.
  \item \textbf{Typed degradation:} reliability loss, widened uncertainty, repaired
  action, fallback mode, and certificate status are explicit runtime objects.
  \item \textbf{Domain-bounded portability:} each adapter owns telemetry semantics and
  safety geometry, but the kernel and benchmark semantics do not change from one row to
  the next.
  \item \textbf{Auditable release:} safety is judged on what actuation is actually
  released, under what certificate, and with what lifecycle trace.
\end{enumerate}

These principles explain why ORIUS should be read as infrastructure for Physical AI
safety rather than as one more domain benchmark result. A battery controller can be
good or bad; an AV planner can be good or bad; a healthcare monitor can be good or
bad. ORIUS asks a different question: when those nominal systems become observation-
degraded, what runtime object decides whether the next action remains legally releasable?

The draft is intentionally ambitious but disciplined. It uses a universal writing
posture, but it keeps one explicit evidence-boundary section in the main body. It
includes formulas, theorem statements, proof sketches, runtime pseudocode, decisive
results tables, and cross-domain synthesis, while pushing proof-depth and artifact-depth
material into two separate appendices.
