\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This thesis presents ORIUS (Observation--Reality Integrity for Universal Safety), a
universal safety layer for physical AI under degraded observation. The central claim is
that a controller can appear safe on the observed state while becoming physically unsafe
on the true state when the observation channel degrades. ORIUS treats that
observation--action safety gap (OASG) as a first-class runtime object and addresses it
through one typed Detect--Calibrate--Constrain--Shield--Certify kernel: detect degraded
telemetry, widen the observation-consistent state set, tighten the admissible action
set, repair or replace the candidate action, and emit a causal certificate that can be
audited after the fact.

The book is organized as a universal-first monograph rather than as a battery-centered
thesis with later extensions. It develops one common architecture, one benchmark and
runtime-governance discipline, one theorem bridge, and six first-class domain chapters:
battery energy storage, autonomous vehicles, industrial process control, medical and
healthcare monitoring, navigation and guidance, and aerospace control. Each domain is
forced through the same safety-layer template: domain problem, degraded-observation
hazard, ORIUS instantiation, safety object, dataset and training surface, replay and
evidence surface, fallback behavior, and explicit non-claims.

The formal contribution is a domain-agnostic safe-control kernel under degraded
observation. The thesis proves that hidden true-state violations can arise even when
observed-state legality appears intact, that reliability-aware action repair is
structurally necessary, and that the ORIUS repair-and-certify stack admits explicit
quantitative safety bounds and transfer obligations under stated assumptions. The
current theorem-to-code-to-artifact chain remains deepest on the battery reference row,
but the monograph packages the broader result as one universal architectural and
semantic claim rather than as a collection of disconnected domain papers.

The empirical contribution is governed by one equal-domain parity gate. Battery remains
the deepest reference row. Industrial process control and healthcare monitoring are
defended proof-validated rows under bounded contracts. The current autonomous-vehicle
row clears bounded proof-validation under a TTC plus predictive-entry-barrier contract.
Navigation remains gated by the absence of a defended real-data row, and aerospace
remains gated by an experimental placeholder flight-task surface. This explicit parity
matrix prevents the universal-first narrative from outrunning the evidence.

The manuscript is intentionally conservative about where theory is stronger than
evidence. The runtime contract and theorem bridge are broader than the current
cross-domain empirical closure, so claims outside the defended rows are written as
bounded architectural or semantic statements rather than as equal-domain validation
claims. The reviewer-analysis appendix is included only as an editorial audit and
revision-traceability aid; it is not part of the scientific evidence surface.

Across the defended rows, ORIUS shows that degraded observation can be turned into a
runtime safety problem with explicit repair, fallback, and certificate semantics rather
than treated as a vague robustness concern. The broader contribution of the thesis is
therefore not a claim that every physical-AI domain is already equally mature. It is
the stronger and more durable claim that one universal safety-layer grammar can govern
all of them, while a visible parity gate determines how far each domain may currently be
promoted.
