%% Chapter 21 — Group-Conditional Coverage and High-Probability Violation Bound
%% Phase 2 of ORIUS gap-closing plan
%% Adds Theorem 9 (Mondrian conditional coverage) and Theorem 10 (Hoeffding bound)

\section{Group-Conditional Coverage and High-Probability Bounds}
\label{sec:ch21_conditional_coverage}

\subsection{Motivation}

The core bound of Chapter~18 guarantees a marginal expectation:
$\mathbb{E}[V] \leq \alpha(1-\bar{w})T$.
Two gaps remain that the PhD committee identified as critical:
\begin{enumerate}
  \item \textbf{Marginal only}: the bound averages over all reliability groups
        and does not guarantee that \emph{each subgroup} of timesteps is well
        covered.  A system that is perfectly safe for high-reliability windows
        but catastrophically unsafe for low-reliability windows could still
        satisfy the marginal bound.
  \item \textbf{Expectation only}: the bound says the \emph{average} violation
        rate is low, but gives no probabilistic guarantee that a \emph{single
        realisation} stays close to that average.
\end{enumerate}

This chapter closes both gaps.

\subsection{Theorem 9 — Group-Conditional Coverage Under Reliability Partitioning}

\begin{theorem}[Mondrian Group-Conditional Coverage]\label{thm:mondrian_conditional}
Partition the reliability scores $w_t \in [0,1]$ into $K$ disjoint bins
$\mathcal{G}_1, \ldots, \mathcal{G}_K$ (e.g.\ equal-mass quantile bins of low,
medium, and high reliability).  For each group $k$, let
$\hat{C}_k(x) = [\hat{y}(x) - q_k \cdot \mathrm{infl}_k,\;
                  \hat{y}(x) + q_k \cdot \mathrm{infl}_k]$
be the DC3S conformal set calibrated on the subset
$\{i : w_i \in \mathcal{G}_k\}$.
Under the exchangeability assumption of the group-specific calibration set:
\[
  P\!\left(Y \in \hat{C}_k(X)\;\middle|\; w_t \in \mathcal{G}_k\right) \geq 1 - \alpha
  \qquad \forall\, k = 1,\ldots,K.
\]
\end{theorem}

\begin{proof}[Proof sketch]
Within group $k$, the calibration pairs $\{(x_i, y_i) : w_i \in \mathcal{G}_k\}$
form an exchangeable sequence (by the same exchangeability assumption used for
marginal conformal prediction, restricted to the subgroup).
By the standard split-conformal guarantee applied to this subgroup,
the quantile $q_k$ satisfies
$P(Y \in \hat{C}_k(X) \mid w_t \in \mathcal{G}_k) \geq 1 - \alpha$.
Since infl$_k \geq 1$ (Assumption~3 of the core bound), the inflated set
$\hat{C}_k$ is a superset of the base conformal set, so coverage is
preserved by monotonicity.  $\square$
\end{proof}

\paragraph{Empirical verification.}
The function \texttt{mondrian\_group\_coverage()} in
\texttt{src/orius/dc3s/coverage\_theorem.py} bins the test trace into $K$
reliability groups and reports per-group PICP.  The evidence gate requires
PICP $\geq 1 - \alpha - 0.02$ for every group.  The defended Battery, AV, and
Healthcare rows pass this gate (see \Cref{tab:group_conditional_coverage}).

\input{assets/tables/generated/tbl_group_conditional_coverage.tex}

\subsection{Theorem 10 — High-Probability Violation Bound (Hoeffding)}

\begin{theorem}[Hoeffding High-Probability Bound]\label{thm:hoeffding_tsvr}
Let $V_1, \ldots, V_T \in \{0,1\}$ be the per-step violation indicators
produced by DC3S over a horizon of length $T$, and let
$\bar{w} = \frac{1}{T}\sum_{t=1}^T w_t$ be the mean reliability score.
Assume $V_t$ are independent and the marginal expectation bound holds:
$\mathbb{E}[V_t] \leq \alpha(1-w_t)$.  Then for any $\varepsilon > 0$:
\[
  P\!\left(\mathrm{TSVR} \;\geq\; \alpha(1-\bar{w}) + \varepsilon\right)
  \;\leq\;
  \exp\!\left(-2T\varepsilon^2\right).
\]
That is, the empirical violation rate exceeds the expectation bound by
$\varepsilon$ with probability at most $\exp(-2T\varepsilon^2)$, which is
exponentially small in $T$.
\end{theorem}

\begin{proof}[Proof sketch]
Define $Z_t = V_t - \alpha(1-w_t)$ so that $\mathbb{E}[Z_t] \leq 0$ and
$Z_t \in [-\alpha(1-w_t),\, 1 - \alpha(1-w_t)] \subseteq [-1, 1]$.
The empirical mean $\bar{Z} = \frac{1}{T}\sum_t Z_t$ satisfies
$\mathrm{TSVR} - \alpha(1-\bar{w}) = \bar{Z}$.
By the Hoeffding inequality for bounded random variables in $[-1, 1]$:
\[
  P(\bar{Z} \geq \varepsilon) \leq \exp\!\left(-\frac{2T^2\varepsilon^2}{T \cdot 2^2}\right)
  = \exp(-2T\varepsilon^2/4 \cdot 4) = \exp(-2T\varepsilon^2).
\]
Wait — applying Hoeffding directly with $b_t - a_t = 1$ (range 1, bounded
in $[0,1]$):
$P(\bar{Z} \geq \varepsilon) \leq \exp(-2T\varepsilon^2)$.  $\square$
\end{proof}

\paragraph{Practical implication.}
For the ORIUS-Bench default $T = 240$ steps (5 seeds $\times$ 48 steps) and
$\varepsilon = 0.05$, the tail probability is
$\exp(-2 \times 240 \times 0.0025) = \exp(-1.2) \approx 0.30$.
For $\varepsilon = 0.10$ it drops to $\exp(-4.8) \approx 0.008$, meaning
fewer than 1-in-125 runs will exceed the bound by 10\,pp.
The function \texttt{hoeffding\_violation\_bound()} in
\texttt{src/orius/dc3s/coverage\_theorem.py} computes this tail probability
for any $(T, \alpha, \bar{w}, \varepsilon)$.

\subsection{Joint Guarantee}

Combining Theorems~9 and~10:
\begin{itemize}
  \item \emph{Per group}: each reliability-group sub-population has coverage
        $\geq 1-\alpha$ (conditional guarantee, Theorem~9).
  \item \emph{Per run}: the TSVR of a single run exceeds the expectation bound
        by $\varepsilon$ with probability $\leq \exp(-2T\varepsilon^2)$
        (concentration guarantee, Theorem~10).
\end{itemize}
Together they replace the earlier marginal-expectation-only bound with a
significantly stronger two-level guarantee that addresses both the committee's
subgroup-coverage and high-probability objections.
