%% ch23_committee_response.tex
%% Anticipated PhD Committee Questions and Structured Responses
%% Addresses the five primary critique angles of an R1 PhD committee.

\subsection{Anticipated Committee Questions and Responses}
\label{sec:committee_response}

This section addresses the most common PhD committee objections to a
conformal-safety-shield thesis, structured by faculty perspective.

\input{assets/tables/generated/tbl_constraint_satisfaction.tex}
\input{assets/tables/generated/tbl_coverage_vs_alpha.tex}
\input{assets/tables/generated/tbl_latency_benchmark.tex}
\input{assets/tables/generated/tbl_complexity_analysis.tex}
\input{assets/tables/generated/tbl_multi_domain_evidence_gate.tex}

% ------------------------------------------------------------------
\subsubsection*{Q1 (Control Theory): Is the repair function always feasible?}

\textbf{Question.} A safety filter that may fail to produce a feasible repaired
action is not a safety guarantee — it is merely a heuristic.

\textbf{Response.}
The DC3S repair stage (\S\ref{sec:shield}) constructs the tightened action set
$[\ell_t, u_t]$ from the uncertainty-inflated constraint bound.  By
Proposition~1 (inflation monotonicity), $\ell_t \leq u_t$ whenever
$\text{infl}_t \geq 1$, which is enforced by \texttt{verify\_inflation\_geq\_one()}
before the shield is invoked.  The repair then sets
$a_\text{safe} = \text{clamp}(a_\text{candidate},\, \ell_t,\, u_t)$,
which is always defined and always satisfies $a_\text{safe} \in [\ell_t, u_t]$.
This is a closed-form projection, not a solver call that can become infeasible.

The empirical counterpart appears in Table~\ref{tbl:constraint_satisfaction}:
across the defended Battery, nuPlan AV, and Healthcare rows, the post-repair
constraint satisfaction rate is \textbf{100\%}.  The table is not used as a
universal deployment claim; it verifies that the closed-form repair semantics
used by the three promoted rows match Proposition~1.

% ------------------------------------------------------------------
\subsubsection*{Q2 (Statistics): Why not use a split-conformal test for coverage?}

\textbf{Question.} Split conformal prediction gives exact marginal coverage.
Why use a reliability-weighted variant that modifies the quantile?

\textbf{Response.}
Standard split conformal prediction assumes exchangeable residuals.  Under
degraded telemetry, residuals are \emph{not} exchangeable: a step with
$w_t = 0.1$ (severe dropout) produces a fundamentally different residual
distribution than a step with $w_t = 0.95$ (clean signal).  Treating them
identically inflates intervals for clean steps (conservative) while failing
to inflate enough for degraded steps (unsafe).

DC3S replaces the fixed quantile $q_{1-\alpha}$ with a reliability-weighted
quantile $q_{1-\alpha(1-w_t)}$, which broadens intervals exactly in proportion to
signal degradation.  Theorem~1 (\S\ref{sec:theory}) proves this still achieves
marginal coverage $\geq 1-\alpha$ in expectation; Theorem~9 proves group-conditional
coverage per reliability bin; Theorem~10 gives the Hoeffding tail bound.

Table~\ref{tbl:coverage_vs_alpha} shows that across the three defended rows and
$\alpha \in \{0.05, 0.10, 0.15, 0.20\}$, empirical PICP exceeds the nominal
$(1-\alpha)$ target, confirming calibration on the promoted evidence surface.

% ------------------------------------------------------------------
\subsubsection*{Q3 (CS/Systems): Are the latency numbers real, and do they scale?}

\textbf{Question.} Papers often report latency for a toy single-domain prototype.
Does DC3S scale beyond the witness row?

\textbf{Response.}
Table~\ref{tbl:latency_benchmark} reports measured wall-clock p50/p95/p99 for
the promoted Battery, nuPlan AV, and Healthcare rows on a single CPU core
(Python 3.11, Intel Xeon).  All three rows achieve p99 $< 0.35$\,ms, far below
the 50\,ms production SLA used as the runtime gate.  The result is not a claim
of road or clinical deployment; it is a compute-budget check for the defended
runtime contract.

The algorithmic complexity analysis (Table~\ref{tbl:complexity_analysis}) shows that
the dominant cost is $O(W \log W)$ for robust OQE (sort for trimmed-mean), where
$W = 20$ by default.  All other stages are $O(d)$ or $O(1)$, so the system scales
linearly in action dimensionality and is independent of the number of domains
(each step processes one domain).

% ------------------------------------------------------------------
\subsubsection*{Q4 (EE/Domain): Which battery metric is governing?}

\textbf{Question.} Earlier diagnostic tables mixed observed-state ORIUS-Bench
battery rows with the locked hidden-state battery witness.  Which row governs
the primary application?

\textbf{Response.}
The promoted battery result is the locked DE/US witness benchmark, where ORIUS
reduces the governing true-SOC TSVR from 0.008333 to 0.000000.  Deprecated
observed-state ORIUS-Bench battery diagnostics are not used as the release
metric because they measure the telemetry surface rather than the hidden physical
state that the witness row is designed to protect.

The ORIUS-Bench battery track measures TSVR on the observed state, not the hidden
state, so the sign reversal is expected and is a known limitation of the synthetic
battery benchmark (discussed in \S\ref{sec:limitations}).  In the promoted
evidence package, Battery is judged by the locked hidden-state witness row, AV by
the all-zip grouped nuPlan replay runtime, and Healthcare by its bounded
monitoring row.

% ------------------------------------------------------------------
\subsubsection*{Q5 (Applied Math): Are Theorems 9–11 independent of domain?}

\textbf{Question.} The theorems appear to be stated for a generic domain.
Are there hidden domain-specific assumptions?

\textbf{Response.}
Theorems~9 and~10 require only: (i) a measurable safety predicate
$\phi: \mathcal{S} \to \{0,1\}$; (ii) a reliability weight $w_t \in [0,1]$;
and (iii) a conformal inflation $\text{infl}_t \geq 1$.  These are satisfied by
definition in every DC3S adapter — the predicate is the domain's \texttt{check\_safety()}
method, the weight comes from \texttt{compute\_reliability()}, and the inflation is
verified by Proposition~1.  No domain-specific distributional assumption is made.

Theorem~11 additionally requires that at most $f < 1/3$ of readings in the OQE window
are adversarially spoofed.  This is a protocol-level bound (depends on the attacker
model), not a domain-physics assumption.  It holds whenever the system is not fully
compromised, which is the standard Byzantine fault-tolerance assumption in distributed
systems \cite{lamport1982byzantine}.

The three-domain empirical validation (Tables~\ref{tbl:multi_domain_evidence_gate}
and~\ref{tab:adversarial_tsvr}) provides the empirical counterpart for the
promoted rows: Battery supplies the witness row, nuPlan AV supplies the
all-zip grouped replay row, and Healthcare supplies the bounded monitoring row.
The theorems themselves are domain-agnostic by construction---Theorems~9 and~10
require no domain-physics assumptions, only conditions (C1)--(C3) of
Definition~\ref{def:compatible_cps}---but the \emph{empirical validation depth}
is tiered: battery carries locked real-data evidence, the full energy-optimisation
pipeline, and hardware-in-loop results that the bounded AV and Healthcare rows do not yet replicate.
The formal guarantee applies universally; the evidence depth does not.

\subsubsection*{Summary}

All five critique axes have formal answers backed by locked empirical evidence.
The key take-away is that DC3S is a \emph{structurally sound} safety layer:
repair feasibility is guaranteed algebraically (not empirically), coverage is
provably maintained across reliability groups, latency is negligible relative to
any realistic SLA, and the theorems are domain-agnostic by construction.
