\section{Closing ORIUS Program Block}\label{sec:closing_longform}

The closing block below brings the full thesis boundary language directly into
the canonical paper build. This keeps \texttt{paper.pdf} honest about what Paper~1
establishes on the battery witness, what Papers~2--6 add as bounded ORIUS
program contributions, and what still remains outside the defended evidence
surface.

\paragraph{Program hierarchy and proof boundary.}
The ORIUS dissertation has a two-spine architecture.  Paper~1 is the full
proof witness: it owns the battery domain, the locked T1--T8 theorem ladder,
and the governed economics.  Papers~2--6 are explicit second-spine
contributions with a clear dependency structure:
\begin{enumerate}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item Papers~2 and~3 depend on the temporal battery ladder (T5--T7) and
        stay bounded to the battery domain.  Their claim surfaces are the
        validity-horizon engine and the four-policy graceful-fallback
        benchmark, respectively.
  \item Paper~4 is a benchmark-contract contribution, not a theorem
        contribution.  Its claim surface is replayability, schema stability,
        and artifact determinism of ORIUS-Bench.
  \item Paper~5 is a bounded composition extension.  It demonstrates a
        non-composition counterexample on a shared-feeder plant and provides
        centralized and distributed coordination protocols, but does not claim
        a complete composition theory.
  \item Paper~6 is a bounded runtime-governance extension.  It provides
        the CertOS lifecycle, three invariants, and an audit chain, scoped to
        the battery operational regime with bounded lifecycle semantics.
\end{enumerate}
No paper in the 2--6 range upgrades bounded evidence into theorem closure.
Battery remains the only full proof witness.  This is an honest program-first
dissertation, not a broader universal claim.

\PaperImportStructuredFile{../chapters/ch33_what_battery_thesis_proves.tex}
\PaperImportStructuredFile{../chapters/ch34_outside_current_evidence.tex}
\PaperImportStructuredFile{../chapters/ch35_deployment_path_verification_discipline.tex}
\PaperImportStructuredFile{../chapters/ch36_conclusion.tex}

