% =============================================================================
% Appendix C: Full Proofs
% ORIUS/DC3S Battery-Safety Thesis
% =============================================================================
\chapter{Full Proofs}
\label{app:proofs}

This appendix is the authoritative proof surface for the defended ORIUS
theorem program, but it is not a flat headline ladder.  The active audit
distinguishes flagship defended rows, supporting defended rows, and visible
draft/non-defended extensions.  T9 and T10 are supporting defended only after
the domain-discharge artifacts close their empirical constants.  The battery witness remains the deepest proof
chain, so the full step-by-step derivations below begin with the reference
ladder:
\begin{enumerate}
  \item OASG existence,
  \item one-step safety preservation,
  \item the ORIUS core bound,
  \item no free safety,
  \item certificate validity horizon,
  \item certificate expiration bound,
  \item feasible fallback existence, and
  \item graceful degradation dominance.
\end{enumerate}
The certificate half-life result remains a corollary, and safe-budget
monotonicity remains a proposition rather than part of the T1--T8 reference
ladder.  The later extension theorems T9--T11 and L1--L4 are indexed here so
that Chapter~\ref{ch:universality-completeness}, this appendix, and
Appendix~\ref{app:verified-theorems-gap-audit} point to the same proof
surface, but those extension rows do \emph{not} count toward the flagship
defended novelty surface unless separately promoted by the active theorem
audit.  The broader integrated theorem-like surface, including precursor
results and supporting chapter statements, is catalogued separately in
Appendix~\ref{app:integrated-theorem-surface-register}.

\paragraph{Mathematical contract vs. domain discharge.}
The proof copies below use only stated assumptions and typed obligations as
mathematical premises.  Domain discharge is evidence, not an extra hidden assumption: the discharge artifacts instantiate constants, verify
boundary-mass and mixing witnesses, and bind the proof kernels to the promoted
bounded Battery, AV, and Healthcare artifacts.  Thus T9/T10/T11 are
contract-universal, not unrestricted-global; new physical-AI domains may reuse
the proof pattern only after their adapters, safe sets, coverage witnesses, and
fallback obligations are discharged by domain-specific evidence.

%---------------------------------------------------------------------
\section{C.1\quad Battery-Domain OASG Existence}\label{sec:proof-oasg}
\begin{theorem}[restated]
Under A1, A2, A4, A11, and A12, for dropout rate $d>0$ and any optimal
arbitrage controller, $\exists\;t^*$ with $\phi_{\mathrm{obs}}(a_{t^*})=1$,
$\phi_{\mathrm{true}}(a_{t^*})=0$.
\end{theorem}
\begin{proof}
Fix $\eta>0$.  By A11, the battery witness controller can be driven to a
boundary-sensitive state under admissible demand, so there exists a step $t^*$
and an admissible released action $a_{t^*}$ such that the observed-state
forward prediction lies on the safe side of the upper SOC boundary while
remaining arbitrarily close to it:
\[
  \SOC_{\max}-\eta
  \le
  \hat{x}_{t^*} + \Delta(a_{t^*})
  \le
  \SOC_{\max}.
\]
By A2 and the standing dropout rate $d>0$, there exists an admissible
positive observation gap $\delta$ on a degraded-telemetry window.  Choose that
gap so that $\eta<\delta\le g_{\max}$ and let
\[
  x_{t^*} = \hat{x}_{t^*} + \delta.
\]
By A12, the fault process is conditionally independent of the controller
action stream given the history, so the boundary-sensitive witness step and
the admissible telemetry fault can co-occur on the same realized history.
Now choose the admissible one-step model residual $\epsilon_{t^*}=0$, which is
allowed by A1 because $0\in[-\epsilon_{\mathrm{model}},\epsilon_{\mathrm{model}}]$.

The controller evaluates the action through the observed state only, so the
observed successor satisfies
\[
  \hat{x}_{t^*+1}
  =
  \hat{x}_{t^*} + \Delta(a_{t^*})
  \le
  \SOC_{\max},
\]
hence $\phi_{\mathrm{obs}}(a_{t^*})=1$.  Using A4 for the shared one-step
dynamics, the true successor is
\[
  x_{t^*+1}
  =
  x_{t^*} + \Delta(a_{t^*}) + \epsilon_{t^*}
  =
  \hat{x}_{t^*} + \Delta(a_{t^*}) + \delta.
\]
Combining this identity with the lower side of the observed-state witness
gives
\[
  x_{t^*+1}
  \ge
  (\SOC_{\max}-\eta) + \delta
  =
  \SOC_{\max} + (\delta-\eta)
  >
  \SOC_{\max}.
\]
Therefore $\phi_{\mathrm{true}}(a_{t^*})=0$ while
$\phi_{\mathrm{obs}}(a_{t^*})=1$, which is exactly the battery witness-row
observation-action safety gap.  \qedhere
\end{proof}

\section{C.2\quad One-Step Safety Preservation}\label{sec:proof-one-step}
\begin{theorem}[restated]
Under A1--A5, A7, if $\asafe\in\Aset$ and $x_t\in\Crac$, then
$\phi(x_{t+1})=1$, where the tightened margin used to define $\Aset$
already absorbs the one-step model-error allowance.
\end{theorem}
\begin{proof}
Fix a step $t$ and assume the standing conditions A1--A5 and A7.  Because
$x_t\in\Crac$, the true state lies in the reliability-adjusted conformal tube
issued from the current observation.  In particular, membership in that tube gives
\[
  |x_t-\hat{x}_t| \le m_t,
\]
where $m_t$ is the inflation-adjusted half-width determined by A2 and A5.
This step is deterministic once the theorem is conditioned on $x_t\in\Crac$;
it does not introduce a new probability statement.  By A5, the shield uses
the absorbed tightening margin
\[
  m_t^\star = m_t + \epsilon_{\mathrm{model}}
\]
when constructing $\Aset$.  Since $\asafe\in\Aset$, the observed-state forward
prediction therefore satisfies
\[
  \SOC_{\min}+m_t^\star
  \le
  \hat{x}_t + \Delta(\asafe)
  \le
  \SOC_{\max}-m_t^\star,
\]
with $\Delta(\asafe)$ denoting the one-step battery state update implied by
the known dynamics from A4.  Now write the true next state as
\[
  x_{t+1} = x_t + \Delta(\asafe) + \epsilon_t,
\]
where the model residual $\epsilon_t$ is bounded by A1.  Substituting
$x_t = \hat{x}_t + (x_t-\hat{x}_t)$ and using $|x_t-\hat{x}_t|\le m_t$ yields
\[
  x_{t+1}
  \ge (\hat{x}_t+\Delta(\asafe)) - m_t - |\epsilon_t|
  \ge \SOC_{\min} + m_t^\star - m_t - \epsilon_{\mathrm{model}}
  =
  \SOC_{\min} + (m_t+\epsilon_{\mathrm{model}}) - m_t - \epsilon_{\mathrm{model}}
  =
  \SOC_{\min}.
\]
The same substitution on the upper side gives
\[
  x_{t+1}
  \le (\hat{x}_t+\Delta(\asafe)) + m_t + |\epsilon_t|
  \le \SOC_{\max} - m_t^\star + m_t + \epsilon_{\mathrm{model}}
  =
  \SOC_{\max} - (m_t+\epsilon_{\mathrm{model}}) + m_t + \epsilon_{\mathrm{model}}
  =
  \SOC_{\max}.
\]
Therefore
\[
  \SOC_{\min} \le x_{t+1} \le \SOC_{\max}.
\]
This is exactly the safety predicate $\phi(x_{t+1})=1$.  The argument uses
only the current-step information certified by A7 and therefore preserves
causality.  Hence any action selected from $\Aset$ while the current state is
inside $\Crac$ keeps the one-step successor in the safe set.  The theorem
stops at this one-step conditional statement; it does not itself quantify how
often $x_t\in\Crac$ occurs.
\qedhere
\end{proof}

\section{C.3\quad Battery ORIUS Core Bound}\label{sec:proof-core-bound}
\begin{theorem}[restated]
Under A1--A5, A7, and the explicit battery risk-envelope contract
$\Prob[Z_t=1\mid\mathcal{H}_t]\le\alpha(1-w_t)$, one has
$\E[V]\le\alpha\,T\,(1-\bar{w})$.
\end{theorem}
\begin{proof}
Let
\[
  V = \sum_{t=1}^{T} Z_t,
  \qquad
  Z_t = \mathbf{1}\!\bigl[\lnot \phi(x_t)\bigr],
\]
so that $V$ counts the number of true-state safety violations across the
episode.  Because each $Z_t$ is an indicator, $\E[Z_t]=\Prob[Z_t=1]$.
Applying linearity of expectation, which does not require independence across
time, gives
\[
  \E[V] = \sum_{t=1}^{T}\E[Z_t] = \sum_{t=1}^{T}\Prob[Z_t=1].
\]
The theorem-local battery contract supplies the per-step conditional envelope
\[
  \Prob[Z_t=1\mid\mathcal{H}_t] \le \alpha(1-w_t).
\]
Taking expectations of both sides gives
\[
  \Prob[Z_t=1] = \E[\Prob[Z_t=1\mid\mathcal{H}_t]] \le \alpha(1-w_t).
\]

Substituting the per-step bound into the expectation identity yields
\[
  \E[V]
  \le \sum_{t=1}^{T} \alpha(1-w_t)
  = \alpha \sum_{t=1}^{T}(1-w_t).
\]
Now expand the sum:
\[
  \sum_{t=1}^{T}(1-w_t)
  = T - \sum_{t=1}^{T} w_t
  = T\left(1 - \frac{1}{T}\sum_{t=1}^{T} w_t\right).
\]
By definition of the episode-average reliability,
\[
  \bar{w} = \frac{1}{T}\sum_{t=1}^{T} w_t.
\]
Replacing the bracketed term with $\bar{w}$ gives
\[
  \E[V] \le \alpha\,T\,(1-\bar{w}).
\]
This is the desired aggregation corollary.  The actual burden is the
per-step envelope hypothesis
$\Prob[Z_t=1\mid\mathcal{H}_t]\le\alpha(1-w_t)$; once that T3a-style
derivation is available, the displayed $\E[V]\le\alpha T(1-\bar w)$ bound
follows by linearity of expectation alone.  The derivation invokes only
finite-horizon identities and the explicit theorem-local battery risk
budget; no temporal independence assumption is introduced.  Therefore the
battery ORIUS core bound holds for any admissible horizon $T$ on the
defended replay surface whenever that per-step budget contract has been
justified.
\qedhere
\end{proof}

\section{C.4\quad Observation Necessity / No Free Safety}\label{sec:proof-no-free}
\begin{theorem}[restated]
Under A1, A2, A4, and A11, for any quality-ignorant $\pi$ and $d>0$, $\exists$
admissible fault sequence with $\TSVR(\pi)>0$.
\end{theorem}
\begin{proof}
Fix a quality-ignorant controller $\pi$.  By definition, its effective safety
margin does not depend on the runtime reliability score, so there exists a
constant margin $m^\pi\ge 0$ used across all observation qualities.  By A11,
the battery witness controller can be driven to a boundary-sensitive state
under admissible demand.  Consequently, for every $\eta>0$ there exists a
step $t^\star$ and an action released by $\pi$ such that the observed-state
forward prediction nearly saturates the controller's fixed margin on the upper
SOC boundary:
\[
  \hat{x}_{t^\star} + \Delta(\pi(\hat{x}_{t^\star}))
  \ge
  \SOC_{\max} - m^\pi - \eta.
\]

Because $d>0$, A2 admits an admissible degraded-observation gap.  Choose a
fault realization with
\[
  \delta \in (m^\pi+\eta,\; g_{\max}],
  \qquad
  x_{t^\star}=\hat{x}_{t^\star}+\delta.
\]
Now set the admissible one-step model residual to zero, which is allowed by A1.
Using A4 for the shared battery dynamics,
\[
  x_{t^\star+1}
  =
  x_{t^\star} + \Delta(\pi(\hat{x}_{t^\star})) + \epsilon_{t^\star}
  =
  \hat{x}_{t^\star} + \Delta(\pi(\hat{x}_{t^\star})) + \delta.
\]
Substituting the observed-state witness inequality gives
\[
  x_{t^\star+1}
  \ge
  \SOC_{\max} - m^\pi - \eta + \delta
  =
  \SOC_{\max} + (\delta-m^\pi-\eta)
  >
  \SOC_{\max}.
\]
The observed state therefore still appears feasible relative to the fixed
margin $m^\pi$, while the same released action produces a true-state boundary
crossing.  Hence the controller incurs at least one true-state violation on
the constructed admissible sequence, so $\TSVR(\pi)>0$.  The witness is
finite-horizon and class-scoped exactly as claimed in T4.  \qedhere
\end{proof}

\section{C.5\quad Certificate Validity Horizon}\label{sec:proof-cert-valid}
\begin{theorem}[restated]
Under A1, A2, A4--A7, the battery safety certificate remains valid for the
largest integer horizon $\tau_t$ whose forward tube stays inside the SOC
constraint set.  In the active audit this row is treated as a constructive
definition rather than as part of the defended headline theorem count.
\end{theorem}
\begin{proof}
The forward tube propagates the reliability-inflated interval under the
certified safe action while adding the worst-case sub-Gaussian drift radius
$\sigma_d\sqrt{\Delta}$.  The certificate remains valid exactly while that
tube is contained in $[\SOC_{\min},\SOC_{\max}]$.  The runtime helper
\texttt{certificate\_validity\_horizon()} computes this maximal integer
horizon directly from the chapter definition.  Both T5 and T6 are
registered in \texttt{THEOREM\_REGISTER}
(\texttt{src/orius/dc3s/theoretical\_guarantees.py}) and enforced at
runtime: the API router checks $\tau_t \ge 1$ at each step and fails the
guarantee when violated.  \qedhere
\end{proof}

\section{C.6\quad Certificate Expiration Bound}\label{sec:proof-expiration}
\begin{theorem}[restated]
Under A4, A7, A9 and confidence level $\delta\in(0,1)$, the battery
certificate horizon satisfies the lower bound
\[
  \tau_t \ge
  \left\lfloor
    \frac{\delta_{\mathrm{bnd},t}^2}
    {2\sigma_d^2 \log(2/\delta)}
  \right\rfloor.
\]
\end{theorem}
\begin{proof}
For a $\sigma_d$-sub-Gaussian disturbance, the reflection-principle /
first-passage tail inequality gives
\[
  \Prob\!\bigl(|S_\Delta| \ge \delta_{\mathrm{bnd},t}\bigr)
  \le
  2\exp\!\left(
    -\frac{\delta_{\mathrm{bnd},t}^2}{2\sigma_d^2\Delta}
  \right).
\]
Demanding the right-hand side be at most $\delta$ and solving for $\Delta$
yields
\[
  \Delta \le
  \frac{\delta_{\mathrm{bnd},t}^2}
       {2\sigma_d^2 \log(2/\delta)}.
\]
Flooring gives the integer lower bound, and the runtime helper
\texttt{certificate\_expiration\_bound()} implements this confidence-aware
expression directly.
\qedhere
\end{proof}

\section{C.7\quad Certificate Half-Life Corollary}\label{sec:proof-halflife}
\begin{corollary}[restated]
The certificate half-life remains a corollary rather than part of the
flagship T1--T8 ladder.
\end{corollary}
\begin{proof}
The current battery blackout evidence supports a bounded safe-hold claim more
strongly than a finite empirically identified half-life law.  The corollary
therefore remains analytical and should be interpreted together with the more
conservative wording in Chapter~\ref{ch:cert-half-life-blackout}.  \qedhere
\end{proof}

\section{C.8\quad Feasible Fallback Existence}\label{sec:proof-fallback}
\begin{theorem}[restated]
Under A1, A4, and A8, there exists a battery piecewise fallback rule that
either safely holds on the interior or executes a boundary-aware safe-landing
recovery action; otherwise the runtime fails closed instead of asserting a
false fallback guarantee.
\end{theorem}
\begin{proof}
Zero dispatch is the constructive hold action on the interior.  If the current
SOC has enough slack against the one-step model error, the next state remains
inside the admissible interval.  If the current SOC is safe but too close to a
boundary for zero dispatch to be certified, the safe-landing controller
computes a charge or discharge action that moves the SOC toward the interior
safe zone while respecting the active battery limits.  The helper
\texttt{certify\_fallback\_existence()} now checks exactly this piecewise
surface and returns infeasible rather than over-claiming when neither case is
safe.  \qedhere
\end{proof}

\section{C.9\quad Graceful Degradation Dominance}\label{sec:proof-graceful}
\begin{theorem}[restated]
Under A1--A5 and A8, for any paired graceful and uncontrolled violation
sequences generated under the same admissible fault trace, the sequence-level
comparison
\[
  z_t^{\mathrm{graceful}} \le z_t^{\mathrm{uncontrolled}}
  \quad\forall t
\]
implies
\[
  V^{\mathrm{graceful}} \le V^{\mathrm{uncontrolled}},
\]
with strict inequality whenever the graceful sequence is strictly smaller on at
least one step.
\end{theorem}
\begin{proof}
Summing the stepwise inequality over the shared trace gives
\[
  \sum_{t=1}^{T} z_t^{\mathrm{graceful}}
  \le
  \sum_{t=1}^{T} z_t^{\mathrm{uncontrolled}},
\]
which is exactly the displayed dominance statement for the cumulative
violation counts.  Strict inequality follows if the graceful sequence is
strictly smaller on at least one step.  The helper
\texttt{evaluate\_graceful\_degradation\_dominance()} implements this
paired-sequence comparison directly.  The appendix does not promote a broader
absorbing-landing dynamics theorem beyond this scoped supporting surface.
\qedhere
\end{proof}

\section{C.10\quad Safe-Budget Monotonicity Proposition}\label{sec:proof-budget}
\begin{proposition}[restated]
Under A1--A5, $B_t=|\Aset|\cdot\Delta t$ is monotone increasing in $w_t$.
\end{proposition}
\begin{proof}
$|\Aset|=\SOC_{\max}-\SOC_{\min}-2m_t$ and
$m_t=q_t/(w_t+\epsilon)$ is monotone decreasing in $w_t$, so $B_t$
increases with $w_t$.  \qedhere
\end{proof}

\section{C.11\quad Universal Impossibility (T9)}\label{sec:proof-universal-impossibility}
\textbf{Theorem reference.}
The authoritative statement is Theorem~\ref{thm:universal-impossibility} in
Chapter~\ref{ch:universality-completeness}.  It is repeated here as a
supporting, assumption-qualified extension so that the full proof appendix
covers the complete theorem inventory rather than only the battery witness
ladder.

\begin{proof}[Appendix proof copy]
By Assumption~A10b, the fault process is $\varphi$-mixing with geometric
decay, so a buffer gap of length $\tau_{\mathrm{mix}}$ makes sufficiently
separated windows approximately independent in total variation.  Partition the
horizon into analysis windows of length $\ell$ separated by those buffers.
Inside each analysis window, invoke the battery no-free-safety witness from
Section~\ref{sec:proof-no-free}: a controller that does not adapt its safety
margin to the reliability process admits an admissible degraded-observation
sequence with positive violation probability.  The boundary-reachability and
exogenous-fault structure are exactly the ingredients isolated in A11 and A12,
so they package that window-level witness into a constant $c>0$ such that
$p_{\mathrm{win}} \ge c\,d$ when the mean degradation rate is $d$.

Let $Z_j^{\mathrm{win}}$ indicate whether the $j$th separated window incurs a
true-state violation and define
\[
  D_j =
  Z_j^{\mathrm{win}} -
  \mathbb{E}\!\left[
    Z_j^{\mathrm{win}}
    \mid
    \mathcal{F}_{(j-1)(\ell+\tau_{\mathrm{mix}})}
  \right].
\]
The sequence $\{D_j\}$ is a martingale difference sequence with
$|D_j| \le 1$.  Azuma--Hoeffding therefore yields exponential concentration of
the centered partial sums around their mean.  Since the expected number of bad
windows is at least $kcd$ over $k$ separated windows, the probability that the
total violation count drops below a constant fraction of $kcd$ decays like
$\exp(-\Omega(dT))$.  Taking expectations gives
$\mathbb{E}[V_T] \ge \Omega(dT)$, which is the claimed
$\mathbb{E}[V_T] \ge c\,d\,T$ lower bound up to renaming the constant.
\end{proof}

\section{C.12\quad Boundary-Indistinguishability Lower Bound (T10)}\label{sec:proof-frontier-lower-bound}
\textbf{Theorem reference.}
The authoritative statement is Theorem~\ref{thm:information-theoretic-lower-bound}
in Chapter~\ref{ch:universality-completeness}.  The proof below preserves its
supporting, assumption-qualified interpretation as a lower-bound frontier
theorem under explicit extra hypotheses, including the supplied boundary-mass
and OASG-severity sequences.

\begin{proof}[Appendix proof copy]
Fix a step $t$ and the binary boundary-testing subproblem with latent
hypotheses $H_{0,t}$ and $H_{1,t}$.  Because the controller only acts through
the degraded observation channel, its decision is a binary test between the
observation laws $P_{0,t}$ and $P_{1,t}$.  By A13,
\[
  \mathrm{TV}(P_{0,t},P_{1,t}) \le w_t.
\]
Therefore Le~Cam's lemma gives
\[
  \max\!\bigl(
    \Prob[\psi_t \text{ errs}\mid H_{0,t}],
    \Prob[\psi_t \text{ errs}\mid H_{1,t}]
  \bigr)
  \ge
  \frac{1-\mathrm{TV}(P_{0,t},P_{1,t})}{2}
  \ge
  \frac{1-w_t}{2}.
\]
Whenever the unsafe-side hypothesis $H_{1,t}$ holds and the controller makes
that binary mistake, the step contributes a true-state violation and at least
$\delta_t$ severity-weighted OASG.  Therefore
\[
  \Prob[Z_t = 1]
  \ge
  \Prob(H_{1,t})\Prob[\psi_t \text{ errs}\mid H_{1,t}]
  \ge
  p_t\,\frac{1-w_t}{2}.
\]
Summing over the horizon and applying linearity of expectation gives
\[
  \mathbb{E}[V_T(\pi)]
  \ge
  \frac{1}{2}\sum_{t=1}^T p_t(1-w_t).
\]
The same events contribute at least $\delta_t$ to the OASG metric, so
\[
  \mathbb{E}[\mathrm{OASG}_T(\pi)]
  \ge
  \frac{1}{2}\sum_{t=1}^T \delta_t p_t(1-w_t).
\]
The special case $p_t \ge \alpha/2$ for all $t$ yields the stated frontier
$\mathbb{E}[V_T(\pi)] \ge \tfrac{1}{4}\alpha(1-\bar{w})T$.  The proof does not
claim a sharper constant without a domain-specific argument for the
boundary-mass and severity sequences $\{p_t,\delta_t\}$, which is exactly why
T10 is defended as an assumption-qualified lower bound rather than as a global
minimax theorem.
\end{proof}

\section{C.13\quad Typed Structural Transfer and Failure-Mode Converse (T11)}\label{sec:proof-typed-transfer}
\textbf{Theorem reference.}
The forward theorem is Theorem~\ref{thm:universality-completeness} and the
converse witness is Proposition~\ref{prop:t11-counterexamples} in
Chapter~\ref{ch:universality-completeness}.  The appendix copies the key proof
logic so the full-proofs surface remains self-contained.

\begin{proof}[Appendix proof copy]
Fix a step $t$ and condition on the causal observation history
$\mathcal{H}_t$.  Let $E_t = \{x_t \in U_t\}$ denote the event that the latent
state lies in the constructed observation-consistent set.  By the first T11
obligation,
\[
  \Prob[E_t \mid \mathcal{H}_t] \ge 1-\alpha.
\]
On $E_t$, the soundness obligation implies that every action in the tightened
safe-action set preserves safety at the next step for every state in $U_t$ and
every admissible disturbance.  If the tightened safe-action set is non-empty,
the repair-membership obligation guarantees that the released repaired action
lies inside that set and is therefore safe on $E_t$.  If the tightened
safe-action set is empty, the fallback-admissibility obligation guarantees a
safe release path from the current safe state.  In either branch,
\[
  \Prob[x_{t+1}\in\mathcal{S}\mid\mathcal{H}_t]
  \ge
  \Prob[E_t\mid\mathcal{H}_t]
  \ge
  1-\alpha.
\]
This proves the forward one-step transfer direction.  The converse is not
claimed as part of T11 itself; it is tracked separately as a structural
failure-mode witness showing that if any one of the four obligations fails,
the corresponding step in the forward proof disappears.
\end{proof}

\begin{corollary}[T10\_T11\_ObservationAmbiguitySandwich: Covered Observation-Ambiguity Optimality]
Let $B(o)=\{x:h(x)=o\}$ denote an observation ambiguity class and let
$K(o)=\bigcap_{x\in B(o)}C(x)$ denote its common safe core.  Every
mandatory-release observation-only controller satisfies
\[
  \operatorname{TSVR}(\pi)
  \ge
  \mathbb{E}_{O}\!\left[
    \min_{a\in\mathcal{A}}
    \Prob(a\notin C(X^\star)\mid O)
  \right].
\]
If ORIUS constructs a set $\mathcal{X}_t$ with
$\Prob(X_t^\star\notin\mathcal{X}_t)\le\alpha$ and releases only
$u_t\in\bigcap_{x\in\mathcal{X}_t}C(x)$, then
$\operatorname{TSVR}_{\mathrm{ORIUS}}\le\alpha$; under exact coverage, the
one-step true-state violation probability is zero.
\end{corollary}

\begin{proof}
Conditioning on $O=o$, an observation-only controller must select one action
for all states in $B(o)$.  Its conditional violation probability is
$\Prob(a\notin C(X^\star)\mid O=o)$, so no such controller can improve on the
minimum over $a$.  Averaging over observations gives the Bayes lower bound.
When $K(o)=\emptyset$, no action is safe for all states in the ambiguity
class, so any mandatory release has positive conditional risk if the unsafe
state has positive conditional mass.

For ORIUS, on the event $X_t^\star\in\mathcal{X}_t$, membership of $u_t$ in
every $C(x)$ for $x\in\mathcal{X}_t$ implies $u_t\in C(X_t^\star)$, hence no
true-state violation.  The remaining violation probability is bounded by the
coverage miss probability $\alpha$.
\end{proof}

\begin{lemma}[T11\_AV\_BrakeHold: AV Brake-Hold Runtime Lemma]
For a promoted ORIUS autonomous-vehicle replay row, suppose the emitted
domain runtime contract witness has source theorem T11, T11 status
\texttt{runtime\_linked}, no failed T11 obligations, a valid certificate,
and a passing true brake-hold postcondition in the runtime trace.  Then that
row is a bounded runtime instance of the forward T11 transfer surface for the
longitudinal brake-hold contract.
\end{lemma}

\begin{proof}
The witness records the four executable T11 obligations through the typed
contract summary and records the AV-specific true postcondition separately
as \texttt{domain\_postcondition\_passed}.  The stated hypotheses therefore
instantiate the coverage, soundness, repair-membership, and fallback
admissibility steps used in the T11 proof, and the runtime postcondition
checks the narrowed brake-hold safety predicate for the emitted next state.
The result is only row-local and contract-local; it does not assert complete
autonomous-driving closure.
\end{proof}

\begin{lemma}[T11\_HC\_FailSafeRelease: Healthcare Fail-Safe Release Runtime Lemma]
For a promoted ORIUS healthcare monitoring row, suppose the emitted domain
runtime contract witness has source theorem T11, T11 status
\texttt{runtime\_linked}, no failed T11 obligations, a valid certificate,
and a passing true fail-safe alert-release postcondition in the runtime
trace.  Then that row is a bounded runtime instance of the forward T11
transfer surface for the healthcare fail-safe release contract.
\end{lemma}

\begin{proof}
The witness records the typed T11 summary emitted by the universal runtime
and separately records the healthcare-specific true alert-release predicate.
If the T11 status is not \texttt{runtime\_linked}, if any obligation fails,
if the certificate is invalid, or if the postcondition fails, the witness is
not passing.  A passing witness therefore supplies exactly the forward
one-step T11 obligations plus the narrowed clinical-monitoring runtime
postcondition.  It does not assert regulated clinical deployment closure.
\end{proof}

\begin{lemma}[T6\_AV\_FallbackValidity: AV One-Step Fallback Certificate Lemma]
For a promoted ORIUS AV replay row, if the certificate records the
\texttt{single\_step\_fallback} validity scope, the released action is the
full-brake fail-safe action, the T11 runtime witness is linked, and the true
brake-hold postcondition passes, then the certificate has a theorem-backed
one-step validity horizon for that row.
\end{lemma}

\begin{proof}
The helper implementing the certificate semantics assigns
$H_t=1$, half-life one, and expiration at the next step only in the fallback
branch where the emitted action is the domain fail-safe action.  The runtime
witness separately checks T11 linkage and the true brake-hold postcondition.
Thus the certificate is valid only for the single emitted transition; no
multi-step validity is inferred from degraded observation.
\end{proof}

\begin{lemma}[T6\_HC\_FallbackValidity: Healthcare One-Step Fallback Certificate Lemma]
For a promoted ORIUS healthcare monitoring row, if the certificate records
the \texttt{single\_step\_fallback} validity scope, the released action is the
max-alert fail-safe action, the T11 runtime witness is linked, and the true
fail-safe alert-release postcondition passes, then the certificate has a
theorem-backed one-step validity horizon for that row.
\end{lemma}

\begin{proof}
The certificate-validity helper gives fallback releases a positive horizon
only for the domain fail-safe action and only for one step.  T11 linkage and
the healthcare postcondition are checked by the runtime witness.  Therefore
the row satisfies the formal validity predicate for the emitted transition,
while degraded observations are not promoted to multi-step certificates.
\end{proof}

\begin{lemma}[T\_EQ\_Battery\_RuntimeArtifactPackage: Battery Equal Artifact Discipline Package]
If the battery report directory contains runtime-denominator comparator,
ablation, negative-control, trace, utility, theorem, proof-appendix, and
reproducibility artifacts satisfying the equal-domain gate, then the battery
row satisfies the repository's equal artifact-discipline contract.
\end{lemma}

\begin{proof}
The equal-domain gate is an executable artifact obligation, not a new safety
theorem.  It checks that the battery evidence package contains a runtime
trace, runtime-native comparator rows, runtime-native ablation rows,
runtime-native negative controls, a proof-surface anchor, and a
reproducibility manifest.  It also checks the non-vacuity condition that the
ORIUS row has zero TSVR, certificate/witness closure, and useful work greater
than the degenerate safe-fallback comparator.  If any required artifact is
missing or proxy-labelled, the validator fails closed.  Therefore a passing
gate is exactly the stated artifact-discipline contract.
\end{proof}

\begin{lemma}[T\_EQ\_AV\_RuntimeArtifactPackage: AV Equal Artifact Discipline Package]
If the AV report directory contains runtime-denominator comparator,
ablation, negative-control, trace, utility, theorem, proof-appendix, and
reproducibility artifacts satisfying the equal-domain gate, then the bounded
AV replay row satisfies the repository's equal artifact-discipline contract.
\end{lemma}

\begin{proof}
The gate requires the AV T11 brake-hold runtime lemma and the AV T6
one-step fallback-validity lemma, then separately requires domain-native
runtime denominator rows for baselines, ablations, negative controls, utility,
and reproducibility.  The ORIUS row must have zero TSVR, valid runtime
certificates, passing runtime witnesses, and useful work greater than the
always-brake comparator.  The conclusion is limited to artifact discipline
for the promoted replay contract; it does not assert full autonomous-driving
closure.
\end{proof}

\begin{lemma}[T\_EQ\_HC\_RuntimeArtifactPackage: Healthcare Equal Artifact Discipline Package]
If the healthcare report directory contains runtime-denominator comparator,
ablation, negative-control, trace, utility, theorem, proof-appendix, and
reproducibility artifacts satisfying the equal-domain gate, then the bounded
healthcare monitoring row satisfies the repository's equal
artifact-discipline contract.
\end{lemma}

\begin{proof}
The gate requires the healthcare T11 fail-safe release runtime lemma and the
healthcare T6 one-step fallback-validity lemma, then requires native runtime
denominator rows for baselines, ablations, negative controls, utility, and
reproducibility.  The ORIUS row must have zero TSVR, valid runtime
certificates, passing runtime witnesses, and useful work greater than the
always-alert comparator.  The result is an artifact-discipline guarantee for
the promoted monitoring contract only; it does not assert regulated clinical
deployment readiness.
\end{proof}

%% ====================================================================
%%  C.14  L1 — Rate-Distortion Safety Law
%% ====================================================================
\section{L1 — Rate-Distortion Safety Law}\label{sec:proof-L1}

\begin{theorem}[L1: Rate-Distortion Safety Law]
Let $X\in\{0,1\}$ denote the latent safe/unsafe state with entropy
$H(X)$, let the observation channel have capacity $C$ bits per use, and
let $\alpha\in(0,1)$ be the conformal miscoverage level.  On the
\emph{stylized binary-source surrogate} used by the ORIUS law helper, we
define the lower-envelope safety distortion by
\[
  D^{*}(C)\;\ge\;\alpha\,\max\!\bigl(0,\,1-C/H(X)\bigr).
\]
\end{theorem}

\begin{proof}
This law is a scoped modeling surrogate rather than a standalone
information-theoretic converse.  The ORIUS extension surface adopts the
normalised unresolved-state fraction
$\max(0,1-C/H(X))$ as a binary-source distortion envelope and scales it
by the runtime risk budget $\alpha$.  Substituting that surrogate
definition yields the displayed lower envelope immediately.  The clamp at
zero handles the full-channel regime $C\ge H(X)$.
\end{proof}

\paragraph{Worked example.}
Battery SOC monitoring: $H(X)=1$ bit (binary safe/unsafe), $C=0.7$ bits
(30\% packet loss degrades the channel), $\alpha=0.10$.
Then $D^{*}\ge 0.10\times(1-0.7)=0.03$: at least a 3\% safety-violation
proxy is reported by the stylized law helper.  This is a bounded design
calculation, not a defended impossibility claim for the live battery
pipeline.

\paragraph{Scope boundary / non-claim.}
L1 is kept in the appendix because it explains the extension-law
calculators, but the active audit does \emph{not} treat this linear
lower envelope as a fully discharged Fano-tight converse.

%% ====================================================================
%%  C.15  L2 — Capacity Bridge
%% ====================================================================
\section{L2 — Capacity Bridge}\label{sec:proof-L2}

\begin{theorem}[L2: Capacity Bridge]
Let $w_t\in[0,1]$ be the OQE reliability weight.  On a fixed evaluation
surface where a domain-specific proxy constant $\kappa_d>0$ has been
calibrated, the ORIUS law helper uses the bridge
\[
  \bar w\;\le\;\kappa_d\,\frac{C}{H(X)},
\]
as a capacity proxy.
\end{theorem}

\begin{proof}
This appendix now treats L2 as an explicit proxy bridge, not as a proved
theorem of the live OQE.  Once $\kappa_d$ is fixed for the chosen
surface, the displayed inequality is exactly the bridge relation that the
code helper checks.  A separate empirical or theoretical identification
of $\kappa_d$ is required outside this appendix.
\end{proof}

\paragraph{Worked example.}
With $C=0.7$ bits, $H(X)=1$ bit, $\kappa_d=0.95$ (estimated on battery
SCADA data), the bridge gives $\bar w\le 0.95\times 0.70=0.665$.  An
episode with $\bar w=0.72$ would violate the bridge and signal
miscalibration.

\paragraph{Scope boundary / non-claim.}
The active audit treats L2 as a domain-scoped proxy relation.  It is not
currently defended as a theorem deriving the heuristic OQE score from
mutual information.

%% ====================================================================
%%  C.16  L3 — Critical Capacity Theorem
%% ====================================================================
\section{L3 — Critical Capacity Theorem}\label{sec:proof-L3}

\begin{theorem}[L3: Critical Capacity]
For any target violation rate $\varepsilon<\alpha$, define the stylized
critical capacity
\[
  C^{*}_d\;=\;\frac{H(X)\,(1-\varepsilon/\alpha)}{\kappa_d}.
\]
Then any certification argument that uses the L2 proxy bridge together
with the executable T3-style upper envelope requires $C\ge C^{*}_d$.
\end{theorem}

\begin{proof}
The executable upper envelope gives
$\mathrm{TSVR}\le\alpha(1-\bar w)$, so demanding
$\mathrm{TSVR}\le\varepsilon$ forces
$\bar w\ge 1-\varepsilon/\alpha$.  Under the L2 proxy bridge,
$\bar w\le\kappa_d\,C/H(X)$, hence
\[
  1-\varepsilon/\alpha\;\le\;\kappa_d\,\frac{C}{H(X)}.
\]
Rearranging yields $C\ge C^{*}_d$.  Thus $C^{*}_d$ is the threshold
required by the stylized bridge-and-achievability argument.
\end{proof}

\paragraph{Worked example.}
$\alpha=0.10$, $\varepsilon=0.02$, $\kappa_d=1$, $H(X)=1$ bit.
$C^{*}_d=1\times(1-0.02/0.10)/1=0.80$ bits.
Within the stylized bridge argument, any channel with $C<0.80$ bits
fails to support a $2\%$ TSVR certificate.

\paragraph{Scope boundary / non-claim.}
L3 is not presented as an independent universal impossibility theorem.
It is an algebraic threshold derived from the L2 proxy and the T3-style
upper envelope.

%% ====================================================================
%%  C.17  L4 — Achievability-Converse Sandwich
%% ====================================================================
\section{L4 — Achievability-Converse Sandwich}\label{sec:proof-L4}

\begin{theorem}[L4: Achievability-Converse Sandwich]
For the binary safe/unsafe state model:
\[
  \frac{\alpha}{K}\,(1-\bar w)
  \;\le\;\mathrm{TSVR}^{*}
  \;\le\;\alpha\,(1-\bar w),
\]
where the upper side is the executable T3-style envelope and the lower
side is the stylized L1/L2 proxy lower envelope.  The constant $K$
captures the surrogate gap and does not grow with $T$.
\end{theorem}

\begin{proof}
\textbf{Upper side (executable achievability).}
The defended T3-style envelope gives
$\mathrm{TSVR}\le\alpha(1-\bar w)$.

\textbf{Lower side (stylized proxy).}
L1 contributes the surrogate lower envelope in capacity space and L2
maps that lower envelope into reliability space through the proxy
constant $\kappa_d$.  Writing $K=1/\kappa_d$ yields the displayed lower
side
\[
  \frac{\alpha}{K}(1-\bar w).
\]
Combining the executable upper side with the stylized lower side gives
the displayed sandwich.
\end{proof}

\paragraph{Worked example.}
$\bar w=0.85$, $\alpha=0.10$, $K=2$.
Lower: $(0.10/2)(1-0.85)=0.0075$.
Upper: $0.10\times(1-0.85)=0.015$.
The stylized law family spans $[0.75\%,\,1.5\%]$.  This interval is a
proxy comparison band, not a sharp theorem about the locked battery
release.

\paragraph{Scope boundary / non-claim.}
The active audit treats L4 as an executable upper bound plus a stylized
proxy lower side.  It is \emph{not} promoted as a closed optimality
characterization.

%% ====================================================================
%%  C.18  T11\_Byzantine — Byzantine-Tolerant OQE Bound
%% ====================================================================
\section{T11\_Byzantine --- Byzantine-Tolerant OQE
Bound}\label{sec:proof-T11-byz}

\begin{theorem}[T11\_Byzantine: Byzantine Trimmed-Mean Bound]
Given a window of $W$ sensor readings with at most fraction $f<1/3$
Byzantine (adversarial) sensors, the symmetric trimmed mean with trim
fraction $\beta\ge f$ on each side and under the explicit
\emph{tail-contamination model} in which every Byzantine reading lies in one
of the trimmed tails satisfies
\[
  |\hat\mu_{\mathrm{trim}}-\mu_{\mathrm{true}}|
  \;\le\;\frac{2\sigma_{\mathrm{honest}}}{\sqrt{W(1-2f)}}
\]
with probability $\ge 1-2e^{-2}\approx 0.729$.
\end{theorem}

\begin{proof}
Under the stated tail-contamination model, trimming $\beta W$ readings from
each tail removes all adversarial values because $\beta\ge f$ on each side.
The remaining
\[
  W_{\mathrm{eff}} = W(1-2f)
\]
readings are therefore honest, and their dispersion is controlled by
$\sigma_{\mathrm{honest}}$.

By Hoeffding's inequality applied to $W_{\mathrm{eff}}$ i.i.d.\
sub-Gaussian readings with parameter $\sigma_{\mathrm{honest}}$:
\[
  \Pr\!\bigl[|\hat\mu_{\mathrm{trim}}-\mu_{\mathrm{true}}|
  > t\bigr]
  \;\le\;2\exp\!\Bigl(-\frac{W_{\mathrm{eff}}\,t^{2}}{2\sigma_{\mathrm{honest}}^{2}}\Bigr).
\]
Setting
\[
  t=\frac{2\sigma_{\mathrm{honest}}}{\sqrt{W_{\mathrm{eff}}}}
\]
gives
\[
  \Pr\!\bigl[|\hat\mu_{\mathrm{trim}}-\mu_{\mathrm{true}}| > t\bigr]
  \le
  2e^{-2}.
\]
Hence the stated width holds with success probability at least
$1-2e^{-2}\approx 0.729$.  The threshold $f<1/3$ ensures
$W_{\mathrm{eff}}>W/3>0$, so the trimmed window is non-empty.
\end{proof}

\paragraph{Worked example.}
$W=30$ readings, $f=0.10$ (3~adversarial), $\sigma_{\mathrm{honest}}=2.0\,\text{MW}$.
$W_{\mathrm{eff}}=30\times 0.80=24$.
Bound: $4.0/\sqrt{24}=0.816\,\text{MW}$.
The trimmed mean is within $0.82\,\text{MW}$ of the true value with
$\ge 73\%$ confidence even under adversarial corruption.

%% ====================================================================
%%  C.19  T\_stale\_decay — Stale-Decay Graceful Degradation
%% ====================================================================
\section{T\_stale\_decay --- Stale-Decay Graceful
Degradation}\label{sec:proof-T-stale}

\begin{theorem}[T\_stale\_decay: Stale-Decay Schedule]
Under stale telemetry persisting for $k>\tau_{\max}$ steps (Assumption
A6), the OQE reliability weight decays geometrically:
\[
  w_t(k)\;=\;w_0\,\gamma^{k-\tau_{\max}},\qquad k>\tau_{\max},
\]
where $\gamma\in(0,1)$ is the decay rate.  The weight reaches the
floor $\varepsilon$ at step
\[
  N_\varepsilon\;=\;\tau_{\max}+\Bigl\lceil
  \frac{\ln(\varepsilon/w_0)}{\ln\gamma}\Bigr\rceil.
\]
Moreover, the episode TSVR under stale decay satisfies
$\mathrm{TSVR}\le\alpha\bigl(1-\bar w_{\mathrm{stale}}\bigr)$ where
$\bar w_{\mathrm{stale}}$ accounts for the geometric series.
\end{theorem}

\begin{proof}
During $k\le\tau_{\max}$, the telemetry is within the detection lag
bound (A6) and $w_t=w_0$.  For $k>\tau_{\max}$, each additional step of
staleness multiplies the previous weight by $\gamma$:
$w_t(k)=w_0\gamma^{k-\tau_{\max}}$ by induction.

Setting $w_0\gamma^{N-\tau_{\max}}=\varepsilon$ and solving:
$N-\tau_{\max}=\ln(\varepsilon/w_0)/\ln\gamma$.  Taking the ceiling
gives $N_\varepsilon$.

The episode-average reliability over $T$ steps with staleness onset at
step $t_0$ is:
\[
  \bar w\;=\;\frac{1}{T}\Bigl[\sum_{k=1}^{\tau_{\max}}w_0
  +\sum_{k=\tau_{\max}+1}^{T}w_0\gamma^{k-\tau_{\max}}\Bigr]
  \;=\;\frac{w_0}{T}\Bigl[\tau_{\max}
  +\frac{1-\gamma^{T-\tau_{\max}}}{1-\gamma}\Bigr].
\]
The TSVR bound follows from T3 (Core Bound):
$\mathrm{TSVR}\le\alpha(1-\bar w)$.

\textbf{Remark.}  The geometric decay is a \emph{design choice}, not a
derived property of the physical system.  It encodes the principle that
stale observations carry exponentially less information, consistent with
the channel-capacity interpretation (capacity of an erasure channel with
increasing erasure probability).
\end{proof}

\paragraph{Worked example.}
$w_0=0.95$, $\gamma=0.85$, $\tau_{\max}=6$ steps, $\varepsilon=0.05$.
$N_\varepsilon=6+\lceil\ln(0.05/0.95)/\ln(0.85)\rceil
=6+\lceil-2.944/(-0.1625)\rceil=6+19=25$ steps.
After 25 steps of persistent staleness the system enters safe-landing
mode ($w_t\le 0.05$).

%% ====================================================================
%%  C.20  T\_minimax — Tight OASG Minimax Tradeoff
%% ====================================================================
\section{T\_minimax --- Tight OASG Minimax
Tradeoff}\label{sec:proof-T-minimax}

\begin{theorem}[T\_minimax: Minimax OASG Tradeoff]
Under degraded observation with mean reliability $\bar w$, the current
extension surface combines a stylized lower envelope with the defended
T3-style upper envelope:
\[
  \frac{\alpha}{K}\sum_{t=1}^{T}(1-w_t)
  \;\le\;\mathbb{E}[V_T]
  \;\le\;\alpha\sum_{t=1}^{T}(1-w_t),
\]
where $K=2$ for binary state spaces and $V_T=\sum_{t=1}^{T}Z_t$ is the
cumulative violation count.
\end{theorem}

\begin{proof}
\textbf{Upper bound.}  By T3 (Core Bound), DC3S ensures
$\Pr[Z_t=1\mid\mathcal{H}_t]\le\alpha(1-w_t)$, so linearity of
expectation gives
$\mathbb{E}[V_T]\le\alpha\sum_t(1-w_t)$.

\textbf{Lower bound.}  The extension surface adopts the stylized L1/L2
proxy lower envelope $r_t\ge(\alpha/K)(1-w_t)$.  This follows because:
\begin{enumerate}
\item L1 gives $D^{*}(C_t)\ge\alpha(1-C_t/H(X))$ where $C_t$ is the
  effective per-step channel capacity.
\item L2 translates: $1-w_t\le 1-\kappa_d C_t/H(X)$, i.e.,
  $1-C_t/H(X)\ge(1-w_t)/\kappa_d$.
\item Substituting: $r_t\ge\alpha(1-w_t)/\kappa_d$.  With
  $K=1/\kappa_d$ (which equals $2$ for BSC at worst-case crossover):
  $r_t\ge(\alpha/K)(1-w_t)$.
\end{enumerate}
Summing over $T$ steps and taking expectations completes the proxy
sandwich.
\end{proof}

\paragraph{Worked example.}
$T=100$ steps, $\bar w=0.90$, $\alpha=0.10$, $K=2$.
Lower: $(0.10/2)\times 100\times 0.10=0.50$ expected violations.
Upper: $0.10\times 100\times 0.10=1.00$ expected violations.
The resulting interval is a stylized comparison band.  It is not a
sharp minimax certificate for the locked deployment artifacts.

\paragraph{Scope boundary / non-claim.}
T\textsubscript{minimax} remains a scoped extension theorem because its
lower side inherits the open L1/L2 proxy bridge.

%% ====================================================================
%%  C.21  T\_sensor — Information-Theoretic Sensor Quality Converse
%% ====================================================================
\section{T\_sensor\_converse --- Sensor Quality
Converse}\label{sec:proof-T-sensor}

\begin{theorem}[T\_sensor\_converse: Sensor Quality Converse]
For any controller $\pi$ and target $\mathrm{TSVR}(\pi)\le\varepsilon$:
\[
  \bar w\;\ge\;1-\frac{\varepsilon}{\alpha}.
\]
This threshold is sufficient by the executable T3-style envelope and
necessary only inside the stylized L2/L3 bridge model.
\end{theorem}

\begin{proof}
\textbf{Stylized necessity.}
Under the L2/L3 proxy bridge, $\bar w<1-\varepsilon/\alpha$ means the
stylized threshold is not met.

\textbf{Executable sufficiency.}
If $\bar w\ge 1-\varepsilon/\alpha$, then
$\alpha(1-\bar w)\le\varepsilon$, and DC3S achieves
$\mathrm{TSVR}\le\alpha(1-\bar w)\le\varepsilon$.
\end{proof}

\paragraph{Worked example.}
$\alpha=0.10$, $\varepsilon=0.02$.
$\bar w\ge 1-0.02/0.10=0.80$.
To guarantee $\le 2\%$ violations, mean sensor reliability must be at
least $80\%$ according to the stylized threshold calculator.

\paragraph{Scope boundary / non-claim.}
The active audit does not treat T\textsubscript{sensor} as a fully
defended universal converse for all controllers; it is the inverse of
the executable upper envelope plus the scoped L2/L3 bridge.

%% ====================================================================
%%  C.22  T\_trajectory\_PAC — Distribution-Free Trajectory Certificate
%% ====================================================================
\section{T\_trajectory\_PAC --- Distribution-Free Trajectory Safety
Certificate}\label{sec:proof-T-trajectory-PAC}

\begin{theorem}[T\_trajectory\_PAC: PAC Trajectory Certificate]
Under A1, A4, A5, A9, and the implemented Bonferroni/union-bound aggregation
over the horizon, for horizon $H$, calibration size $n_{\mathrm{cal}}$, and
confidence target $1-\delta$:
\[
  \Pr\!\bigl[\text{all }H\text{ steps safe}\bigr]
  \;\ge\;1-H\alpha(1-\bar w)-\varepsilon_{\mathrm{fs}}-\delta/2,
\]
where the implemented finite-sample correction is
\[
  \varepsilon_{\mathrm{fs}}
  =
  \sqrt{\frac{\ln(4/\delta)}{2n_{\mathrm{eff}}}},
  \qquad
  n_{\mathrm{eff}}
  =
  \Bigl\lfloor n_{\mathrm{cal}} \min_{t\le H} w_t \Bigr\rfloor.
\]
The reliability-deflated $n_{\mathrm{eff}}$ is a conservative executable
calibration rule used by the helper, not a claim of optimal split-conformal
sample complexity.

The maximum trajectory horizon certified at confidence $1-\delta$ by this
bound is
\[
  H_{\max}
  =
  \max\!\left\{
    0,\;
    \Bigl\lfloor
      \frac{\delta/2-\varepsilon_{\mathrm{fs}}}{\alpha(1-\bar w)}
    \Bigr\rfloor
  \right\}.
\]
\end{theorem}

\begin{proof}
\textbf{Step 1 (per-step risk).}
By T3, $\Pr[Z_t=1\mid\mathcal{H}_t]\le\alpha(1-w_t)$.

\textbf{Step 2 (trajectory aggregation by union bound).}
Let $E_t=\{Z_t=1\}$ denote a stepwise true-state violation.  Then
\[
  \Pr\!\Bigl[\bigcup_{t=1}^{H} E_t\Bigr]
  \;\le\;\sum_{t=1}^{H}\Pr[E_t]
  \;\le\;\sum_{t=1}^{H}\alpha(1-w_t)
  \;\le\;H\alpha(1-\bar w).
\]

\textbf{Step 3 (finite-sample correction).}
The conformal quantile $\hat q$ estimated from $n_{\mathrm{cal}}$ points
satisfies (Vovk, 2005):
$\Pr[\hat q<q_{\mathrm{true}}]\le\varepsilon_{\mathrm{fs}}$.
The executable helper conservatively deflates the effective calibration size by
the worst reliability over the certified prefix:
\[
  n_{\mathrm{eff}}
  =
  \Bigl\lfloor n_{\mathrm{cal}} \min_{t\le H} w_t \Bigr\rfloor.
\]

\textbf{Step 4 (combining).}
\begin{align*}
  &\Pr[\exists\,t\in[H]:\text{violation}]\\
  &\quad\le\underbrace{H\alpha(1-\bar w)}_{\text{risk budget}}
  +\underbrace{\varepsilon_{\mathrm{fs}}}_{\text{calibration error}}
  +\underbrace{\delta/2}_{\text{exit-time budget}}
\end{align*}
\[
  \Rightarrow\;\Pr[\text{all safe}]\ge 1-H\alpha(1-\bar w)
  -\varepsilon_{\mathrm{fs}}-\delta/2.
\]
To certify the trajectory at confidence $1-\delta$, it is sufficient that
\[
  1-H\alpha(1-\bar w)-\varepsilon_{\mathrm{fs}}-\delta/2
  \ge 1-\delta.
\]
Rearranging gives
\[
  H\alpha(1-\bar w)\le \delta/2-\varepsilon_{\mathrm{fs}}.
\]
If the right-hand side is negative, the bound certifies no positive horizon.
Otherwise flooring the quotient gives the displayed $H_{\max}$.
\end{proof}

\paragraph{Worked example.}
$\bar w=0.90$, $\alpha=0.10$, $n_{\mathrm{cal}}=300$, $\delta=0.05$,
$\min_t w_t=0.60$.
$n_{\mathrm{eff}}=\lfloor 300\times 0.60\rfloor=180$.
$\varepsilon_{\mathrm{fs}}=\sqrt{\ln(80)/(360)}=\sqrt{4.382/360}=0.110$.
Because $\delta/2-\varepsilon_{\mathrm{fs}}=0.025-0.110=-0.085<0$, the
certificate is vacuous at the $95\%$ confidence target and
\[
  H_{\max}=0.
\]
With this calibration size and reliability floor, the implemented helper
honestly reports that no positive trajectory horizon is certified at the
requested confidence level.

\paragraph{Scope boundary / non-claim.}
The defended executable surface for T\textsubscript{PAC} is the
Bonferroni/union-bound certificate above.  No separate
Ville/supermartingale strengthening is claimed in the live repo.
