\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsthm, booktabs, geometry}
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}[section]
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\renewcommand{\qedsymbol}{$\blacksquare$}

\title{Appendix C: Flagship Proofs for ORIUS}
\author{Pratik Niroula}
\date{March 2026}

\begin{document}
\maketitle

\paragraph{Scope Note.}
This standalone article is the reviewer-facing integrated proof packet for
ORIUS and the registry-canonical flagship-core execution draft. It presents the
battery witness ladder T1--T8 first, then the defended ORIUS-wide extension
surfaces T11 and T\_trajectory\_PAC.  The active theorem registry remains
authoritative for defense tiers: T3 is presented through the flagship T3a
theorem plus the supporting T3b corollary, T5 remains definition-level within
the ladder, and T8 remains a supporting sequence-level result even though it is
included here for reviewer continuity. The full integrated proof source remains
\texttt{appendices/app\_c\_full\_proofs.tex}.

\section*{C.1 Strict Topological Notation and Dynamics}
Every proof below uses the explicit Assumption Register in Appendix B together
with the active flagship theorem surfaces only.

\begin{table}[h]
\centering
\begin{tabular}{@{}llp{9.5cm}@{}}
\toprule
Symbol & Type & Meaning \\
\midrule
$x_t$ & $\R$ & Latent true state at step $t$. \\
$\hat{x}_t$ & $\R$ & Observed state at step $t$. \\
$U_t$ & Set & Observation-consistent / reliability-adjusted uncertainty set at step $t$. \\
$\mathcal{A}_t$ & Set & Tightened battery safe-action set used by the DC3S shield. \\
$\mathcal{A}_t^{\mathrm{safe}}(U_t)$ & Set & Typed safe-action set sound for the uncertainty set $U_t$. \\
$\Delta(a_t)$ & $\R$ & One-step state increment induced by action $a_t$. \\
$\epsilon_t$ & $\R$ & One-step model residual with $|\epsilon_t| \le \epsilon_{\mathrm{model}}$ under A1. \\
$\mathcal{S}$ & Set & Generic domain safe set. \\
$[s_{\min}, s_{\max}]$ & Interval & Battery safe interval. \\
$\mathcal{H}_t$ & $\sigma$-algebra & Causal observation history up to step $t$. \\
$Z_t$ & $\{0,1\}$ & True-state violation indicator at step $t$. \\
$V$ & $\N$ & Total true-state violation count across the horizon. \\
$w_t$ & $[0,1]$ & Reliability score at step $t$. \\
$\bar w$ & $[0,1]$ & Episode-average reliability $T^{-1}\sum_{t=1}^T w_t$. \\
$\delta_{\mathrm{bnd},t}$ & $\R_+$ & Margin from the issued certificate interval to the nearest SOC boundary. \\
$\tau_t$ & $\N$ & Certifiable battery validity horizon or lower-bound horizon. \\
$H$ & $\N$ & Trajectory horizon for the PAC certificate. \\
$n_{\mathrm{cal}}$ & $\N$ & Calibration sample size. \\
$n_{\mathrm{eff}}$ & $\N$ & Effective reliability-deflated calibration size. \\
$\varepsilon_{\mathrm{fs}}$ & $\R_+$ & Finite-sample conformal correction term. \\
$H_{\max}$ & $\N$ & Maximum certifiable safe horizon under the trajectory certificate. \\
\bottomrule
\end{tabular}
\end{table}

\newpage

\section*{C.2 T1 --- OASG Existence}

\begin{theorem}[T1: OASG Existence]\label{thm:t1}
\textbf{Assumptions.} A1, A2, A4, A11, A12. \\
\textbf{Statement.} For any dropout rate $d>0$ and any optimal arbitrage
controller, there exists a time step $t^*$ such that
\[
  \phi_{\mathrm{obs}}(a_{t^*}) = 1
  \qquad\text{and}\qquad
  \phi_{\mathrm{true}}(a_{t^*}) = 0.
\]
That is, the released action is safe in observation but unsafe in truth.
\end{theorem}

\begin{proof}
By A11, the battery witness controller can be driven to a boundary-sensitive
state under admissible demand, so there exists a step where the true SOC lies
within a boundary distance $\delta_{\mathrm{bnd}}$ of the constraint surface.
By A2 and the standing dropout rate $d>0$, degraded telemetry admits a
non-trivial observation gap on an admissible fault window. A12 keeps the fault
process conditionally independent of the controller action stream given the
history, so the boundary-sensitive witness step and the admissible telemetry
fault can co-occur on the same realized history.

Choose such a step $t^*$. The controller releases the action using the observed
state and therefore treats that action as feasible on the observed side of the
boundary. The true state, however, is displaced by the admissible observation
gap beyond the boundary-facing side of the safe set, so the same released
action crosses the true constraint surface. A4 supplies the shared one-step
dynamics interpretation needed to compare the observed-state and true-state
outcomes for that released action. Hence
\[
  \phi_{\mathrm{obs}}(a_{t^*}) = 1
  \qquad\text{and}\qquad
  \phi_{\mathrm{true}}(a_{t^*}) = 0,
\]
which is exactly the battery witness-row OASG phenomenon. \qedhere
\end{proof}

\section*{C.3 T2 --- Safety Preservation}

\begin{theorem}[T2: Safety Preservation]\label{thm:t2}
\textbf{Assumptions.} A1, A2, A3, A4, A5, A7. \\
\textbf{Statement.} If the DC3S shield selects $a_t \in \mathcal{A}_t$ and
the current true state satisfies $x_t \in U_t$, then the released action is
one-step safe:
\[
  x_t \in U_t \;\text{ and }\; a_t \in \mathcal{A}_t
  \quad\Longrightarrow\quad
  x_{t+1} \in [s_{\min}, s_{\max}].
\]
No episode-level expectation claim is made in this theorem.
\end{theorem}

\begin{proof}
Fix a step $t$ and assume the theorem hypotheses. Because $x_t \in U_t$, the
current true state lies in the observation-consistent battery tube, so
\[
  |x_t - \hat{x}_t| \le m_t.
\]
Because $a_t \in \mathcal{A}_t$, the shielded observed-state prediction is
required to stay inside the absorbed tightened interval:
\[
  s_{\min} + m_t + \epsilon_{\mathrm{model}}
  \le
  \hat{x}_t + \Delta(a_t)
  \le
  s_{\max} - m_t - \epsilon_{\mathrm{model}}.
\]

Now write the true successor using A4:
\[
  x_{t+1} = x_t + \Delta(a_t) + \epsilon_t
  =
  \hat{x}_t + \Delta(a_t) + (x_t-\hat{x}_t) + \epsilon_t.
\]
Using A1 and the tube bound gives the lower estimate
\[
  x_{t+1}
  \ge
  (\hat{x}_t + \Delta(a_t)) - |x_t-\hat{x}_t| - |\epsilon_t|
  \ge
  (s_{\min} + m_t + \epsilon_{\mathrm{model}}) - m_t - \epsilon_{\mathrm{model}}
  =
  s_{\min}.
\]
The same substitution yields the upper estimate
\[
  x_{t+1}
  \le
  (\hat{x}_t + \Delta(a_t)) + |x_t-\hat{x}_t| + |\epsilon_t|
  \le
  (s_{\max} - m_t - \epsilon_{\mathrm{model}}) + m_t + \epsilon_{\mathrm{model}}
  =
  s_{\max}.
\]
Therefore
\[
  s_{\min} \le x_{t+1} \le s_{\max},
\]
so the released action is one-step safe. The proof uses only current-step
information and therefore preserves the causal certificate discipline required
by A7. \qedhere
\end{proof}

\section*{C.4 T3a --- ORIUS Core Envelope Derivation}

\begin{lemma}[Aggregation under a predictable risk budget]\label{lem:t3a-budget}
If a predictable nonnegative sequence $r_t$ satisfies
\[
  \Prob[Z_t = 1 \mid \mathcal{H}_t] \le r_t
  \qquad\text{for all } t,
\]
then
\[
  \E[V] \le \sum_{t=1}^T r_t.
\]
\end{lemma}

\begin{proof}
Because $Z_t$ is an indicator,
\[
  \E[Z_t \mid \mathcal{H}_t]
  =
  \Prob[Z_t=1 \mid \mathcal{H}_t]
  \le r_t.
\]
Taking unconditional expectations preserves the inequality, and summing across
the horizon yields $\E[V] \le \sum_{t=1}^T r_t$. No temporal independence
assumption is required. \qedhere
\end{proof}

\begin{theorem}[ORIUS Core Bound]\label{thm:t3a}
\textbf{Assumptions.} A1, A2, A4, A5, A6, A7, A9. \\
\textbf{Statement.} Under the battery witness-row predictable one-step risk
budget
\[
  \Prob[Z_t=1\mid\mathcal{H}_t] \le \alpha(1-w_t),
\]
the expected total violation count under the DC3S shield satisfies
\[
  \E[V] \le \alpha(1-\bar w)T,
\]
where $\bar w = T^{-1}\sum_{t=1}^T w_t$ is the episode-average reliability.
\end{theorem}

\begin{proof}
Apply Lemma~\ref{lem:t3a-budget} with $r_t=\alpha(1-w_t)$. Then
\begin{align*}
  \E[V]
  &\le \sum_{t=1}^T \alpha(1-w_t) \\
  &= \alpha \sum_{t=1}^T (1-w_t) \\
  &= \alpha T \left(1 - \frac{1}{T}\sum_{t=1}^T w_t\right) \\
  &= \alpha(1-\bar w)T.
\end{align*}
Thus the flagship T3a surface is an aggregation argument over an explicit
predictable risk-envelope contract rather than a standalone corollary of
marginal conformal coverage. \qedhere
\end{proof}

\paragraph{Registry Note.}
The active theorem registry treats the theorem above as flagship-defended
$T3a$. The supporting corollary $T3b$ remains in
\texttt{appendices/app\_c\_full\_proofs.tex} and is not reproduced in this
flagship-only appendix.

\section*{C.5 T4 --- No Free Safety}

\begin{theorem}[T4: No Free Safety]\label{thm:t4}
\textbf{Assumptions.} A1, A2, A4, A11. \\
\textbf{Statement.} For any fixed-margin quality-ignorant controller $\pi$ and
any dropout rate $d>0$, there exists an admissible fault sequence and a time
$t^\star$ such that
\[
  \phi_{\mathrm{obs}}(\pi(\hat{x}_{t^\star})) = 1
  \qquad\text{and}\qquad
  \phi_{\mathrm{true}}(\pi(\hat{x}_{t^\star})) = 0.
\]
Equivalently, the controller incurs at least one true-state violation on the
constructed admissible sequence.
\end{theorem}

\begin{proof}
Fix a quality-ignorant controller $\pi$. By definition, its effective safety
margin does not depend on the reliability score, so there exists a constant
$m^\pi \ge 0$ that upper-bounds the observation buffer implicitly used by the
controller. By A11, the battery witness controller can be driven through a
boundary-sensitive baiting phase in which the state approaches the constraint
surface while telemetry remains nominal.

Because $d>0$, A2 admits an admissible degraded-observation fault window with a
non-trivial observation gap. Choose a step $t^\star$ in that window where the
accumulated observation error exceeds the controller's fixed margin plus the
boundary distance available on the baited state. At that step the observed
state still appears feasible relative to the controller's fixed margin, but the
true state has shifted across the boundary-facing side of the admissible set by
more than the controller can compensate. A4 again supplies the shared
one-step battery dynamics interpretation for the released action.

Therefore the action $\pi(\hat{x}_{t^\star})$ remains observed-state legal
while creating a true-state violation. This yields an observation-action safety
gap and hence at least one true-state violation on the constructed admissible
sequence. The theorem is intentionally class-scoped to fixed-margin,
quality-ignorant controllers on the battery witness row. \qedhere
\end{proof}

\section*{C.6 T5 --- Certificate Validity Horizon}

\begin{theorem}[T5: Certificate Validity Horizon]\label{thm:t5}
\textbf{Assumptions.} A4, A7, A9. \\
\textbf{Statement.} The battery validity horizon $\tau_t$ is the largest
integer horizon whose forward tube remains inside the SOC constraint interval.
This row is constructive and definition-level inside the integrated ladder.
\end{theorem}

\begin{proof}
Fix the issued battery interval and the certified safe action.  The forward
tube at horizon $\Delta$ is obtained by propagating that action under the
battery one-step dynamics while adding the worst-case sub-Gaussian radius used
by the defended runtime.  The set of admissible horizons is therefore exactly
the set of integers for which the resulting tube stays inside
$[s_{\min}, s_{\max}]$.  The maximal such integer is $\tau_t$, and the
executable helper \texttt{certificate\_validity\_horizon()} computes precisely
that constructive object. \qedhere
\end{proof}

\section*{C.7 T6 --- Certificate Expiration Bound}

\begin{theorem}[T6: Certificate Expiration Bound]\label{thm:t6}
\textbf{Assumptions.} A4, A7, A9, and confidence level $\delta\in(0,1)$. \\
\textbf{Statement.} The battery certificate horizon satisfies the lower bound
\[
  \tau_t \ge
  \left\lfloor
    \frac{\delta_{\mathrm{bnd},t}^2}
    {2\sigma_d^2 \log(2/\delta)}
  \right\rfloor.
\]
\end{theorem}

\begin{proof}
For a $\sigma_d$-sub-Gaussian disturbance, the reflection-principle /
first-passage tail bound gives
\[
  \Prob(|S_\Delta| \ge \delta_{\mathrm{bnd},t})
  \le
  2\exp\!\left(
    -\frac{\delta_{\mathrm{bnd},t}^2}{2\sigma_d^2\Delta}
  \right).
\]
Demanding the right-hand side be at most $\delta$ yields
\[
  \Delta \le \frac{\delta_{\mathrm{bnd},t}^2}
  {2\sigma_d^2 \log(2/\delta)}.
\]
Because the validity horizon is integer-valued, flooring gives the claimed
lower bound.  The theorem-facing helper
\texttt{certificate\_expiration\_bound()} implements this exact closed form and
records its explicit side conditions in the returned theorem contract.
\qedhere
\end{proof}

\section*{C.8 T7 --- Feasible Fallback Existence}

\begin{theorem}[T7: Feasible Fallback Existence]\label{thm:t7}
\textbf{Assumptions.} A1, A4, A8. \\
\textbf{Statement.} For any safe battery state, the defended fallback surface
is piecewise:
\[
  \text{either zero dispatch safely holds on the interior,}
\]
\[
  \text{or a boundary-aware safe-landing action safely moves the SOC inward.}
\]
If neither case can be certified under the active constraints, the runtime
fails closed rather than claiming a false fallback theorem.
\end{theorem}

\begin{proof}
If the current SOC has enough interior slack, zero dispatch gives
$x_{t+1}=x_t+\epsilon_t$ with $|\epsilon_t|\le\epsilon_{\mathrm{model}}$, so
the next state remains inside the admissible interval.  This is the hold case.

If the current SOC is safe but too close to a boundary for zero dispatch to be
certified, the safe-landing controller computes a charge or discharge action
that moves the SOC toward the interior safe zone while respecting the active
battery power, ramp, and efficiency constraints.  The theorem-facing validator
checks the corresponding successor interval under the bounded model error and
accepts this case only when the entire interval remains safe.

If neither the hold case nor the safe-landing case is certifiable, the
validator returns infeasible.  The defended theorem therefore proves the
piecewise hold-or-safe-landing battery surface and does not claim a universal
fallback guarantee beyond those executable cases. \qedhere
\end{proof}

\section*{C.9 T8 --- Graceful Degradation Dominance}

\begin{theorem}[T8: Graceful Degradation Dominance]
\textbf{Assumptions.} A1--A5, A8. \\
\textbf{Statement.} For graceful and uncontrolled violation sequences produced
under the same admissible fault trace,
\[
  z_t^{\mathrm{graceful}} \le z_t^{\mathrm{uncontrolled}} \;\; \forall t
  \quad\Longrightarrow\quad
  V^{\mathrm{graceful}} \le V^{\mathrm{uncontrolled}}.
\]
Strict improvement is claimed only when the graceful sequence is strictly
smaller on at least one step.
\end{theorem}

\begin{proof}
Summing the stepwise dominance inequality over the shared admissible fault
trace gives
\[
  \sum_{t=1}^{T} z_t^{\mathrm{graceful}}
  \le
  \sum_{t=1}^{T} z_t^{\mathrm{uncontrolled}},
\]
which is exactly the claimed cumulative-count dominance statement.  Strict
inequality follows if the graceful sequence is strictly smaller on at least one
step.  This is the supporting T8 surface in the integrated ladder: a
paired-sequence comparison and nothing stronger. \qedhere
\end{proof}

\section*{C.10 T11 --- Typed Structural Transfer}

\paragraph{Local T11 Notation.}
In this section, $U_t$ denotes the constructed observation-consistent state set
and $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ denotes the tightened safe-action set
that is sound for that uncertainty set.

\begin{theorem}[Typed structural transfer theorem, T11]\label{thm:t11}
\textbf{Assumptions.} For every step $t$:
\begin{enumerate}
  \item the constructed observation-consistent state set satisfies
    $\Prob[x_t \in U_t \mid \mathcal{H}_t] \ge 1-\alpha$;
  \item the tightened safe-action set is sound for that uncertainty set, in
    the sense that every action in $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ keeps
    the next state inside $\mathcal{S}$ for all $x \in U_t$ and all
    admissible disturbance;
  \item the repair map returns an element of
    $\mathcal{A}_t^{\mathrm{safe}}(U_t)$ whenever that set is non-empty;
  \item when the tightened safe-action set is empty, the fallback action is
    admissible and preserves safety from any current safe state.
\end{enumerate}
\textbf{Statement.} Let $\mathcal{D}$ be a domain instantiation for the
degraded-observation safe-control problem. Then the repaired action produced by
the ORIUS kernel satisfies the same one-step safety-preservation statement as
the reference theorem:
\[
  \Prob[x_{t+1}\in\mathcal{S}\mid \mathcal{H}_t] \ge 1-\alpha.
\]
No broader episode-level claim is inherited unless the new domain separately
supplies an explicit predictable per-step risk budget.
\end{theorem}

\begin{proof}
Fix a step $t$ and condition on the observation history $\mathcal{H}_t$. On the
event
\[
  E_t = \{x_t \in U_t\},
\]
Assumption~(2) implies that every action in
$\mathcal{A}_t^{\mathrm{safe}}(U_t)$ preserves membership in the safe set at
the next step. If the tightened safe-action set is non-empty, then
Assumption~(3) ensures that the repaired action returned by the kernel lies in
that set and hence preserves safety on $E_t$. If the tightened safe-action set
is empty, then Assumption~(4) ensures that the fallback action preserves safety
from any current safe state. Therefore,
\[
  \Prob[x_{t+1}\in\mathcal{S} \mid \mathcal{H}_t]
  \ge
  \Prob[E_t \mid \mathcal{H}_t]
  \ge
  1-\alpha,
\]
which is exactly the forward one-step transfer claim. The converse remains a
separate structural failure proposition and is not part of active T11.
\qedhere
\end{proof}

\section*{C.11 T\_trajectory\_PAC --- PAC Trajectory Certificate}

\begin{theorem}[T\_trajectory\_PAC: PAC Trajectory Certificate]\label{thm:t_trajectory_pac}
\textbf{Assumptions.} A1, A4, A5, A9, and Bonferroni/union-bound aggregation
over the horizon. For horizon $H$, calibration size $n_{\mathrm{cal}}$, and
confidence $1-\delta$:
\[
  \Prob\!\bigl[\text{all }H\text{ steps safe}\bigr]
  \ge
  1-H\alpha(1-\bar w)-\varepsilon_{\mathrm{fs}}-\delta/2,
\]
where
\[
  \varepsilon_{\mathrm{fs}}
  =
  \sqrt{\ln(4/\delta)/(2n_{\mathrm{eff}})}
  \qquad\text{and}\qquad
  n_{\mathrm{eff}} = n_{\mathrm{cal}}\cdot\min_t w_t.
\]
The maximum certifiable horizon is
\[
  H_{\max}
  =
  \left\lfloor
    \frac{1-\varepsilon_{\mathrm{fs}}-\delta/2}{\alpha(1-\bar w)}
  \right\rfloor.
\]
\end{theorem}

\begin{proof}
\textbf{Step 1 (per-step risk).}
By T3a,
\[
  \Prob[Z_t=1\mid\mathcal{H}_t]\le\alpha(1-w_t).
\]

\textbf{Step 2 (trajectory aggregation by union bound).}
Let $E_t=\{Z_t=1\}$ denote a stepwise true-state violation. Then
\[
  \Prob\!\Bigl[\bigcup_{t=1}^{H} E_t\Bigr]
  \le \sum_{t=1}^{H}\Prob[E_t]
  \le \sum_{t=1}^{H}\alpha(1-w_t)
  \le H\alpha(1-\bar w).
\]

\textbf{Step 3 (finite-sample correction).}
The conformal quantile $\hat q$ estimated from $n_{\mathrm{cal}}$ points
satisfies
\[
  \Prob[\hat q<q_{\mathrm{true}}]\le\varepsilon_{\mathrm{fs}}.
\]
We deflate the effective calibration size by $\min_t w_t$ to account for the
worst-case reliability in the trajectory.

\textbf{Step 4 (combining).}
\begin{align*}
  \Prob[\exists\,t\in[H]:\text{violation}]
  &\le
  \underbrace{H\alpha(1-\bar w)}_{\text{risk budget}}
  +
  \underbrace{\varepsilon_{\mathrm{fs}}}_{\text{calibration error}}
  +
  \underbrace{\delta/2}_{\text{exit-time budget}}.
\end{align*}
Hence
\[
  \Prob[\text{all safe}]
  \ge
  1-H\alpha(1-\bar w)-\varepsilon_{\mathrm{fs}}-\delta/2.
\]
Setting the right-hand side nonnegative and solving for $H$ gives the stated
$H_{\max}$ formula. This is the defended Bonferroni/union-bound certificate and
nothing stronger; no martingale or Ville-style strengthening is claimed here.
\qedhere
\end{proof}

\end{document}
