% =============================================================================
% Appendix M: Audited Theorems and Gap Register
% ORIUS/DC3S Battery-Safety Thesis
% =============================================================================
\chapter{Audited Theorems and Gap Register}
\label{app:verified-theorems-gap-audit}
\begingroup
\footnotesize
\sloppy

This appendix is the human-readable companion to the machine-generated
active theorem audit in \texttt{reports/publication/active\_theorem\_audit.*}.
It does \emph{not} treat the live theorem inventory as one flat ladder.
Instead it mirrors the registry-canonical split emitted by
\texttt{defended\_theorem\_core.*}:
\begin{itemize}
  \item \textbf{Flagship defended}: T1, T2, T3a, T4, T6, T7, T11, and T\_trajectory\_PAC.
  \item \textbf{Supporting defended}: T3b, T8, T11\textsubscript{Byz}, and T\textsubscript{stale}.
  \item \textbf{Draft / non-defended}: T5, T9, T10, L1--L4,
  T\textsubscript{minimax}, and T\textsubscript{sensor}.
\end{itemize}
This appendix audits the full live surface, but only the flagship and
supporting tiers belong to the defended June core. The draft tier remains
visible to prevent hidden drift, not to inflate the headline theorem count.

The certificate half-life remains a corollary, and safe-budget monotonicity
remains a proposition. The tracked integrated theorem surface register still
audits the older hard-gated traceability surface, but it is a
\emph{traceability gate}, not a proof-soundness gate.

\section{Theorem Statements}

\subsection{Theorem 1: OASG Existence}
\label{sec:thm-oasg}

\textbf{Statement.} There exist scenarios where an action appears safe in observed state but is unsafe in true physical state. Formally: $\exists$ episodes with degraded telemetry such that $\text{observed-safe}(a \mid x_{\text{obs}}) \neq \text{true-safe}(a \mid x_{\text{true}})$.

\textbf{Assumptions.} A2 (bounded telemetry error), A4 (known dynamics). Fault injection produces non-trivial $x_{\text{obs}} \neq x_{\text{true}}$ separation.

\textbf{Proof status.} Structural: code explicitly separates observed and true state; empirical: violation rates under dropout/drift exceed nominal.

\textbf{Code anchor.} \texttt{src/orius/cpsbench\_iot/}: \texttt{scenarios.py} (fault generation, $x_{\text{obs}}$, $x_{\text{true}}$, \texttt{event\_log}); \texttt{plant.py} (unclamped truth SOC in \texttt{BatteryPlant.step()}); \texttt{runner.py} (true-state violation computation).

\subsection{Theorem 2: One-Step Safety Preservation}
\label{sec:thm-safety-preservation}

\textbf{Statement.} If the repaired action lies inside the tightened safe set
and the current state lies inside the inflated certificate, then the next
battery state remains safe.

\textbf{Assumptions.} A1--A5, A7.

\textbf{Proof status.} Proved as a conditional one-step theorem plus
deterministic runtime guarantee checks.  Coverage frequency is supplied
separately and is not bundled into T2.

\textbf{Code anchor.} \texttt{src/orius/dc3s/guarantee\_checks.py} and
\texttt{src/orius/dc3s/shield.py}.

\subsection{Theorem 3a: ORIUS Core Envelope Derivation}
\label{sec:thm-core-bound}

\textbf{Statement.} The active defended T3a surface is the per-step envelope
\[
  \Prob[Z_t = 1 \mid \mathcal{H}_t] \le \alpha(1-w_t),
\]
under the explicit battery risk-budget contract and the narrowed
reliability-score interpretation.

\textbf{Assumptions.} A1, A2, A4, A5, A6, A7, A9, plus the theorem-local
battery calibration bridge recorded in the active theorem audit.

\textbf{Proof status.} Paper-rigorous flagship theorem with explicit runtime
contract reporting.  The code now emits a theorem-contract summary that checks
the executable envelope algebra and keeps the calibration bridge explicit as a
declared theorem-local contract rather than hiding it inside generic conformal
language.

\textbf{Code anchor.} \texttt{src/orius/universal\_theory/risk\_bounds.py}
(\texttt{build\_t3a\_contract\_summary()}), plus the backward-compatible
wrapper \texttt{src/orius/dc3s/coverage\_theorem.py}.

\subsection{Corollary 3b: ORIUS Core Aggregation Corollary}
\label{sec:thm-core-bound-aggregation}

\textbf{Statement.} Aggregating the predictable per-step budget from T3a gives
the episode envelope
\[
  \mathbb{E}[V] \le \alpha(1-\bar{w})T.
\]

\textbf{Assumptions.} The assumptions of T3a, plus linearity of expectation.

\textbf{Proof status.} Supporting defended corollary.  Once T3a is fixed, the
aggregation is algebraic bookkeeping rather than a new theorem burden.

\textbf{Code anchor.} \texttt{src/orius/universal\_theory/risk\_bounds.py}
(\texttt{compute\_episode\_risk\_bound()}).

\subsection{Theorem 4: No Free Safety}
\label{sec:thm-no-free}

\textbf{Statement.} Within the fixed-margin, quality-ignorant controller
class, there exists an admissible degraded-observation sequence that produces
an observation-action safety gap and hence a true-state violation.

\textbf{Assumptions.} A3 (feasible safe repair exists), A8 (admissible fallback policy). Quality-ignorant baselines lack reliability-aware inflation.

\textbf{Proof status.} Constructive witness theorem on the battery reference
surface, plus empirical ablation support.

\textbf{Code anchor.} Supporting evidence in \texttt{baselines.py},
\texttt{scenarios.py}, and \texttt{runner.py}.

\subsection{Definition 5: Certificate Validity Horizon}
\label{sec:thm-cert-horizon}

\textbf{Statement.} T5 is kept as the constructive maximal forward-tube
containment definition: the validity horizon is the largest integer horizon
whose forward tube remains inside the battery SOC bounds.

\textbf{Assumptions.} A1, A2, A4--A7.

\textbf{Proof status.} Explicit definition, not part of the defended theorem
count.  The helper computes the maximal containment horizon directly and no
probabilistic future-step theorem is claimed at this surface.

\textbf{Code anchor.} \texttt{src/orius/dc3s/temporal\_theorems.py}
(\texttt{forward\_tube()}, \texttt{certificate\_validity\_horizon()});
registered as T5 in \texttt{THEOREM\_REGISTER}
(\texttt{src/orius/dc3s/theoretical\_guarantees.py}).
Runtime expiry logic consumes this quantity, but the active theorem audit does
not count T5 as a defended theorem.

\subsection{Theorem 6: Certificate Expiration Bound}
\label{sec:thm-cert-expiration}

\textbf{Statement.} The battery certificate horizon satisfies the canonical
delta-aware lower bound
\[
  \tau_t \ge
  \left\lfloor
    \frac{\delta_{\mathrm{bnd},t}^2}
    {2\sigma_d^2\log(2/\delta)}
  \right\rfloor.
\]

\textbf{Assumptions.} A4--A7.

\textbf{Proof status.} Flagship defended closed-form theorem with
machine-checkable helper linkage.  The theorem-facing API requires
\texttt{delta} explicitly, records the first-passage side conditions, and
marks legacy no-\texttt{delta} semantics as non-defended.

\textbf{Code anchor.} \texttt{src/orius/universal\_theory/battery\_instantiation.py}
(\texttt{certificate\_expiration\_bound()}), which now emits theorem metadata
and a fail-closed T6 contract summary.

\subsection{Theorem 7: Feasible Fallback Existence}
\label{sec:thm-fallback}

\textbf{Statement.} The defended T7 surface is piecewise: zero dispatch
preserves safety on the interior, a boundary-aware safe-landing action
preserves safety near the boundary, and the runtime fails closed when neither
case can be certified.

\textbf{Assumptions.} A1, A4, A8.

\textbf{Proof status.} Flagship defended battery fallback theorem with an
explicit piecewise executable witness.  The promoted surface is battery-only
and does not claim cross-domain transfer on its own.

\textbf{Code anchor.} \texttt{src/orius/dc3s/temporal\_theorems.py}
(\texttt{zero\_dispatch\_fallback()}, \texttt{certify\_fallback\_existence()}).

\subsection{Theorem 8: Graceful Degradation Dominance}
\label{sec:thm-graceful-dominance}

\textbf{Statement.} The defended T8 surface is the paired sequence-level
comparison: for graceful and uncontrolled violation sequences produced under
the same admissible fault trace,
\[
  z_t^{\mathrm{graceful}} \le z_t^{\mathrm{uncontrolled}} \;\; \forall t
  \quad\Longrightarrow\quad
  V^{\mathrm{graceful}} \le V^{\mathrm{uncontrolled}}.
\]
Strict improvement is claimed only when the graceful sequence is strictly
smaller on at least one step.

\textbf{Assumptions.} A1--A5, A8.

\textbf{Proof status.} Supporting defended comparison theorem at the
sequence-comparison level only.  The stronger absorbing-landing interpretation
remains supporting evidence rather than a promoted theorem.

\textbf{Code anchor.} \texttt{src/orius/universal\_theory/battery\_instantiation.py}
(\texttt{evaluate\_graceful\_degradation\_dominance()}), with
\texttt{safe\_landing} kept as supporting implementation evidence only.

\subsection{Theorem 9: Universal Impossibility}
\label{sec:thm-universal-impossibility}

\textbf{Statement.} For any quality-ignorant controller~$\pi$ and any
$\varphi$-mixing fault process with mean degradation rate~$d > 0$,
$\mathbb{E}[V_T(\pi)] \ge c\,d\,T$ for some constant $c > 0$.

\textbf{Assumptions.} A4 (boundary reachability), A6$'$ ($\varphi$-mixing
fault process).

\textbf{Proof status.} Structural impossibility theorem under explicit
$\varphi$-mixing and boundary-reachability assumptions; battery T4 is the
reference witness for the per-window violation probability and the witness
constant remains theorem-local.

\textbf{Code anchor.} No dedicated module; the battery
\texttt{cpsbench\_iot/runner.py} provides the reference witness;
the proof is summarized in Section~\ref{sec:thm-universal-impossibility}.

\subsection{Theorem 10: Boundary-Indistinguishability Lower Bound}
\label{sec:thm-info-lower-bound}

\textbf{Statement.} Under the explicit boundary-mass and
indistinguishability hypotheses of T10, every controller~$\pi$ satisfies the
scoped lower bound
$\mathbb{E}[V_T(\pi)] \ge \tfrac{1}{4}\,\alpha(1-\bar{w})T$.

\textbf{Assumptions.} A1, A2, A5$'$ (calibration regularity), conformal
coverage condition.

\textbf{Proof status.} Narrowed stylized lower-bound surface under extra
calibration/indistinguishability assumptions.  The active audit records that
the one-sided Le~Cam step remains open, so the code witness is treated as a
scoped lower-bound calculator rather than a sharp frontier theorem.

\textbf{Code anchor.} \texttt{src/orius/universal/reliability\_safety\_frontier.py}
(\texttt{compute\_frontier}, \texttt{is\_dc3s\_optimal}); 25 unit tests
in \texttt{tests/test\_reliability\_frontier.py}.

\subsection{Theorem 11: Typed Structural Transfer}
\label{sec:thm-universality-completeness}

\textbf{Statement.} Any CPS domain instantiation satisfying the structural
transfer obligations of T11---coverage of the observation-consistent state
set, soundness of the tightened safe-action set, repair membership, and
safe fallback admissibility---inherits the T2--T3 safety-guarantee pattern
without new battery-specific algebra.

\textbf{Assumptions.} The universal adapter contract and the four structural
obligations stated in Section~\ref{sec:thm-universality-completeness}.

\textbf{Proof status.} Forward one-step transfer theorem.  The runtime now
builds the four T11 obligation checks directly from typed artifacts, while the
converse is tracked separately as a structural failure-mode proposition rather
than as part of the defended T11 statement.

\textbf{Code anchor.} \texttt{src/orius/universal\_theory/contracts.py} and
\texttt{src/orius/universal\_theory/kernel.py}, with a compact reference
harness in \texttt{src/orius/universal/contract.py}; focused tests in
\texttt{tests/test\_universal\_contract.py},
\texttt{tests/test\_unification.py}, and
\texttt{tests/test\_reliability\_frontier.py}, plus
\texttt{tests/test\_universal\_theory.py}.

\section{Gap Audit Table}

Table~\ref{tab:theorem-gap-audit} summarises the active T1--T11 audit surface,
with the understanding that only a strict subset belongs to the June defended core.

\begin{table}[htbp]
\centering
\caption{Active T1--T11 theorem audit summary with flagship/supporting/draft posture}
\label{tab:theorem-gap-audit}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llllp{3.2cm}@{}}
\toprule
\textbf{Theorem} & \textbf{Proof status} & \textbf{Empirical validation} & \textbf{Code matches statement} & \textbf{Notes} \\
\midrule
T1 OASG Existence & Structural & Yes & Yes & Observed-vs-true split is explicit in benchmark code and locked evidence \\
T2 One-Step Safety Preservation & Yes (conditional one-step) & Yes & Yes & Shield + guarantee checks + unit tests anchor the theorem; coverage frequency handled separately \\
T3a ORIUS Core Envelope Derivation & Paper-rigorous & Yes & Yes & Runtime theorem-contract summary now separates executable envelope algebra from the declared calibration bridge \\
T3b ORIUS Core Aggregation Corollary & Corollary only & Yes & Yes & Supporting aggregation corollary derived from T3a by linearity of expectation \\
T4 No Free Safety & Constructive witness & Yes & Partial & Battery proof gives an explicit degraded-observation sequence; broader baseline comparison remains empirical \\
T5 Certificate Validity Horizon Definition & Explicit definition & Yes & Yes & Maximal forward-tube containment helper is constructive and excluded from defended theorem counts \\
T6 Certificate Expiration Bound & V2-linked flagship theorem & Yes & Yes & Canonical helper is delta-aware, closed-form, and records explicit first-passage side conditions \\
T7 Feasible Fallback Existence & Piecewise hold-or-safe-landing theorem & Yes & Yes & Runtime certifies interior safe hold, boundary recovery by safe landing, and fail-closed infeasibility \\
T8 Graceful Degradation Dominance & Supporting sequence comparison & Partial & Yes & Defended only as a paired-sequence comparator under a shared fault trace; absorbing landing remains supporting-only \\
\midrule
T9 Universal Impossibility & Structural / scoped witness constant & Yes & Partial & $\varphi$-mixing + Azuma windowing remain explicit assumptions; battery T4 is the reference witness; Section~\ref{sec:thm-universal-impossibility} \\
T10 Boundary-Indistinguishability Lower Bound & Narrowed stylized lower bound & Partial & Partial & Le~Cam step remains one-sided; the executable witness is a scoped lower-bound calculator, not a sharp frontier theorem \\
T11 Typed Structural Transfer & Forward-only one-step transfer & Yes & Yes & Typed runtime artifacts now carry four-obligation summaries; the converse and mini-harness remain separate supporting surfaces \\
\bottomrule
\end{tabular}%
}
\end{table}

\begin{table}[htbp]
\centering
\caption{Audited extension laws and depth-theorem surfaces}
\label{tab:extension-theorem-gap-audit}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llllp{3.1cm}@{}}
\toprule
\textbf{ID} & \textbf{Status} & \textbf{Code surface} & \textbf{Audit posture} & \textbf{Notes} \\
\midrule
L1 & Broken as standalone converse & \texttt{orius\_law.py} & Stylized surrogate only & Linear lower envelope retained for extension calculators only \\
L2 & Broken as theorem of live OQE & \texttt{orius\_law.py}, \texttt{quality.py} & Proxy bridge only & Requires external justification of $\kappa_d$ and the capacity proxy \\
L3 & Has a hole & \texttt{orius\_law.py} & Narrowed threshold calculator & Depends on L2 proxy plus the executable T3 upper envelope \\
L4 & Has a hole & \texttt{orius\_law.py} & Stylized sandwich & Executable upper side plus proxy lower side; not a closed characterization \\
T11\_Byzantine & Plausible but informal & \texttt{theoretical\_guarantees.py} & Bounded extension & Matches the trimmed-mean witness and tests \\
T\_stale\_decay & Plausible but informal & \texttt{theoretical\_guarantees.py} & Bounded extension & Geometric decay is a design schedule, not a derived physical law \\
T\_minimax & Broken as defended minimax theorem & \texttt{theoretical\_guarantees.py} & Stylized lower-envelope witness & Inherits the open L1/L2 proxy bridge \\
T\_sensor\_converse & Broken as universal converse & \texttt{theoretical\_guarantees.py} & Stylized inverse threshold & Necessity only holds inside the L2/L3 proxy model \\
T\_trajectory\_PAC & Plausible but informal & \texttt{risk\_bounds.py} & Defended as Bonferroni certificate & No separate Ville-strengthened theorem is claimed \\
\bottomrule
\end{tabular}%
}
\end{table}

\section{Remaining Claim-Boundary Gaps}

The audit above separates proof completeness from promotion completeness.
The remaining gaps that still bound the dissertation's final claim are
therefore explicit:
\begin{enumerate}
  \item \textbf{Autonomous vehicles.} The current AV row is no longer claimed
    as full autonomy closure. It is promoted only as a bounded nuPlan replay
    row under TTC and predictive-entry-barrier semantics. Remaining AV work is
    closed-loop road deployment evidence, not a contradiction of T1--T11.
  \item \textbf{Future mobility and flight-style rows.} Earlier drafts named
    additional exploratory rows, but the current monograph does not promote
    them. They remain planning surfaces until they supply defended replay,
    runtime, governance, and post-repair evidence under the same gate as the
    active Battery + AV + Healthcare lane.
\end{enumerate}

These are the remaining thesis-level gaps after the monograph theorem pass.
They do not weaken the battery proof witness; they bound how far the transfer
claim is currently allowed to extend.

\section{Locked Empirical Evidence}

Battery theorem verification relies on a mixture of code anchors, tests, and
locked outputs in \texttt{reports/publication/}:
\begin{itemize}
\item \texttt{dc3s\_main\_table\_ci.csv}: violation rates, PICP, scenario breakdown
\item \texttt{reliability\_group\_coverage\_phase3.csv}: coverage by reliability bin
\item \texttt{blackout\_study.csv}: certificate-safe-hold evidence for the temporal extension
\item \texttt{graceful\_degradation\_trace.csv}: behavioral evidence for the landing controller
\item \texttt{claim\_evidence\_matrix.csv}: claim-to-evidence linkage
\end{itemize}

The theorem-to-evidence map is maintained in \texttt{orius-plan/theorem\_to\_evidence\_map.md}.

\section{Integrated Release-Gate Status}

The flagship T1--T8 ladder above remains the governance-critical proof spine,
and the broader defended theorem program now extends through T11.  The
generated summary below reports the older hard-gated release status for the 10
source-backed theorem claims in the precursor plus reference ladder together
with the 8 Appendix~C theorem restatements.  The larger T1--T11 program is
kept aligned by the theorem-surface validator and the appendix proof crosswalk
rather than by that legacy 18-row gate alone.

\IfFileExists{reports/publication/integrated_theorem_gate_summary.tex}{%
  \input{reports/publication/integrated_theorem_gate_summary}%
}{%
  \input{../reports/publication/integrated_theorem_gate_summary}%
}

\IfFileExists{reports/publication/integrated_theorem_gate_matrix.tex}{%
  \input{reports/publication/integrated_theorem_gate_matrix}%
}{%
  \input{../reports/publication/integrated_theorem_gate_matrix}%
}
\endgroup
