% =============================================================================
% Appendix A: Notation Sheet
% ORIUS/DC3S Battery-Safety Thesis
% =============================================================================
\chapter{Notation Sheet}
\label{app:notation}

This appendix collects every symbol used in the ORIUS battery-safety
framework.  Symbols are grouped by role: telemetry and state variables,
quality and drift indicators, uncertainty and action objects, battery
plant parameters, metrics, and tuning constants.

\section{Telemetry and State Variables}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Telemetry and state symbols.}\label{tab:notation-state}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{(continued)}}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endhead
\bottomrule
\endfoot
$\zt$         & $\R^n$     & Clean (ground-truth) telemetry vector at step $t$. \\
$\ztt$        & $\R^n$     & Degraded telemetry vector after dropout, delay, or drift. \\
$\ot$         & $\R^n$     & Observed state passed to the controller. \\
$\xt$         & $\R^n$     & True (latent) physical state; not directly accessible. \\
$\hat{x}_t$   & $\R^n$     & State estimate reconstructed from $\ztt$. \\
$\xi_t$       & $\R^n$     & Observation error: $\xi_t = \hat{x}_t - \xt$. \\
$\epsilon_t$  & $\R$       & One-step model (dynamics) prediction error. \\
\end{longtable}

\section{Quality, Drift, and Sensitivity Indicators}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Quality and detector symbols.}\label{tab:notation-quality}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{(continued)}}\\
\toprule \textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endhead
\bottomrule
\endfoot
$\wt$  & $[0,1]$    & Observation-quality (reliability) score.  Aggregates freshness, consistency, and drift evidence. \\
$\dt$  & $\{0,1\}$  & Drift-detected flag from the Page--Hinkley detector. \\
$\st$  & $\R_+$     & Sensitivity / staleness indicator; large $\st$ implies stale data. \\
\end{longtable}

\section{Uncertainty and Action Objects}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Uncertainty-set and action symbols.}\label{tab:notation-action}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{(continued)}}\\
\toprule \textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endhead
\bottomrule
\endfoot
$\Chat$  & interval   & Base conformal prediction interval at coverage $1-\alpha$. \\
$\Crac$  & interval   & Reliability-adaptive conformal interval after RUI inflation. \\
$\Ut$    & set        & Uncertainty set for robust dispatch: $\Ut = [\hat{x}_t - m_t,\;\hat{x}_t + m_t]$. \\
$m_t$    & $\R_+$     & Half-width (margin) of the uncertainty set: $m_t = q_t/(w_t + \epsilon)$. \\
$\astar$ & $\R^m$     & Candidate action proposed by the base controller. \\
$\asafe$ & $\R^m$     & Repaired action after DC3S safety projection. \\
$\Aset$  & set        & Tightened safe-action set under reliability-aware constraints. \\
$\phi(\cdot)$ & $\{0,1\}$ & Safety predicate: $\phi(x)=1$ iff $x$ satisfies all constraints. \\
\end{longtable}

\section{Battery Plant Parameters}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Battery plant parameters.}\label{tab:notation-battery}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\bottomrule
\endfoot
$\SOC_t$   & $[0, \Emax]$  & State of charge at step $t$ (MWh). \\
$\Emax$    & $\R_+$        & Rated energy capacity (MWh). \\
$\Pmax$    & $\R_+$        & Maximum charge/discharge power (MW). \\
$\Rmax$    & $\R_+$        & Maximum ramp rate (MW/step). \\
$\etac$    & $(0,1]$       & Charging efficiency. \\
$\etad$    & $(0,1]$       & Discharging efficiency. \\
$\Delta t$ & $\R_+$        & Dispatch interval length (hours). \\
\end{longtable}

\section{Metrics}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Evaluation metrics.}\label{tab:notation-metrics}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\bottomrule
\endfoot
$\TSVR$ & $[0,1]$ & True-SOC Violation Rate: fraction of steps with $\SOC_t \notin [\SOC_{\min}, \SOC_{\max}]$. \\
$\IR$    & $[0,1]$ & Intervention Rate: fraction of steps where DC3S repairs the action. \\
$\OASG$ & $[0,1]$ & Observed-safe / Actually-unsafe State Gap rate. \\
$\PICP$ & $[0,1]$ & Prediction-Interval Coverage Probability. \\
\end{longtable}

\section{Tuning Constants and Greek Letters}

\begin{longtable}{@{}clp{8.2cm}@{}}
\caption{Tuning constants and thresholds.}\label{tab:notation-tuning}\\
\toprule
\textbf{Symbol} & \textbf{Domain} & \textbf{Description} \\
\midrule
\endfirsthead
\bottomrule
\endfoot
$\alpha$      & $(0,1)$  & Conformal miscoverage level (target: 0.10). \\
$\kappa_r$    & $\R_+$   & Reliability decay rate in the OQE score. \\
$\kappa_d$    & $\R_+$   & Drift penalty coefficient. \\
$\kappa_s$    & $\R_+$   & Staleness penalty coefficient. \\
$w_{\min}$    & $(0,1)$  & Quality-score floor to prevent division by zero. \\
$q_t$         & $\R_+$   & Conformal quantile at step $t$. \\
$\epsilon$    & $\R_+$   & Small constant in margin denominator: $m_t = q_t/(w_t + \epsilon)$. \\
\end{longtable}

\section{Set Notation and Operators}

\begin{longtable}{@{}cp{10.5cm}@{}}
\caption{Set notation and mathematical operators.}\label{tab:notation-ops}\\
\toprule
\textbf{Symbol} & \textbf{Description} \\
\midrule
\endfirsthead
\bottomrule
\endfoot
$\R, \R_+$           & Real numbers; non-negative reals. \\
$\mathcal{U}$        & Power- and ramp-feasible action set (before safety tightening). \\
$\Prob[\cdot]$       & Probability measure. \\
$\E[\cdot]$          & Expectation operator. \\
$\argmin, \argmax$   & Argument of the minimum / maximum. \\
$\mathbf{1}[\cdot]$  & Indicator function. \\
$|\cdot|$            & Absolute value or set cardinality, depending on context. \\
$\|\cdot\|$          & Euclidean norm unless otherwise stated. \\
\end{longtable}
