\chapter*{References}
\addcontentsline{toc}{chapter}{References}

\begin{thebibliography}{99}

\bibitem{romano2019}
Y.~Romano, E.~Patterson, and R.~J.~Candès,
``Conformalized quantile regression,''
in \emph{Advances in Neural Information Processing Systems}, vol.~32, 2019.

\bibitem{gibbs2021}
I.~Gibbs and E.~J.~Candès,
``Adaptive conformal inference under distribution shift,''
in \emph{Advances in Neural Information Processing Systems}, vol.~34, 2021.

\bibitem{vovk2005}
V.~Vovk, A.~Gammerman, and G.~Shafer,
\emph{Algorithmic Learning in a Random World}.
Springer, 2005.

\bibitem{chen2016}
T.~Chen and C.~Guestrin,
``XGBoost: A scalable tree boosting system,''
in \emph{Proc.\ ACM SIGKDD}, 2016.

\bibitem{ke2017}
G.~Ke \emph{et al.},
``LightGBM: A highly efficient gradient boosting decision tree,''
in \emph{Advances in Neural Information Processing Systems}, vol.~30, 2017.

\bibitem{oreshkin2020}
B.~N.~Oreshkin, D.~Carp\'{o}v, N.~Chapados, and Y.~Bengio,
``N-BEATS: Neural basis expansion analysis for interpretable time series forecasting,''
in \emph{Proc.\ ICLR}, 2020.

\bibitem{lim2021}
B.~Lim, S.~Ar{\i}k, N.~Loeff, and T.~Pfister,
``Temporal Fusion Transformers for interpretable multi-horizon time series forecasting,''
\emph{Int.\ J.\ Forecasting}, vol.~37, no.~4, pp.~1748--1764, 2021.

\bibitem{nie2023}
Y.~Nie \emph{et al.},
``A time series is worth 64 words: Long-term forecasting with transformers,''
in \emph{Proc.\ ICLR}, 2023.

\bibitem{wachi2024}
A.~Wachi, Y.~Sui, Y.~Yue, and M.~Ono,
``Safe reinforcement learning in constrained Markov decision processes,''
in \emph{Proc.\ ICML}, 2024.

\bibitem{amodei2016}
D.~Amodei \emph{et al.},
``Concrete problems in AI safety,''
\emph{arXiv:1606.06565}, 2016.

\bibitem{romano2019conformalized}
Y.~Romano, E.~Patterson, and R.~J.~Candès,
``Conformalized quantile regression,''
in \emph{Advances in Neural Information Processing Systems}, vol.~32, 2019.

\bibitem{gibbs2021adaptive}
I.~Gibbs and E.~J.~Candès,
``Adaptive conformal inference under distribution shift,''
in \emph{Advances in Neural Information Processing Systems}, vol.~34, 2021.

\bibitem{vovk2005algorithmic}
V.~Vovk, A.~Gammerman, and G.~Shafer,
\emph{Algorithmic Learning in a Random World}.
Springer, 2005.

\bibitem{ames2019control}
A.~D.~Ames, S.~Coogan, M.~Egerstedt, G.~Notomista, K.~Sreenath, and P.~Tabuada,
``Control barrier functions: Theory and applications,''
in \emph{Proc.\ European Control Conference (ECC)}, pp.~3420--3431, 2019.

\bibitem{bemporad2007robust}
A.~Bemporad and M.~Morari,
``Robust model predictive control: A survey,''
in \emph{Robustness in Identification and Control}, vol.~245 of \emph{Lecture Notes in Control and Information Sciences}, pp.~207--226. Springer, 2007.

\bibitem{bloem2015shield}
R.~Bloem, B.~K\"onighofer, R.~K\"onighofer, and C.~Wang,
``Shield synthesis: Runtime enforcement for reactive systems,''
in \emph{Proc.\ TACAS}, pp.~533--548, 2015.

\bibitem{cardenas2011attacks}
A.~A.~Cardenas, S.~Amin, Z.-S.~Lin, Y.-L.~Huang, C.-Y.~Huang, and S.~Sastry,
``Attacks against process control systems: Risk assessment, detection, and response,''
in \emph{Proc.\ ACM ASIACCS}, pp.~355--366, 2011.

\bibitem{liang2017review}
Y.~Liang, C.-Z.~Zhao, H.~Yuan, Z.~Chen, W.~Zhang, J.-Q.~Huang, D.~Yu,
Y.~Liu, M.-M.~Titirici, Y.-L.~Chueh, H.~Yu, and Q.~Zhang,
``A review of rechargeable batteries for portable electronic devices,''
\emph{InfoMat}, vol.~1, pp.~6--32, 2019.

\bibitem{shapiro2021lectures}
A.~Shapiro, D.~Dentcheva, and A.~Ruszczy\'nski,
\emph{Lectures on Stochastic Programming: Modeling and Theory}, 3rd~ed.
SIAM, Philadelphia, PA, 2021.

\bibitem{sun2020adversarial}
H.~Sun, J.~Chen, Q.~Ye, and J.~Shao,
``Adversarial attacks on neural networks for logical reasoning,''
\emph{arXiv:2007.13830}, 2020.

\bibitem{xu2020optimal}
C.~Xu and Y.~Xie,
``Conformal prediction interval for dynamic time-series,''
in \emph{Proc.\ ICML}, 2021.

\bibitem{mayne2005robust}
D.~Q.~Mayne, M.~M.~Seron, and S.~V.~Rakovi\'{c},
``Robust model predictive control of constrained linear systems with bounded disturbances,''
\emph{Automatica}, vol.~41, no.~2, pp.~219--224, 2005.

\bibitem{garcia2015comprehensive}
J.~Garc\'{\i}a and F.~Fern\'{a}ndez,
``A comprehensive survey on safe reinforcement learning,''
\emph{Journal of Machine Learning Research}, vol.~16, no.~1, pp.~1437--1480, 2015.

% -----------------------------------------------------------------------
% Conformal Prediction — additional references
% -----------------------------------------------------------------------

\bibitem{angelopoulos2023gentle}
A.~N.~Angelopoulos and S.~Bates,
``Conformal prediction: A gentle introduction,''
\emph{Foundations and Trends in Machine Learning}, vol.~16, no.~1--2, pp.~1--253, 2023.

\bibitem{barber2023beyond}
R.~F.~Barber, E.~J.~Cand\`{e}s, A.~Ramdas, and R.~J.~Tibshirani,
``Conformal prediction beyond exchangeability,''
\emph{Annals of Statistics}, vol.~51, no.~2, pp.~816--845, 2023.

\bibitem{tibshirani2019covariate}
R.~J.~Tibshirani, B.~Efron, R.~F.~Barber, E.~J.~Cand\`{e}s, and A.~Ramdas,
``Conformal prediction under covariate shift,''
in \emph{Advances in Neural Information Processing Systems}, vol.~32, 2019.

\bibitem{papadopoulos2002inductive}
H.~Papadopoulos, K.~Proedrou, V.~Vovk, and A.~Gammerman,
``Inductive confidence machines for regression,''
in \emph{Proc.\ European Conference on Machine Learning (ECML)},
\emph{Lecture Notes in Computer Science}, vol.~2430, pp.~345--356. Springer, 2002.

\bibitem{zaffran2022adaptive}
M.~Zaffran, O.~F\'{e}ron, Y.~Goude, J.~Jolivet, and A.~Dieuleveut,
``Adaptive conformal predictions for time series,''
in \emph{Proc.\ International Conference on Machine Learning (ICML)},
\emph{Proceedings of Machine Learning Research}, vol.~162, pp.~25834--25866, 2022.

\bibitem{stankeviciute2021conformal}
K.~Stankeviciute, A.~M'Charrak, R.~Cornish, and T.~Rainforth,
``Conformal time-series forecasting,''
in \emph{Advances in Neural Information Processing Systems}, vol.~34, 2021.

\bibitem{lei2018distribution}
J.~Lei, M.~G'Sell, A.~Rinaldo, R.~J.~Tibshirani, and L.~Wasserman,
``Distribution-free predictive inference for regression,''
\emph{Journal of the American Statistical Association}, vol.~113, no.~523,
pp.~1094--1111, 2018.

\bibitem{bates2023testing}
S.~Bates, E.~J.~Cand\`{e}s, L.~Lei, Y.~Romano, and M.~Sesia,
``Testing for outliers with conformal p-values,''
\emph{Annals of Statistics}, vol.~51, no.~1, pp.~149--178, 2023.

% -----------------------------------------------------------------------
% Control Barrier Functions and Safety Filters
% -----------------------------------------------------------------------

\bibitem{ames2014cbf}
A.~D.~Ames, J.~W.~Grizzle, and P.~Tabuada,
``Control barrier function based quadratic programs with application to adaptive cruise control,''
in \emph{Proc.\ IEEE Conference on Decision and Control (CDC)}, pp.~6271--6278, 2014.

\bibitem{prajna2004safety}
S.~Prajna and A.~Jadbabaie,
``Safety verification of hybrid systems using barrier certificates,''
in \emph{Hybrid Systems: Computation and Control},
\emph{Lecture Notes in Computer Science}, vol.~2993, pp.~477--492. Springer, 2004.

\bibitem{wabersich2021predictive}
K.~P.~Wabersich and M.~N.~Zeilinger,
``A predictive safety filter for learning-based control of constrained nonlinear dynamical systems,''
\emph{Automatica}, vol.~129, p.~109597, 2021.

\bibitem{taylor2020adaptive}
A.~J.~Taylor and A.~D.~Ames,
``Adaptive safety with control barrier functions,''
in \emph{Proc.\ American Control Conference (ACC)}, pp.~1399--1405, 2020.

\bibitem{wieland2007constructive}
P.~Wieland and F.~Allg\"{o}wer,
``Constructive safety using control barrier functions,''
in \emph{IFAC Proceedings Volumes}, vol.~40, no.~12, pp.~462--467, 2007.

% -----------------------------------------------------------------------
% Robust and Tube MPC
% -----------------------------------------------------------------------

\bibitem{langson2004tubes}
W.~Langson, I.~Chryssochoos, S.~V.~Rako\v{v}i\'{c}, and D.~Q.~Mayne,
``Robust model predictive control using tubes,''
\emph{Automatica}, vol.~40, no.~1, pp.~125--133, 2004.

\bibitem{rawlings2017mpc}
J.~B.~Rawlings, D.~Q.~Mayne, and M.~M.~Diehl,
\emph{Model Predictive Control: Theory, Computation, and Design}, 2nd~ed.
Nob Hill Publishing, Madison, WI, 2017.

\bibitem{rakovic2005invariant}
S.~V.~Rako\v{v}i\'{c}, E.~C.~Kerrigan, K.~I.~Kouramas, and D.~Q.~Mayne,
``Invariant approximations of the minimal robust positively invariant set,''
\emph{IEEE Transactions on Automatic Control}, vol.~50, no.~3, pp.~406--410, 2005.

\bibitem{lorenzen2019constraint}
M.~Lorenzen, F.~Dabbene, R.~Tempo, and F.~Allg\"{o}wer,
``Constraint-tightening and stability in stochastic model predictive control,''
\emph{IEEE Transactions on Automatic Control}, vol.~64, no.~11, pp.~4625--4637, 2019.

% -----------------------------------------------------------------------
% Safe Reinforcement Learning
% -----------------------------------------------------------------------

\bibitem{achiam2017constrained}
J.~Achiam, D.~Held, A.~Tamar, and P.~Abbeel,
``Constrained policy optimization,''
in \emph{Proc.\ International Conference on Machine Learning (ICML)},
\emph{Proceedings of Machine Learning Research}, vol.~70, pp.~22--31, 2017.

\bibitem{brunke2022safe}
L.~Brunke, M.~Greeff, A.~W.~Hall, Z.~Yuan, S.~Zhou, J.~Panerati, and A.~P.~Schoellig,
``Safe learning in robotics: From learning-based control to safe reinforcement learning,''
\emph{Annual Review of Control, Robotics, and Autonomous Systems}, vol.~5, pp.~411--444, 2022.

\bibitem{moldovan2012safe}
T.~M.~Moldovan and P.~Abbeel,
``Safe exploration in Markov decision processes,''
in \emph{Proc.\ International Conference on Machine Learning (ICML)}, pp.~1451--1458, 2012.

\bibitem{ray2019benchmarking}
A.~Ray, J.~Achiam, and D.~Amodei,
``Benchmarking safe exploration in deep reinforcement learning,''
\emph{arXiv preprint arXiv:1910.01708}, 2019.

\bibitem{altman1999constrained}
E.~Altman,
\emph{Constrained Markov Decision Processes}.
Chapman \& Hall/CRC, Boca Raton, FL, 1999.

% -----------------------------------------------------------------------
% Runtime Verification and Monitoring
% -----------------------------------------------------------------------

\bibitem{falcone2012tutorial}
Y.~Falcone, K.~Havelund, and G.~Reger,
``A tutorial on runtime verification,''
in \emph{Engineering Dependable Software Systems},
NATO Science for Peace and Security Series, pp.~141--175. IOS Press, 2013.

\bibitem{bartocci2018specification}
E.~Bartocci, Y.~Falcone, A.~Francalanza, and G.~Reger,
``Introduction to runtime verification,''
in \emph{Lectures on Runtime Verification}, \emph{Lecture Notes in Computer Science},
vol.~10457, pp.~1--33. Springer, 2018.

\bibitem{pnueli2006psl}
A.~Pnueli and A.~Zaks,
``PSL model checking and run-time verification via testers,''
in \emph{Formal Methods (FM 2006)}, \emph{Lecture Notes in Computer Science},
vol.~4085, pp.~573--586. Springer, 2006.

\bibitem{leucker2009brief}
M.~Leucker and C.~Schallhart,
``A brief account of runtime verification,''
\emph{Journal of Logic and Algebraic Programming}, vol.~78, no.~5, pp.~293--303, 2009.

\bibitem{maler2004monitoring}
O.~Maler and D.~Nickovic,
``Monitoring temporal properties of continuous signals,''
in \emph{Formal Techniques, Modelling and Analysis of Timed and Fault-Tolerant Systems},
\emph{Lecture Notes in Computer Science}, vol.~3253, pp.~152--166. Springer, 2004.

% -----------------------------------------------------------------------
% Cyber-Physical Systems Security
% -----------------------------------------------------------------------

\bibitem{teixeira2015secure}
A.~Teixeira, I.~Shames, H.~Sandberg, and K.~H.~Johansson,
``A secure control framework for resource-limited adversaries,''
\emph{Automatica}, vol.~51, pp.~135--148, 2015.

\bibitem{urbina2016limiting}
D.~I.~Urbina, J.~A.~Giraldo, A.~A.~Cardenas, N.~O.~Tippenhauer, J.~Valente,
M.~Faisal, J.~Ruths, R.~Candell, and H.~Sandberg,
``Limiting the impact of stealthy attacks on industrial control systems,''
in \emph{Proc.\ ACM SIGSAC Conference on Computer and Communications Security (CCS)},
pp.~1092--1105, 2016.

\bibitem{mo2014secure}
Y.~Mo and B.~Sinopoli,
``Secure estimation in the presence of integrity attacks,''
\emph{IEEE Transactions on Automatic Control}, vol.~60, no.~4, pp.~1145--1151, 2015.

\bibitem{giraldo2018survey}
J.~A.~Giraldo, D.~Urbina, A.~A.~Cardenas, J.~Valente, M.~Faisal, J.~Ruths,
N.~O.~Tippenhauer, H.~Sandberg, and R.~Candell,
``A survey of physics-based attack detection in cyber-physical systems,''
\emph{ACM Computing Surveys}, vol.~51, no.~4, pp.~76:1--76:36, 2018.

% -----------------------------------------------------------------------
% Battery Systems and Energy Storage
% -----------------------------------------------------------------------

\bibitem{hu2012comparative}
X.~Hu, S.~Li, and H.~Peng,
``A comparative study of equivalent circuit models for Li-ion batteries,''
\emph{Journal of Power Sources}, vol.~198, pp.~359--367, 2012.

\bibitem{tarascon2001issues}
J.~M.~Tarascon and M.~Armand,
``Issues and challenges facing rechargeable lithium batteries,''
\emph{Nature}, vol.~414, pp.~359--367, 2001.

\bibitem{dunn2011electrical}
B.~Dunn, H.~Kamath, and J.-M.~Tarascon,
``Electrical energy storage for the grid: A battery of choices,''
\emph{Science}, vol.~334, no.~6058, pp.~928--935, 2011.

\bibitem{bose2019stochastic}
S.~Bose, S.~Fattahi, and A.~Swami,
``Stochastic battery operation in a competitive electricity market,''
\emph{IEEE Transactions on Power Systems}, vol.~34, no.~6, pp.~4809--4817, 2019.

\bibitem{rosewater2019battery}
D.~M.~Rosewater, D.~A.~Copp, T.~A.~Nguyen, R.~H.~Byrne, and S.~Santoso,
``Battery energy storage models for optimal control,''
\emph{IEEE Access}, vol.~7, pp.~178497--178506, 2019.

% -----------------------------------------------------------------------
% Probabilistic Forecasting and Uncertainty Quantification
% -----------------------------------------------------------------------

\bibitem{hong2016probabilistic}
T.~Hong and S.~Fan,
``Probabilistic electric load forecasting: A tutorial review,''
\emph{International Journal of Forecasting}, vol.~32, no.~3, pp.~914--938, 2016.

\bibitem{gneiting2007strictly}
T.~Gneiting and A.~E.~Raftery,
``Strictly proper scoring rules, prediction, and estimation,''
\emph{Journal of the American Statistical Association}, vol.~102, no.~477,
pp.~359--378, 2007.

\bibitem{makridakis2020m4}
S.~Makridakis, E.~Spiliotis, and V.~Assimakopoulos,
``The M4 Competition: 100,000 time series and 61 forecasting methods,''
\emph{International Journal of Forecasting}, vol.~36, no.~1, pp.~54--74, 2020.

\bibitem{taylor2003electricity}
J.~W.~Taylor,
``Short-term electricity demand forecasting using double seasonal exponential smoothing,''
\emph{Journal of the Operational Research Society}, vol.~54, no.~8, pp.~799--805, 2003.

% -----------------------------------------------------------------------
% Autonomous Vehicle Safety
% -----------------------------------------------------------------------

\bibitem{koopman2016av}
P.~Koopman and M.~Wagner,
``Autonomous vehicle safety: An interdisciplinary challenge,''
\emph{IEEE Intelligent Transportation Systems Magazine}, vol.~8, no.~1, pp.~90--96, 2016.

\bibitem{shalev2017formal}
S.~Shalev-Shwartz, S.~Shammah, and A.~Shashua,
``On a formal model of safe and scalable self-driving cars,''
\emph{arXiv preprint arXiv:1708.06374}, 2017.

\bibitem{althoff2014online}
M.~Althoff and J.~M.~Dolan,
``Online verification of automated road vehicles using reachability analysis,''
\emph{IEEE Transactions on Robotics}, vol.~30, no.~4, pp.~903--918, 2014.

\bibitem{nolte2017assuring}
T.~Nolte, M.~Sjodin, K.~Akesson, K.~Lundqvist, M.~Behnam, M.~Lindberg,
and C.~Normann,
``Towards bringing research on functional safety in automotive systems to practice,''
\emph{IEEE Transactions on Industrial Informatics}, vol.~13, no.~1, pp.~338--349, 2017.

% -----------------------------------------------------------------------
% Medical Device Safety and ICU Monitoring
% -----------------------------------------------------------------------

\bibitem{lee2010medical}
I.~Lee and O.~Sokolsky,
``Medical cyber physical systems,''
in \emph{Proc.\ ACM/IEEE Design Automation Conference (DAC)}, pp.~743--748, 2010.

\bibitem{saeed2011mimic}
M.~Saeed, M.~Villarroel, A.~T.~Reisner, G.~Clifford, L.~W.~Lehman, G.~Moody,
T.~Heldt, T.~H.~Kyaw, B.~Moody, and R.~G.~Mark,
``Multiparameter intelligent monitoring in intensive care II (MIMIC-II):
A public-access intensive care unit database,''
\emph{Critical Care Medicine}, vol.~39, no.~5, pp.~952--960, 2011.

\bibitem{clifford2012robust}
G.~D.~Clifford, W.~J.~Long, G.~B.~Moody, and P.~Szolovits,
``Robust parameter extraction for decision support using multimodal intensive care data,''
\emph{Philosophical Transactions of the Royal Society A}, vol.~367, no.~1887,
pp.~411--429, 2009.

\bibitem{fda2019samd}
U.S.\ Food and Drug Administration,
``Software as a medical device (SaMD): Clinical evaluation,''
FDA Guidance Document, 2019.

% -----------------------------------------------------------------------
% Aerospace Safety and Certification Standards
% -----------------------------------------------------------------------

\bibitem{rtca2011do178c}
RTCA,
\emph{DO-178C: Software Considerations in Airborne Systems and Equipment Certification}.
RTCA, Washington, D.C., 2011.

\bibitem{rushby1993formal}
J.~Rushby,
``Formal methods and the certification of critical systems,''
Technical Report SRI-CSL-93-7, Computer Science Laboratory,
SRI International, Menlo Park, CA, 1993.

\bibitem{habli2010model}
I.~Habli and T.~Kelly,
``Model-based assurance of control system safety and security,''
in \emph{Proc.\ International Symposium on Software Reliability Engineering (ISSRE)},
pp.~31--40, 2010.

% -----------------------------------------------------------------------
% Partial Observability and State Estimation
% -----------------------------------------------------------------------

\bibitem{kaelbling1998planning}
L.~P.~Kaelbling, M.~L.~Littman, and A.~R.~Cassandra,
``Planning and acting in partially observable stochastic domains,''
\emph{Artificial Intelligence}, vol.~101, no.~1--2, pp.~99--134, 1998.

\bibitem{julier2004unscented}
S.~J.~Julier and J.~K.~Uhlmann,
``Unscented filtering and nonlinear estimation,''
\emph{Proceedings of the IEEE}, vol.~92, no.~3, pp.~401--422, 2004.

\bibitem{rockafellar2000optimization}
R.~T.~Rockafellar and S.~Uryasev,
``Optimization of conditional value-at-risk,''
\emph{Journal of Risk}, vol.~2, no.~3, pp.~21--42, 2000.

% -----------------------------------------------------------------------
% Industrial Process Control
% -----------------------------------------------------------------------

\bibitem{qin2003survey}
S.~J.~Qin and T.~A.~Badgwell,
``A survey of industrial model predictive control technology,''
\emph{Control Engineering Practice}, vol.~11, no.~7, pp.~733--764, 2003.

\bibitem{yin2014review}
S.~Yin, S.~X.~Ding, X.~Xie, and H.~Luo,
``A review on basic data-driven approaches for industrial process monitoring,''
\emph{IEEE Transactions on Industrial Electronics}, vol.~61, no.~11, pp.~6418--6428, 2014.

\bibitem{isermann2005model}
R.~Isermann,
``Model-based fault-detection and diagnosis---status and applications,''
\emph{Annual Reviews in Control}, vol.~29, no.~1, pp.~71--85, 2005.

% -----------------------------------------------------------------------
% Anomaly Detection
% -----------------------------------------------------------------------

\bibitem{liu2008isolation}
F.~T.~Liu, K.~M.~Ting, and Z.-H.~Zhou,
``Isolation forest,''
in \emph{Proc.\ IEEE International Conference on Data Mining (ICDM)},
pp.~413--422, 2008.

\bibitem{chandola2009anomaly}
V.~Chandola, A.~Banerjee, and V.~Kumar,
``Anomaly detection: A survey,''
\emph{ACM Computing Surveys}, vol.~41, no.~3, pp.~15:1--15:58, 2009.

% -----------------------------------------------------------------------
% Fault Detection and Diagnosis
% -----------------------------------------------------------------------

\bibitem{tsybakov2009introduction}
A.~B.~Tsybakov,
\emph{Introduction to Nonparametric Estimation}.
Springer, New York, 2009.

\bibitem{yu1997assouad}
B.~Yu,
``Assouad, Fano, and Le~Cam,''
in \emph{Festschrift for Lucien Le~Cam}, D.~Pollard, E.~Torgersen,
and G.~L.~Yang, Eds.
Springer, New York, 1997, pp.~423--435.

\bibitem{blanke2006diagnosis}
M.~Blanke, M.~Kinnaert, J.~Lunze, and M.~Staroswiecki,
\emph{Diagnosis and Fault-Tolerant Control}, 2nd~ed.
Springer, Berlin, 2006.

\bibitem{zhang2008bibliographical}
Y.~Zhang and J.~Jiang,
``Bibliographical review on reconfigurable fault-tolerant control systems,''
\emph{Annual Reviews in Control}, vol.~32, no.~2, pp.~229--252, 2008.

\bibitem{lindemann2023conformal}
L.~Lindemann, M.~Cleaveland, G.~Shim, and G.~J.~Pappas,
``Safe planning in dynamic environments using conformal prediction,''
\emph{IEEE Robotics and Automation Letters}, vol.~8, no.~8,
pp.~5116--5123, 2023.

\bibitem{dixit2023adaptive}
A.~Dixit, L.~Lindemann, S.~X.~Wei, M.~Cleaveland, G.~J.~Pappas,
and J.~V.~Deshmukh,
``Adaptive conformal prediction for motion planning among dynamic agents,''
in \emph{Proc.\ IEEE/RSJ IROS}, Detroit, MI, 2023, pp.~5765--5772.

\bibitem{angelopoulos2023conformal}
A.~N.~Angelopoulos, S.~Bates, E.~J.~Cand\`{e}s, M.~I.~Jordan,
and L.~Lei,
``Conformal risk control,''
in \emph{Proc.\ ICLR}, 2024.

\end{thebibliography}
